
# Normal

```{r}
theme_grey
```

```{r}
set.seed(123)
N <- 100
xx <- rnorm(N)
yy <- rnorm(N, xx)

dat <- tibble(xx, yy)
```

```{r}

p1 <- ggplot(dat, aes(x = xx, y = yy)) +
  geom_point() +
  theme_grey()

p1

```


```{r}
p1 + theme(axis.title.x =  element_text(size = rel(2)))
```



# upload

sudo scp -r -i ~/.ssh/id_rsa ~/Dropbox/MS/LMAms/data/nature02403-s2.csv mattocci@178.128.60.227:/home/mattocci/moge.csv

sudo scp -r -i ~/.ssh/id_rsa ~/Dropbox/MS/LMAms/data/nature02403-s2.csv mattocci@178.128.60.227:/home/mattocci/moge2.csv


# download

sudo scp -r -i ~/.ssh/id_rsa mattocci@178.128.60.227:/home/mattocci/moge.csv ~/Desktop/mogemoge.csv

sudo scp -r -i ~/.ssh/id_rsa mattocci@178.128.60.227:/home/mattocci/LMAms/data/GLpara.csv ~/Desktop/GLpara.csv
sudo scp -r -i ~/.ssh/id_rsa mattocci@178.128.60.227:/home/mattocci/LMAms/data/PApara.csv ~/Desktop/PApara.csv


sudo scp -r -i ~/.ssh/id_rsa mattocci@178.128.60.227:/home/mattocci/LMAms/fig_code/fig.html ~/Desktop/fig.html

sudo scp -r -i ~/.ssh/id_rsa mattocci@178.128.60.227:/home/mattocci/LMAms/data.zip ~/Desktop/data.zip


sudo scp -r -i ~/.ssh/id_rsa mattocci@178.128.60.227:/home/mattocci/LMAms/data/GL_LMAms_obs.rda ~/Dropbox/MS/LMAms/data/GL_LMAms_obs_do.rda 

sudo scp -r -i ~/.ssh/id_rsa mattocci@178.128.60.227:/home/mattocci/LMAms/data/PA_LMAms_L0_more_obs.rda ~/Dropbox/MS/LMAms/data/PA_LMAms_L0_more_obs_do.rda



sudo scp -r -i ~/.ssh/id_rsa mattocci@206.189.155.122 'cat /home/mattocci/.ssh/id_rsa.pub' | pbcopy

git clone git@github.com:mattocci27/LMAms.git ~/LMAms

GL_LMA_CV_more_obs.rda
GL_LMA_CV_obs.rda
GL_LMA_CV_obs_cv.rda
GL_LMA_more_obs.rda
GL_LMA_obs.rda
GL_LMAms_CV_obs.rda
GL_LMAms_CV_obs_cv.rda
GL_LMAms_more_obs.rda
GL_LMAms_obs.rda
PA_LD_L_CV_obs_cv.rda
PA_LMA_CV_obs_cv.rda
PA_LMA_L_CV_obs_cv.rda
PA_LMAms_CV_obs_cv.rda
PA_LMAms_L0_more_obs.rda
PA_LMAms_L_CV_obs_cv.rda
PA_LMAms_L_more_obs.rda
PA_LMAms_more_obs.rda



ssh -i ~/.ssh/id_rsa mattocci@178.128.60.227 'cat /home/mattocci/.ssh/id_rsa.pub' | pbcopy


git clone git@github.com:mattocci27/LMAms.git ~/LMAms

git clone git@github.com:mattocci27/LMAms.git ~/ufrc/LMAms

ssh-keygen -t rsa -b 4096 -C "mattocci27@gmail.com"



ssh -i ma@178.128.60.227 'cat /home/mattocci/.ssh/id_rsa.pub' | pbcopy

```

OBS=obs
DATA=PA
for MODEL in PA_LMAms0 PA_LMAms

do
  for OBS in obs
  do
R --vanilla --slave --args ${MODEL} ${DATA} 5000 4000 1 ${OBS} < ./model/model_more.r > ./log/${MODEL}_${OBS}.log
  done
done


OBS=obs
DATA=GL
for MODEL in GL_LMAs
do
  for OBS in obs
  do
R --vanilla --slave --args ${MODEL} ${DATA} 5000 4000 1 ${OBS} < ./model/model.r > ./log/${MODEL}_${OBS}.log
  done
done
```


```{r}

a1 <- 5
a2 <- 2

x <- seq(1,50, length = 100)
y <- a1 * x / (a2 + x)

par(mfrow = c(2,2))
plot(x,y)
plot(1/x,1/y)
plot(x,y, log = "xy")
par(mfrow = c(1,1))

lm(y~x) %>% summary
lm(log(y)~log(x)) %>% summary

xx <- seq(0, 1, length = 1000)
par(mfrow = c(2,2))
yy <- dbeta(xx, 1, 1, log = F)
plot(xx, yy, type = "l")
yy <- dbeta(xx, 0.5, 0.5, log = F)
plot(xx, yy, type = "l")
yy <- dbeta(xx, 2, 2, log = F)
plot(xx, yy, type = "l")
yy <- dbeta(xx, 10, 5, log = F)
plot(xx, yy, type = "l")

```

