---
title: GLOPNET results using normal distribution (potential)
author: Masatoshi Katabuchi
date: "`r format(Sys.time(), '%B %d, %Y')`"
---

Note: Still using a name "LMAp". .

\begin{align}
  &E[\mathrm{net \; photosynthesis}]
  = E[A_{\mathrm{area} \; i}]
  = \alpha_0\mathrm{LMAp}^{\alpha_p} \\
  &E[\mathrm{LL}_{i}] = \beta_{0} \mathrm{LMAs}_i^{\beta_s} \\
  &E[R_{\mathrm{area} \; i}] = r_0 \mathrm{LMAp}^{r_p}_{i} \mathrm{LMAs}^{r_s}_{i}
\end{align}


```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=TRUE,
                      warning=FALSE,
                      message=FALSE)
```


```{r, echo = F, message = F}
library(loo)
library(rstan)
library(dplyr)
library(scales)
library(memisc)


theme_set(theme_light())

settings <- yaml::yaml.load_file("../settings.yml")
```


```{r, echo = F, message = F}

#load("./data/GL_sitePL_LD_L_MVN_obs.rda")
#load("./data/PA_LMAm_LMAsLT2_obs.rda")
#summary_GL_LDL <- data.frame(summary(res)$summary)

load("../data/GL_m0_std_obs.rda")

pairs(res, pars = c("beta[1]", "beta[2]", "lp__"))

GL_summary <- data.frame(summary(res)$summary) %>%
  round(3)
DT::datatable(GL_summary)
GL <- dat 
P_vec <- paste("p[", 1:nrow(GL), "]" ,sep = "")

GL <- GL %>%
  mutate(DE = ifelse(GL$DE == "", "U", as.character(DE))) 


#LMA <- rlnorm(100, log(80), 1)
#p <- rbeta(100, 40, 5)
#p <- rbeta(100, 10, 5)
#p2 <- 1-p
#
#plot(LMA*p, LMA*p2, log = "xy")

mu2_vec <- paste("mu2[", 1:nrow(GL), "]", sep = "")
#Mu_vec <- paste("mu[", 1:nrow(GL), ",2]", sep = "")
temp_LMAp <- GL_summary[P_vec, "mean"] * GL$LMA
temp_LMAs <- GL$LMA - temp_LMAp
temp_LMAp <- rstan::extract(res, "log_LMAp_ori")[[1]] %>% exp
temp_LMAs <- rstan::extract(res, "log_LMAs_ori")[[1]] %>% exp
temp_LL <- exp(GL_summary[mu2_vec, "mean"])
#temp_LL <- rstan::extract(res, "mu2")[[1]][5,] %>% exp

GL <- GL %>%
  mutate(LMAp = apply(temp_LMAp, 2, median)) %>%
  mutate(LMAp_lo = apply(temp_LMAp, 2, function(x)quantile(x, 0.025))) %>%
  mutate(LMAp_up = apply(temp_LMAp, 2, function(x)quantile(x, 0.975))) %>%
  mutate(LMAs = apply(temp_LMAs, 2, median)) %>%
  mutate(LMAs_lo = apply(temp_LMAs, 2, function(x)quantile(x, 0.025))) %>%
  mutate(LMAs_up = apply(temp_LMAs, 2, function(x)quantile(x, 0.975))) %>%
  mutate(preLL = temp_LL) 


preLL <- rstan::extract(res, "mu2")[[1]]

tmp <- GL %>%
  dplyr::select(sp, DE) 

d <- read_csv("../data/nature02403-s2.csv", skip = 10)

dd <- data.frame(Narea = 10^d$`log Narea`,
        Parea = 10^d$`log Parea`,
        sp = d$Species) %>%
        group_by(sp) %>%
        summarize(Narea = mean(Narea, na.omit = T),
            Parea = mean(Parea, na.omit = T))

GL <- left_join(GL, dd, by = "sp") %>%
  mutate(gr = factor(DE,
    labels = c("Deciduous",
               "Evergreen",
               "Unclassified"
                      ))) %>%
  arrange(desc(gr))

```


```{r, echo = F, eval = T}
waic(extract_log_lik(res,"log_lik"))$waic
loo(extract_log_lik(res,"log_lik"))
moge <- loo(extract_log_lik(res,"log_lik")) 
plot(moge)
pareto_k_ids(moge, threshold = 0.3)
```

# Paramters 

```{r, echo = F, eval = T}

plot(res)

```

```{r, echo = F}

var_p <- function(data)cov(data$LMA, data$LMAp)
var_s <- function(data)cov(data$LMA, data$LMAs)
var_t <- function(data)var(data$LMA)

var_p <- function(data)var(data$LMAp)
var_s <- function(data)var(data$LMAs)
cov_ps <- function(data)cov(data$LMAp, data$LMAs)
var_t <- function(data)var(data$LMA)

var_dat <- GL %>%
  mutate(site = "GL") %>%
  group_by(site) %>%
  nest %>%
  mutate(var_LMAm = map_dbl(data, var_p)) %>%
  mutate(var_LMAs = map_dbl(data, var_s)) %>%
  mutate(cov_ms = 2 * map_dbl(data, cov_ps)) %>%
  mutate(var_total = map_dbl(data, var_t)) %>%
  dplyr::select(-data)

var_dat_s <- var_dat %>%
  mutate(LMAm_variance = var_LMAm / (var_LMAm + var_LMAs) * 100) %>%
  mutate(LMAs_variance = var_LMAs / (var_LMAm + var_LMAs) * 100) %>%
  gather(var, val, c(LMAm_variance, LMAs_variance))

var_dat_s2 <- var_dat %>%
  mutate(LMAm_variance = var_LMAm / (var_total) * 100) %>%
  mutate(LMAs_variance = var_LMAs / (var_total) * 100) %>%
  mutate(cov_LMAm_LMAs = cov_ms / (var_total) * 100) %>%
  gather(var, val, c(LMAm_variance, LMAs_variance, cov_LMAm_LMAs)) 


```



```{r, echo = F, eval = F}
# R values

#R values are based on posterior distributions of LMAs

par(mfrow = c(1,2))
apply(log_LMAs, 1, function(x)cor(x, log(GL$Aarea))) %>% hist(main = "LMAs vs Aarea, R")
apply(log_LMAs, 1, function(x)cor(x, log(GL$Aarea))^2) %>% hist(main = "LMAs vs Aarea, R2")

apply(log_LMAs, 1, function(x)cor(x, log(GL$Aarea + GL$Rarea))) %>% hist(main = "LMAs vs Aarea + Rarea, R")
apply(log_LMAs, 1, function(x)cor(x, log(GL$Aarea + GL$Rarea))^2) %>% hist(main = "LMAs vs Aarea + Rarea, R2")


apply(log_LDs, 1, function(x)cor(x, log(GL$Aarea))) %>% hist(main = "LMAs/LT vs Aarea, R")
apply(log_LDs, 1, function(x)cor(x, log(GL$Aarea))^2) %>% hist(main = "LMAs/LT vs Aarea, R2")


apply(log_LDs, 1, function(x)cor(x, log(GL$Aarea + GL$Rarea))) %>% hist(main = "LMAs/LT vs Aarea + Rarea, R")
apply(log_LDs, 1, function(x)cor(x, log(GL$Aarea + GL$Rarea))^2) %>% hist(main = "LMAs/LT vs Aarea + Rarea, R2")
```


# Trait correlations

## A~area~, R~area~, LL

Open symbols indicate sun leaves and closed symbols indicate shade leaves. The results look nice and do not alter main conclusions.

```{r, echo = F, message = F}

GL2 <- GL %>%
  gather(LMA, val, c("LMA", "LMAp", "LMAs")) %>%
  gather(trait, val2, c("Aarea", "LL", "Rarea")) 

GL_lo <- GL %>%
  gather(LMA , lo, c("LMAp_lo", "LMAs_lo")) %>%
  dplyr::select(sp, lo, LMA) %>%
  mutate(LMA = ifelse(LMA == "LMAp_lo", "LMAp", "LMAs") )
GL_up <- GL %>%
  gather(LMA , up, c("LMAp_up", "LMAs_up")) %>%
  dplyr::select(sp, up, LMA) %>%
  mutate(LMA = ifelse(LMA == "LMAp_up", "LMAp", "LMAs") )

GL2 <- GL2 %>%
  full_join(., GL_lo, by = c("sp", "LMA")) %>%
  full_join(., GL_up, by = c("sp", "LMA")) 

GL3 <- GL %>%
  gather(LMA, val, c("LMA", "LMAp", "LMAs")) %>%
  #gather(trait, val2, c("Aarea", "LL", "Rarea")) %>%
  gather(trait2, val3, c("Narea","Parea")) 

fills <- c("Deciduous" = settings$fills$D,
          "Evergreen" = settings$fills$E,
          #"Unclassified" = settings$fills$U)
          "Unclassified" = "grey")

cols <- c("Deciduous" = settings$colors$D,
          "Evergreen" = settings$colors$E,
          "Unclassified" = settings$colors$U)

ggplot(GL2, aes(x = val, y = val2, fill = gr, col = gr)) +
  geom_errorbarh(aes(xmin = lo, xmax=up, colour = gr), alpha = 0.8) +
  geom_point(shape = 21, col = "black") +
  facet_grid(trait ~ LMA, scale = "free") +
  xlab("LMA") +
  ylab("Traits") +
  scale_fill_manual(values = fills) +
  scale_colour_manual(values = fills) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))

```

## Observed LL vs predicted LL

Note: Light and site effects were not included in this model.

```{r, echo = F, message = F}

ggplot(GL, aes(x = LL, y = preLL, fill = gr, col = gr)) +
  geom_point(shape = 21) +
  xlab("obsLL") +
  ylab("preLL") +
  scale_fill_manual(values = fills) +
  scale_colour_manual(values = cols) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))


cor.test(log(GL$LL), log(GL$preLL))


```


##  N~area~, P~area~

```{r, echo = F, message = F}

ggplot(GL3, aes(x = val, y = val3, fill = gr, col = gr)) +
  geom_point(shape = 21) +
  facet_grid(trait2 ~ LMA, scale = "free") +
  xlab("LMA") +
  ylab("Traits") +
  scale_fill_manual(values = fills) +
  scale_colour_manual(values = cols) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))

cor.test(log(GL$LMA), log(GL$Narea))
cor.test(log(GL$LMAp), log(GL$Narea))

cor.test(log(GL$LMA), log(GL$Parea))
cor.test(log(GL$LMAp), log(GL$Parea))


```

## LMAp vs LMAs

```{r, echo = F, message = F}
ggplot(GL, aes(x = LMAp, y = LMAs, fill = gr, col = gr)) +
  geom_point(shape = 21) +
  xlab("LMAm") +
  ylab("LMAs") +
  scale_fill_manual(values = fills) +
  scale_colour_manual(values = cols) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))
```

## Trait means

The results look nice and do not alter main conclusions.

```{r, echo = F, message = F}

ggplot(GL2, aes(x = gr, y = val,
                                         fill = gr,
                                         col = gr)) +
  geom_boxplot() +
  facet_grid(.~LMA, scale = "free") +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_fill_manual(values = fills) +
  scale_colour_manual(values = cols) 


```


# Decomposition of LMA variation 

1) var(LMAm) & var(LMAs) & 2*cov(LMAm, LMAs), 2)relative contribution: var(LMAm)/(var(LMAm) + var(LMAs)) & var(LMAs) /(var(LMAm) + var(LMAs))

```{r, echo = F}

ggplot(var_dat_s2, aes(x = site, y = val, fill = var)) +
  geom_col() +
  ylab("val (%)") 

ggplot(var_dat_s, aes(x = site, y = val, fill = var)) +
  geom_bar(position = "fill",stat = "identity") +
  scale_y_continuous(labels = percent_format())

```


```{r, echo = F, message = F, eval = F}

cor.test(log(GL$LMA), log(GL$cell_area))
cor.test(log(GL$LMAs), log(GL$cell_area))

cor.test(log(GL$LD), log(GL$cell_vol))
cor.test(log(GL$LDs), log(GL$cell_vol))


cor.test(log(GL$LMA), log(GL$Narea))
cor.test(log(GL$LMAp), log(GL$Narea))

cor.test(log(GL$LMA), log(GL$Parea))
cor.test(log(GL$LMAp), log(GL$Parea))

apply(log_LMAp, 1, function(x)cor(x, log(GL$Narea))) %>% hist(main = "LMAp vs Narea, R")
apply(log_LMAp, 1, function(x)cor(x, log(GL$Parea))) %>% summary

apply(log_LMAp, 1, function(x)cor(x, log(GL$Narea))) %>% hist(main = "LMAp vs Narea, R")

apply(log_LMAp, 1, function(x)cor(x, log(GL$Rarea))) %>% quantile(0.025)
cor.test(log(GL$LMA), log(GL$Rarea))

apply(log_LMAs, 1, function(x)cor(x, log(GL$cell_area))) %>% hist(main = "LMAp vs Narea, R")

```


```{r, echo = F, message = F, eval = F}

ggplot(tmp3, aes(x = LD_shade, y = LD_sun)) +
  geom_point() +
  facet_grid(~site) +
  geom_abline(slope=1, lty = 2) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  coord_fixed()


ggplot(tmp3, aes(x = cell_vol_shade, y = cell_vol_sun)) +
  geom_point() +
  facet_grid(~site) +
  geom_abline(slope=1, lty = 2) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  coord_fixed()


ggplot(GL, aes(x = site_strata, y = cell_vol,
               fill = site_strata,
               col = site_strata)) +
  geom_boxplot() +
  #geom_violin() +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_fill_manual(values = c("DRY_CAN" = "white",
                               "WET_CAN" = "white",
                               "DRY_UNDER" = "#8B7355",
                               "WET_UNDER" = "#1874CD"
                               )) +
  scale_colour_manual(values = c("DRY_CAN" = "#8B7355",
                               "WET_CAN" = "#1874CD",
                               "DRY_UNDER" = "black",
                               "WET_UNDER" = "black"
                               ))

ggplot(GL, aes(x = site_strata, y = LD,
               fill = site_strata,
               col = site_strata)) +
  geom_boxplot() +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_fill_manual(values = c("DRY_CAN" = "white",
                               "WET_CAN" = "white",
                               "DRY_UNDER" = "#8B7355",
                               "WET_UNDER" = "#1874CD"
                               )) +
  scale_colour_manual(values = c("DRY_CAN" = "#8B7355",
                               "WET_CAN" = "#1874CD",
                               "DRY_UNDER" = "black",
                               "WET_UNDER" = "black"
                               ))



ggplot(GL, aes(x = site_strata, y = cell_area,
               fill = site_strata,
               col = site_strata)) +
  #geom_violin() +
  geom_boxplot() +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_fill_manual(values = c("DRY_CAN" = "white",
                               "WET_CAN" = "white",
                               "DRY_UNDER" = "#8B7355",
                               "WET_UNDER" = "#1874CD"
                               )) +
  scale_colour_manual(values = c("DRY_CAN" = "#8B7355",
                               "WET_CAN" = "#1874CD",
                               "DRY_UNDER" = "black",
                               "WET_UNDER" = "black"
                               ))


ggplot(GL, aes(x = LD, y = Aarea, fill = site, col = site, shape = strata)) +
  geom_point()  +
  facet_grid(site ~ strata)  +
  scale_shape_manual(values = c("CAN" = 1, "UNDER" = 21)) +
  scale_fill_manual(values = c("PNM" = "#8B7355", "PNSL" = "#1874CD")) +
  scale_colour_manual(values = c("PNM" = "#8B7355", "PNSL" = "#1874CD")) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))

ggplot(GL, aes(x = LD, y = Aarea, fill = site, col = site, shape = strata)) +
  geom_point()  +
  facet_grid( ~ strata)  +
  scale_shape_manual(values = c("CAN" = 1, "UNDER" = 21)) +
  scale_fill_manual(values = c("PNM" = "#8B7355", "PNSL" = "#1874CD")) +
  scale_colour_manual(values = c("PNM" = "#8B7355", "PNSL" = "#1874CD")) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))


ggplot(GL, aes(x = cell_vol, y = Aarea, fill = site, col = site, shape = strata)) +
  geom_point()  +
  facet_grid(site ~ strata)  +
  scale_shape_manual(values = c("CAN" = 1, "UNDER" = 21)) +
  scale_fill_manual(values = c("PNM" = "#8B7355", "PNSL" = "#1874CD")) +
  scale_colour_manual(values = c("PNM" = "#8B7355", "PNSL" = "#1874CD")) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))


ggplot(GL, aes(x = cell_vol, y = Aarea, fill = site, col = site, shape = strata)) +
  geom_point()  +
  facet_grid( ~ strata)  +
  scale_shape_manual(values = c("CAN" = 1, "UNDER" = 21)) +
  scale_fill_manual(values = c("PNM" = "#8B7355", "PNSL" = "#1874CD")) +
  scale_colour_manual(values = c("PNM" = "#8B7355", "PNSL" = "#1874CD")) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))


ggplot(GL, aes(x = cell_area, y = Aarea, fill = site, col = site, shape = strata)) +
  geom_point()  +
  facet_grid(site ~ strata)  +
  scale_shape_manual(values = c("CAN" = 1, "UNDER" = 21)) +
  scale_fill_manual(values = c("PNM" = "#8B7355", "PNSL" = "#1874CD")) +
  scale_colour_manual(values = c("PNM" = "#8B7355", "PNSL" = "#1874CD")) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))


 x <- rlnorm(100)

 y <- x^0.2

 plot(y ~ x)


LMAp <- x^0.2


```
