
---
title: "Null model check for Panama"
output:
  bookdown::pdf_book:
    fig_caption: true
    keep_tex: yes
    toc: no
    number_sections: no
    pandoc_args:
      - "--filter"
      - "pandoc-crossref"
fontsize: 12pt
sansfont: Times New Roman
geometry: margin=1in
link-citations: yes
---

## Optimal LL
 According to optimal LL theory, the optimal leaf lifespan (LL~opt~) maximizes a leaf’s lifetime carbon gain per-unit time, which is determined by potential LL, net photosynthetic rate, and construction cost per unit leaf area:

\begin{align}
  &{\rm LL_{opt}} = \sqrt{2bC/(a-m)}(\#eq:kikuzawa)
\end{align}

where *a* is the realized (i.e., light-dependent) gross photosynthetic rate per unit leaf area, *m* is the realized daily respiration rate per unit leaf area, *b* is the rate of decline in photosynthetic capacity with leaf age (which determines potential LL in the Kikuzawa model), and *C* is the construction cost per unit leaf area. To implement this Optimal LL Model, we assumed that (i) potential LL is proportional to LMAs; and (ii) leaf construction cost per unit area (g glucose per unit leaf area) is proportional to LMA. Assumption (ii) is justified because leaf construction cost per unit mass (g glucose per unit leaf mass) is strongly conserved. Thus, Eq. \@ref(eq:kikuzawa) can be written in terms of the traits considered in our analysis as:

### LMAsLMA (simple orginal version)

\begin{align}
  {\rm E[LL}_{i}] &=\beta_{2} \sqrt{{\rm LMAs}_{i}{\rm LMA}_{i}
  / (\theta_{\rm L} A_{{\rm area} \; i} - R_{{\rm area} \; i})}(\#eq:simple)
\end{align}

where $\beta_{2}$ is a constant, and $\theta_{\rm L} (0 < \theta_{\rm L} < 1)$ is a scaling parameter that accounts for the effects of light availability on the realized photosynthetic rate. The expected value of the logarithm of LL for the Optimal LL Model is:

- Potential LL is proportional to LMAs.

### $LMAs^{\beta_2}LMA$ (power-law)

\begin{align}
  {\rm E[LL}_{i}] &=\beta_{2} \sqrt{{\rm LMAs^{\beta_3}}_{i}{\rm LMA}_{i}
  / (\theta_{\rm L} A_{{\rm area} \; i} - R_{{\rm area} \; i})} \notag \\
  (\#eq:power)
\end{align}


- Potential LL and LMAs have a power-law relationship

### $LMAs^{\beta_2}$ (power-law, simple)

\begin{align}
  {\rm E[LL}_{i}] &=\beta_{2} \sqrt{{\rm LMAs^{\beta_3}}_{i}
  / (\theta_{\rm L} A_{{\rm area} \; i} - R_{{\rm area} \; i})} \notag \\
  (\#eq:power-simple)
\end{align}

- Model looks simple but no clear relationship between Kikuzawa model

\newpage
## LMAsLMA
```{r, fig.width = 5, fig.height=5, echo = FALSE}
setwd("~/Dropbox/MS/LES_MS/LMApsModel/data/")
library(loo)
# setwd("~/Dropbox/MS/LES_MS/LMApsModel/md/")
# render("PA_check.rmd", "bookdown::pdf_book")
# use null result 1
null_n <- 3

# PA data -------------------------------------------------------------------
load("PA_model3r_2017-07-07_.RData")
waic(extract_log_lik(res,"log_lik"))$waic
loo(extract_log_lik(res,"log_lik"))
s3r <- fit.summary
PA <- data

PA <- data %>%
  mutate(site2 = ifelse(site == "PNM", "DRY", "WET")) %>%
  mutate(sp_site_strata = paste(sp, site2, strata, sep = "_")) %>%
  mutate(site_strata = paste(site2, strata, sep = "_"))

P_vec <- paste("p[", 1:nrow(PA), "]", sep = "")
Mu_vec <- paste("mu[", 1:nrow(PA), ",2]", sep = "")
temp_LMAp <- fit.summary[P_vec, "mean"] * PA$LMA
temp_LMAs <- PA$LMA - temp_LMAp
temp_LL <- exp(fit.summary[Mu_vec, "mean"])

PA <- PA %>%
  mutate(LMAp = temp_LMAp) %>%
  mutate(LMAs = temp_LMAs) %>%
  mutate(preLL = temp_LL)

# rand LMA
load("PA_model3r2_LMA.ra.data_2017-06-30_.RData")
# waic(extract_log_lik(res,"log_lik"))$waic

s3rLMA <- fit.summary.list[[null_n]]
temp <- LMA.ra.data[,null_n]
temp_LL <- exp(s3rLMA[Mu_vec, "mean"])

PA <- PA %>%
  mutate(LMA_ra = temp) %>%
  mutate(LMAp_ra =  s3rLMA[P_vec, "mean"] * LMA_ra) %>%
  mutate(LMAs_ra = LMA_ra - LMAp_ra) %>%
  mutate(preLL_ra = exp(s3rLMA[Mu_vec, "mean"])
         )

write.csv(PA, "~/Dropbox/MS/LES_MS/LMApsModel/data/PA20170911.csv")

Lim.func <- function(trait){
  mid_len <- (log10(max(trait, na.rm = T)) - log10(min(trait, na.rm = T))) /2
  max_lim <- log10(max(trait, na.rm = T)) + mid_len * 0.3
  min_lim <- log10(min(trait, na.rm = T)) - mid_len * 0.3
  return(10^(c(min_lim, max_lim)))
}

PA.each <- function(x, y, data, log = log, y_text = NULL, x_text = NULL, subscript = NULL){
  sun.wet <- data %>% filter(strata == "CAN" & site == "PNSL")
  sun.dry <- data %>% filter(strata == "CAN" & site == "PNM")
  shade.wet <- data %>% filter(strata == "UNDER" & site == "PNSL")
  shade.dry <- data %>% filter(strata == "UNDER" & site == "PNM")

  rand.x <- paste(x, "rc", sep = "_")

  plot(data[, x], data[, y],
            log = log, axes = F, type = "n",
            xlim = Lim.func(data[, x]), ylim = Lim.func(data[, y]))

  mtext(subscript, side = 3, line = -1, adj = 0, cex = 0.65)
  box()


  points(sun.wet[, x], sun.wet[, y], col = "dodgerblue3",
    pch = 21, bg = "white", lwd = 1.2)

  points(sun.dry[, x], sun.dry[, y], col = "burlywood4",
    pch = 21, bg = "white", lwd = 1.2)

  points(shade.wet[, x], shade.wet[, y], bg = "dodgerblue3", col = "black",
    pch = 21, lwd = 0.8)

  points(shade.dry[, x], shade.dry[, y], bg = "burlywood4", col = "black",
    pch = 21, lwd = 0.8)


  if(is.null(y_text) != TRUE){
    axis(2, tick = FALSE,line = -0.8, cex.axis = 0.9)
    axis(2, tcl = 0.2, labels = FALSE)
    mtext(y_text, side = 2, line = 1.2, cex = 0.8)
  }

  if(is.null(x_text) != TRUE){
    axis(1, tick = FALSE, line = -0.8, cex.axis = 0.9)
    axis(1, tcl = 0.2, labels = FALSE, cex = 0.8)
    mtext(x_text, side = 1, line = 1.5, cex = 0.8)
  }
}

PA.each2 <- function(x, y, data, log = log, y_text = NULL, x_text = NULL, subscript = NULL){
  sun.wet <- data %>% filter(strata_all == "CAN" & site_all == "PNSL")
  sun.dry <- data %>% filter(strata_all == "CAN" & site_all == "PNM")
  shade.wet <- data %>% filter(strata_all == "UNDER" & site_all == "PNSL")
  shade.dry <- data %>% filter(strata_all == "UNDER" & site_all == "PNM")

  rand.x <- paste(x, "rc", sep = "_")

  plot(data[, x], data[, y],
            log = log, axes = F, type = "n",
            xlim = Lim.func(data[, x]), ylim = Lim.func(data[, y]))

  mtext(subscript, side = 3, line = -1, adj = 0, cex = 0.65)
  box()


  points(sun.wet[, x], sun.wet[, y], col = "dodgerblue3",
    pch = 21, bg = "white", lwd = 1.2)

  points(sun.dry[, x], sun.dry[, y], col = "burlywood4",
    pch = 21, bg = "white", lwd = 1.2)

  points(shade.wet[, x], shade.wet[, y], bg = "dodgerblue3", col = "black",
    pch = 21, lwd = 0.8)

  points(shade.dry[, x], shade.dry[, y], bg = "burlywood4", col = "black",
    pch = 21, lwd = 0.8)


  if(is.null(y_text) != TRUE){
    axis(2, tick = FALSE,line = -0.8, cex.axis = 0.9)
    axis(2, tcl = 0.2, labels = FALSE)
    mtext(y_text, side = 2, line = 1.2, cex = 0.8)
  }

  if(is.null(x_text) != TRUE){
    axis(1, tick = FALSE, line = -0.8, cex.axis = 0.9)
    axis(1, tcl = 0.2, labels = FALSE, cex = 0.8)
    mtext(x_text, side = 1, line = 1.5, cex = 0.8)
  }
}

#head(fit.summary)
```

```{r, fig.width = 5, fig.height=5, echo = FALSE}
moge_plt <- function() {
  par(mfrow = c(3, 3))
  par(mar = c(0, 0, 0, 0))
  par(oma = c(3, 3, 2, 2))

  PA.each(y = "Aarea", x = "LMA", data = PA, log = "xy",
    subscript = " (a)",
    y_text = expression(paste(italic(A)[area]," (", mu, "mol ", m^-2," ", s^-1, ")")))


  par(xpd = NA)
  legend(10, 85, c("Sun-Wet", "Sun-Dry", "Shade-Wet", "Shade-Dry", "Rand"),
           pch = c(21, 21, 21, 21, 21),
           col = c("dodgerblue3", "burlywood4", "black", "black", "gray"),
           pt.bg = c("white", "white", "dodgerblue3", "burlywood4", "gray"),
           bty = "n",
           # lwd = 1,
           cex = 0.9,
           pt.cex = 1,
           pt.lwd = 0.8,
           ncol = 5)
  par(xpd = TRUE)

  PA.each(y = "Aarea", x = "LMAp",
    data = PA, log = "xy", subscript = " (b)")

  PA.each(y = "Aarea", x = "LMAs",
    data = PA, log = "xy", subscript = " (c)")

  PA.each(y = "Rarea", x = "LMA", data = PA, log = "xy",
    subscript = " (d)",
    y_text = expression(paste(italic(R)[area]," (", mu, "mol ", m^-2," ", s^-1, ")")))

  PA.each(y = "Rarea", x = "LMAp",
    data = PA, log = "xy", subscript = " (e)")

  PA.each(y = "Rarea", x = "LMAs",
    data = PA, log = "xy", subscript = " (f)",)

  PA.each(y = "LL", x = "LMA", data = PA, log = "xy",
    subscript = " (g)",
    y_text = expression(paste("LL (months)")),
    x_text = expression(paste("LMA (g ",m^-2,")")))

  PA.each(y = "LL", x = "LMAp",
    data = PA, log = "xy",
    subscript = " (h)",
    x_text = expression(paste("LMAp (g ",m^-2,")")))

  PA.each(y = "LL", x = "LMAs",
    data = PA, log = "xy",
    subscript = " (i)",
    x_text = expression(paste("LMAs (g ",m^-2,")")))

par(mfrow = c(1,1))
}

moge_plt_ra <- function() {
  par(mfrow = c(3, 3))
  par(mar = c(0, 0, 0, 0))
  par(oma = c(3, 3, 2, 2))

  PA.each(y = "Aarea", x = "LMA_ra", data = PA, log = "xy",
    subscript = " (a)",
    y_text = expression(paste(italic(A)[area]," (", mu, "mol ", m^-2," ", s^-1, ")")))


  par(xpd = NA)
  legend(10, 85, c("Sun-Wet", "Sun-Dry", "Shade-Wet", "Shade-Dry", "Rand"),
           pch = c(21, 21, 21, 21, 21),
           col = c("dodgerblue3", "burlywood4", "black", "black", "gray"),
           pt.bg = c("white", "white", "dodgerblue3", "burlywood4", "gray"),
           bty = "n",
           # lwd = 1,
           cex = 0.9,
           pt.cex = 1,
           pt.lwd = 0.8,
           ncol = 5)
  par(xpd = TRUE)

  PA.each(y = "Aarea", x = "LMAp_ra",
    data = PA, log = "xy", subscript = " (b)")

  PA.each(y = "Aarea", x = "LMAs_ra",
    data = PA, log = "xy", subscript = " (c)")

  PA.each(y = "Rarea", x = "LMA_ra", data = PA, log = "xy",
    subscript = " (d)",
    y_text = expression(paste(italic(R)[area]," (", mu, "mol ", m^-2," ", s^-1, ")")))

  PA.each(y = "Rarea", x = "LMAp_ra",
    data = PA, log = "xy", subscript = " (e)")

  PA.each(y = "Rarea", x = "LMAs_ra",
    data = PA, log = "xy", subscript = " (f)",)

  PA.each(y = "LL", x = "LMA_ra", data = PA, log = "xy",
    subscript = " (g)",
    y_text = expression(paste("LL (months)")),
    x_text = expression(paste("LMA (g ",m^-2,")")))

  PA.each(y = "LL", x = "LMAp_ra",
    data = PA, log = "xy",
    subscript = " (h)",
    x_text = expression(paste("LMAp (g ",m^-2,")")))

  PA.each(y = "LL", x = "LMAs_ra",
    data = PA, log = "xy",
    subscript = " (i)",
    x_text = expression(paste("LMAs (g ",m^-2,")")))

  par(mfrow = c(1,1))
}


moge_plt_all <- function() {
  par(mfrow = c(3, 3))
  par(mar = c(0, 0, 0, 0))
  par(oma = c(3, 3, 2, 2))

  PA.each2(y = "Aarea_all", x = "LMA_all", data = PA, log = "xy",
    subscript = " (a)",
    y_text = expression(paste(italic(A)[area]," (", mu, "mol ", m^-2," ", s^-1, ")")))


  par(xpd = NA)
  legend(10, 85, c("Sun-Wet", "Sun-Dry", "Shade-Wet", "Shade-Dry", "Rand"),
           pch = c(21, 21, 21, 21, 21),
           col = c("dodgerblue3", "burlywood4", "black", "black", "gray"),
           pt.bg = c("white", "white", "dodgerblue3", "burlywood4", "gray"),
           bty = "n",
           # lwd = 1,
           cex = 0.9,
           pt.cex = 1,
           pt.lwd = 0.8,
           ncol = 5)
  par(xpd = TRUE)

  PA.each2(y = "Aarea_all", x = "LMAp_all",
    data = PA, log = "xy", subscript = " (b)")

  PA.each2(y = "Aarea_all", x = "LMAs_all",
    data = PA, log = "xy", subscript = " (c)")

  PA.each2(y = "Rarea_all", x = "LMA_all", data = PA, log = "xy",
    subscript = " (d)",
    y_text = expression(paste(italic(R)[area]," (", mu, "mol ", m^-2," ", s^-1, ")")))

  PA.each2(y = "Rarea_all", x = "LMAp_all",
    data = PA, log = "xy", subscript = " (e)")

  PA.each2(y = "Rarea_all", x = "LMAs_all",
    data = PA, log = "xy", subscript = " (f)",)

  PA.each2(y = "LL_all", x = "LMA_all", data = PA, log = "xy",
    subscript = " (g)",
    y_text = expression(paste("LL (months)")),
    x_text = expression(paste("LMA (g ",m^-2,")")))

  PA.each2(y = "LL_all", x = "LMAp_all",
    data = PA, log = "xy",
    subscript = " (h)",
    x_text = expression(paste("LMAp (g ",m^-2,")")))

  PA.each2(y = "LL_all", x = "LMAs_all",
    data = PA, log = "xy",
    subscript = " (i)",
    x_text = expression(paste("LMAs (g ",m^-2,")")))

  par(mfrow = c(1,1))
}

# NPC
moge_plt2 <- function() {
  par(mfrow = c(3, 3))
  par(mar = c(0, 0, 0, 0))
  par(oma = c(3, 3, 2, 2))
  PA.each(y = "Narea", x = "LMA", data = PA2, log = "xy",
  subscript = " (a)",
  y_text = expression(paste(italic(N)[area]," (g ",m^-2,")")))
  par(xpd = NA)
  legend(10, 0.7, c("Sun-Wet", "Sun-Dry", "Shade-Wet", "Shade-Dry", "Rand"),
           pch = c(21, 21, 21, 21, 21),
           col = c("dodgerblue3", "burlywood4", "black", "black", "gray"),
           pt.bg = c("white", "white", "dodgerblue3", "burlywood4", "gray"),
           bty = "n",
           # lwd = 1,
           cex = 0.9,
           pt.cex = 1,
           pt.lwd = 0.8,
           ncol = 5)
  par(xpd = TRUE)

  PA.each(y = "Narea", x = "LMAp",
    data = PA2, log = "xy", subscript = " (b)")

  PA.each(y = "Narea", x = "LMAs",
    data = PA2, log = "xy", subscript = " (c)")

  PA.each(y = "Parea", x = "LMA", data = PA2, log = "xy",
    subscript = " (d)",
    y_text = expression(paste(italic(P)[area]," (g ",m^-2,")")))

  PA.each(y = "Parea", x = "LMAp",
    data = PA2, log = "xy", subscript = " (e)")

  PA.each(y = "Parea", x = "LMAs",
    data = PA2, log = "xy", subscript = " (f)",)

  PA.each(y = "cell_area", x = "LMA", data = PA2, log = "xy",
    subscript = " (g)",
    y_text = expression(paste(italic(CL)[area]," (g ",m^-2,")")),
    x_text = expression(paste("LMA (g ",m^-2,")")))

  PA.each(y = "cell_area", x = "LMAp",
    data = PA2, log = "xy",
    subscript = " (h)",
    x_text = expression(paste("LMAp (g ",m^-2,")")))

  PA.each(y = "cell_area", x = "LMAs",
    data = PA2, log = "xy",
    subscript = " (i)",
    x_text = expression(paste("LMAs (g ",m^-2,")")))
    par(mfrow = c(3, 3))
    par(mar = c(0, 0, 0, 0))
    par(oma = c(3, 3, 2, 2))

}

```

```{r, fig.width = 4.33, fig.height=4.33, echo = FALSE}
setwd("~/Dropbox/MS/LES_MS/LMApsModel/data/")
load("PA_model3r_LMA.all.data2_2017-07-12_.RData")
s3r_all <- fit.summary.list[[null_n]]

temp <- LMA.all.data[[null_n]]

#plot(Aarea ~ Rarea, log = "xy", temp)

PA <- PA %>%
  mutate(LMA_all = temp[ ,"LMA"]) %>%
  mutate(LMAp_all =  s3r_all[P_vec, "mean"] * LMA_all) %>%
  mutate(LMAs_all = LMA_all - LMAp_all) %>%
  mutate(LL_all = temp[ ,"LL"]) %>%
  mutate(Aarea_all = temp[, "Aarea"]) %>%
  mutate(Rarea_all = temp[, "Rarea"]) %>%
  mutate(site_all = temp[, "site"]) %>%
  mutate(strata_all = temp[, "strata"])

cat("\\newpage")
print("LMAsLMA; obs")
moge_plt()

cat("\\newpage")
print("LMAsLMA; LMA only")
moge_plt_ra()

cat("\\newpage")
print("LMAsLMA; LMA all")
moge_plt_all()

print("LMAsLMA; obs")
head(s3r, 10)

print("LMAsLMA; LMA only")
head(s3rLMA, 10)

print("LMAsLMA; LMA all")
head(s3r_all, 10)

# fiber data -------------------------------------------------------------------
fiber <- read.csv("~/Dropbox/LES/fiber_analysis.csv")
trait <- read.csv("~/Dropbox/LES/LFTRAITS.csv")

trait2 <- trait %>%
  mutate(strata = ifelse(STRATA. == "CANOPY", "CAN", "UNDER")) %>%
  mutate(site = ifelse(trait$SITE. == "PNM", "PNM", "PNLS")) %>%
  mutate(site2 = ifelse(trait$SITE. == "PNM", "DRY", "WET")) %>%
  mutate(sp_site_strata = paste(SP4., site2, strata, sep = "_")) %>%
  mutate(LMA_LEAF = 1/SLA_LEAF * 10000) %>%
  mutate(LMA = 1/SLA_DISC * 10000) %>%
  mutate(Narea = LMA_LEAF * N_PCT / 1000) %>%
  mutate(Parea = LMA_LEAF * P_PCT / 1000) %>%
  mutate(LD = LMA / LFTHICK / 1000) %>%
  mutate(LAMTUF = LAMTUF / 1000, MDRBTUF = MDRBTUF / 1000,
      VEINTUF = VEINTUF / 1000)

fiber2 <- fiber %>%
  mutate(site2 = ifelse(site == "PNM", "DRY", "WET")) %>%
  mutate(sp_site_strata = paste(sp, site2, position, sep = "_"))

fdata <- full_join(trait2, fiber2, by = "sp_site_strata") %>%
  select(sp_site_strata, X.ADF, X.NPE, X.Lignin,
     LFTHICK, LD, Narea, Parea, LAMTUF, MDRBTUF, VEINTUF) %>%
  rename(ADF = X.ADF, NPE = X.NPE, Lig = X.Lignin, LT = LFTHICK)

PA2 <- left_join(PA, fdata, by = "sp_site_strata") %>%
  mutate(cell_mass = ADF - Lig) %>%
  mutate(cell_area = cell_mass / 100 * LMA) %>%
  mutate(cell_vol = cell_area / LT / 100 / 100 / 0.1) # g/cm3

moge_plt2()

````









\newpage
## $$LMAs^{\beta_2}LMA$$

```{r, fig.width = 4.33, fig.height=4.33, echo = FALSE}
setwd("~/Dropbox/MS/LES_MS/LMApsModel/data/")
load("PA_model3r2_2017-07-07_.RData")
waic(extract_log_lik(res,"log_lik"))$waic
loo(extract_log_lik(res,"log_lik"))

s3r <- fit.summary
PA <- data

PA <- data %>%
  mutate(site2 = ifelse(site == "PNM", "DRY", "WET")) %>%
  mutate(sp_site_strata = paste(sp, site2, strata, sep = "_")) %>%
  mutate(site_strata = paste(site2, strata, sep = "_"))

P_vec <- paste("p[", 1:nrow(PA), "]", sep = "")
Mu_vec <- paste("mu[", 1:nrow(PA), ",2]", sep = "")
temp_LMAp <- fit.summary[P_vec, "mean"] * PA$LMA
temp_LMAs <- PA$LMA - temp_LMAp
temp_LL <- exp(fit.summary[Mu_vec, "mean"])

PA <- PA %>%
  mutate(LMAp = temp_LMAp) %>%
  mutate(LMAs = temp_LMAs) %>%
  mutate(preLL = temp_LL)

moge_plt()


cat("\\newpage")
load("PA_model3r2_LMA.all.data2_2017-07-12_.RData")
s3r_all <- fit.summary.list[[null_n]]

temp <- LMA.all.data[[null_n]]

#plot(Aarea ~ Rarea, log = "xy", temp)

PA <- PA %>%
  mutate(LMA_all = temp[ ,"LMA"]) %>%
  mutate(LMAp_all =  s3r_all[P_vec, "mean"] * LMA_all) %>%
  mutate(LMAs_all = LMA_all - LMAp_all) %>%
  mutate(LL_all = temp[ ,"LL"]) %>%
  mutate(Aarea_all = temp[, "Aarea"]) %>%
  mutate(Rarea_all = temp[, "Rarea"]) %>%
  mutate(site_all = temp[, "site"]) %>%
  mutate(strata_all = temp[, "strata"])

print("LMAs^beta2 LMA; LMA all")
moge_plt_all()


print("LMAs^betaLMA; obs")
head(s3r, 12)


print("LMAs^betaLMA; LMA all")
head(s3r_all, 12)


# fiber data -------------------------------------------------------------------
fiber <- read.csv("~/Dropbox/LES/fiber_analysis.csv")
trait <- read.csv("~/Dropbox/LES/LFTRAITS.csv")

trait2 <- trait %>%
  mutate(strata = ifelse(STRATA. == "CANOPY", "CAN", "UNDER")) %>%
  mutate(site = ifelse(trait$SITE. == "PNM", "PNM", "PNLS")) %>%
  mutate(site2 = ifelse(trait$SITE. == "PNM", "DRY", "WET")) %>%
  mutate(sp_site_strata = paste(SP4., site2, strata, sep = "_")) %>%
  mutate(LMA_LEAF = 1/SLA_LEAF * 10000) %>%
  mutate(LMA = 1/SLA_DISC * 10000) %>%
  mutate(Narea = LMA_LEAF * N_PCT / 1000) %>%
  mutate(Parea = LMA_LEAF * P_PCT / 1000) %>%
  mutate(LD = LMA / LFTHICK / 1000) %>%
  mutate(LAMTUF = LAMTUF / 1000, MDRBTUF = MDRBTUF / 1000,
      VEINTUF = VEINTUF / 1000)

fiber2 <- fiber %>%
  mutate(site2 = ifelse(site == "PNM", "DRY", "WET")) %>%
  mutate(sp_site_strata = paste(sp, site2, position, sep = "_"))

fdata <- full_join(trait2, fiber2, by = "sp_site_strata") %>%
  select(sp_site_strata, X.ADF, X.NPE, X.Lignin,
     LFTHICK, LD, Narea, Parea, LAMTUF, MDRBTUF, VEINTUF) %>%
  rename(ADF = X.ADF, NPE = X.NPE, Lig = X.Lignin, LT = LFTHICK)

PA2 <- left_join(PA, fdata, by = "sp_site_strata") %>%
  mutate(cell_mass = ADF - Lig) %>%
  mutate(cell_area = cell_mass / 100 * LMA) %>%
  mutate(cell_vol = cell_area / LT / 100 / 100 / 0.1) # g/cm3

moge_plt2()



```  









\newpage
## potential LL

```{r, fig.width = 4.33, fig.height=4.33, echo = FALSE}
setwd("~/Dropbox/MS/LES_MS/LMApsModel/data/")
load("PA_model1r_2017-07-07_.RData")
waic(extract_log_lik(res,"log_lik"))$waic
loo(extract_log_lik(res,"log_lik"))

s3r <- fit.summary
PA <- data

PA <- data %>%
  mutate(site2 = ifelse(site == "PNM", "DRY", "WET")) %>%
  mutate(sp_site_strata = paste(sp, site2, strata, sep = "_")) %>%
  mutate(site_strata = paste(site2, strata, sep = "_"))

P_vec <- paste("p[", 1:nrow(PA), "]", sep = "")
Mu_vec <- paste("mu[", 1:nrow(PA), ",2]", sep = "")
temp_LMAp <- fit.summary[P_vec, "mean"] * PA$LMA
temp_LMAs <- PA$LMA - temp_LMAp
temp_LL <- exp(fit.summary[Mu_vec, "mean"])

PA <- PA %>%
  mutate(LMAp = temp_LMAp) %>%
  mutate(LMAs = temp_LMAs) %>%
  mutate(preLL = temp_LL)

moge_plt()


cat("\\newpage")
load("PA_model1r_LMA.all.data2_2017-07-12_.RData")
s3r_all <- fit.summary.list[[null_n]]

temp <- LMA.all.data[[null_n]]

#plot(Aarea ~ Rarea, log = "xy", temp)

PA <- PA %>%
  mutate(LMA_all = temp[ ,"LMA"]) %>%
  mutate(LMAp_all =  s3r_all[P_vec, "mean"] * LMA_all) %>%
  mutate(LMAs_all = LMA_all - LMAp_all) %>%
  mutate(LL_all = temp[ ,"LL"]) %>%
  mutate(Aarea_all = temp[, "Aarea"]) %>%
  mutate(Rarea_all = temp[, "Rarea"]) %>%
  mutate(site_all = temp[, "site"]) %>%
  mutate(strata_all = temp[, "strata"])

print("potential; LMA all")
moge_plt_all()


print("potential; obs")
head(s3r, 12)


print("poential; LMA all")
head(s3r_all, 12)


# fiber data -------------------------------------------------------------------
fiber <- read.csv("~/Dropbox/LES/fiber_analysis.csv")
trait <- read.csv("~/Dropbox/LES/LFTRAITS.csv")

trait2 <- trait %>%
  mutate(strata = ifelse(STRATA. == "CANOPY", "CAN", "UNDER")) %>%
  mutate(site = ifelse(trait$SITE. == "PNM", "PNM", "PNLS")) %>%
  mutate(site2 = ifelse(trait$SITE. == "PNM", "DRY", "WET")) %>%
  mutate(sp_site_strata = paste(SP4., site2, strata, sep = "_")) %>%
  mutate(LMA_LEAF = 1/SLA_LEAF * 10000) %>%
  mutate(LMA = 1/SLA_DISC * 10000) %>%
  mutate(Narea = LMA_LEAF * N_PCT / 1000) %>%
  mutate(Parea = LMA_LEAF * P_PCT / 1000) %>%
  mutate(LD = LMA / LFTHICK / 1000) %>%
  mutate(LAMTUF = LAMTUF / 1000, MDRBTUF = MDRBTUF / 1000,
      VEINTUF = VEINTUF / 1000)

fiber2 <- fiber %>%
  mutate(site2 = ifelse(site == "PNM", "DRY", "WET")) %>%
  mutate(sp_site_strata = paste(sp, site2, position, sep = "_"))

fdata <- full_join(trait2, fiber2, by = "sp_site_strata") %>%
  select(sp_site_strata, X.ADF, X.NPE, X.Lignin,
     LFTHICK, LD, Narea, Parea, LAMTUF, MDRBTUF, VEINTUF) %>%
  rename(ADF = X.ADF, NPE = X.NPE, Lig = X.Lignin, LT = LFTHICK)

PA2 <- left_join(PA, fdata, by = "sp_site_strata") %>%
  mutate(cell_mass = ADF - Lig) %>%
  mutate(cell_area = cell_mass / 100 * LMA) %>%
  mutate(cell_vol = cell_area / LT / 100 / 100 / 0.1) # g/cm3

moge_plt2()


```  




## not root

```{r, fig.width = 4.33, fig.height=4.33, echo = FALSE}
setwd("~/Dropbox/MS/LES_MS/LMApsModel/data/")
load("PA_model3r5_2017-07-11_.RData")
waic(extract_log_lik(res,"log_lik"))$waic
loo(extract_log_lik(res,"log_lik"))

s3r <- fit.summary
PA <- data

PA <- data %>%
  mutate(site2 = ifelse(site == "PNM", "DRY", "WET")) %>%
  mutate(sp_site_strata = paste(sp, site2, strata, sep = "_")) %>%
  mutate(site_strata = paste(site2, strata, sep = "_"))

P_vec <- paste("p[", 1:nrow(PA), "]", sep = "")
Mu_vec <- paste("mu[", 1:nrow(PA), ",2]", sep = "")
temp_LMAp <- fit.summary[P_vec, "mean"] * PA$LMA
temp_LMAs <- PA$LMA - temp_LMAp
temp_LL <- exp(fit.summary[Mu_vec, "mean"])

PA <- PA %>%
  mutate(LMAp = temp_LMAp) %>%
  mutate(LMAs = temp_LMAs) %>%
  mutate(preLL = temp_LL)

moge_plt()


cat("\\newpage")
load("PA_model3r5_LMA.all.data2_2017-07-12_.RData")
s3r_all <- fit.summary.list[[null_n]]

temp <- LMA.all.data[[null_n]]

#plot(Aarea ~ Rarea, log = "xy", temp)

PA <- PA %>%
  mutate(LMA_all = temp[ ,"LMA"]) %>%
  mutate(LMAp_all =  s3r_all[P_vec, "mean"] * LMA_all) %>%
  mutate(LMAs_all = LMA_all - LMAp_all) %>%
  mutate(LL_all = temp[ ,"LL"]) %>%
  mutate(Aarea_all = temp[, "Aarea"]) %>%
  mutate(Rarea_all = temp[, "Rarea"]) %>%
  mutate(site_all = temp[, "site"]) %>%
  mutate(strata_all = temp[, "strata"])

print("not root; LMA all")
moge_plt_all()


print("not root; obs")
head(s3r, 12)


print("not root; LMA all")
head(s3r_all, 12)


# fiber data -------------------------------------------------------------------
fiber <- read.csv("~/Dropbox/LES/fiber_analysis.csv")
trait <- read.csv("~/Dropbox/LES/LFTRAITS.csv")

trait2 <- trait %>%
  mutate(strata = ifelse(STRATA. == "CANOPY", "CAN", "UNDER")) %>%
  mutate(site = ifelse(trait$SITE. == "PNM", "PNM", "PNLS")) %>%
  mutate(site2 = ifelse(trait$SITE. == "PNM", "DRY", "WET")) %>%
  mutate(sp_site_strata = paste(SP4., site2, strata, sep = "_")) %>%
  mutate(LMA_LEAF = 1/SLA_LEAF * 10000) %>%
  mutate(LMA = 1/SLA_DISC * 10000) %>%
  mutate(Narea = LMA_LEAF * N_PCT / 1000) %>%
  mutate(Parea = LMA_LEAF * P_PCT / 1000) %>%
  mutate(LD = LMA / LFTHICK / 1000) %>%
  mutate(LAMTUF = LAMTUF / 1000, MDRBTUF = MDRBTUF / 1000,
      VEINTUF = VEINTUF / 1000)

fiber2 <- fiber %>%
  mutate(site2 = ifelse(site == "PNM", "DRY", "WET")) %>%
  mutate(sp_site_strata = paste(sp, site2, position, sep = "_"))

fdata <- full_join(trait2, fiber2, by = "sp_site_strata") %>%
  select(sp_site_strata, X.ADF, X.NPE, X.Lignin,
     LFTHICK, LD, Narea, Parea, LAMTUF, MDRBTUF, VEINTUF) %>%
  rename(ADF = X.ADF, NPE = X.NPE, Lig = X.Lignin, LT = LFTHICK)

PA2 <- left_join(PA, fdata, by = "sp_site_strata") %>%
  mutate(cell_mass = ADF - Lig) %>%
  mutate(cell_area = cell_mass / 100 * LMA) %>%
  mutate(cell_vol = cell_area / LT / 100 / 100 / 0.1) # g/cm3

moge_plt2()

cor.test(log(PA2$Rarea_all), log(PA2$LMAp_all))

cor.test(log(PA2$LL), log(PA2$preLL))
cor.test(log(PA2$LL), log(PA2$LMAs))

cor.test(log(PA2$Narea), log(PA2$LMA))
cor.test(log(PA2$Narea), log(PA2$LMAp))

lm(log(Aarea + Rarea) ~log(LMAp), data = PA) %>% summary
lm(log(Rarea) ~log(LMAp) + log(LMAs), data = PA) %>% summary

cor.test(log(PA2$Parea), log(PA2$LMA))
cor.test(log(PA2$Parea), log(PA2$LMAp))

cor.test(log(PA2$cell_area), log(PA2$LMA))
cor.test(log(PA2$cell_area), log(PA2$LMAp))
cor.test(log(PA2$cell_area), log(PA2$LMAs))

```  









