---
title: Panama results using MVN (Kikuzawa model)
author: Masatoshi Katabuchi
date: "`r format(Sys.time(), '%B %d, %Y')`"
---

More species than LT dataset

- alpha_0 LMAp^alpha_p 
- beta_0 {LMALMAs/(qA-R)}^0.5

Those update include:

- y = ax to y = ax^b^
    - Originally we were using y = ax, which is simple but not biologically correct and could produce some degrees of correlations even in randomized dataset.
    - The new model includes scaling parameters and now randomized dataset never produced correlations. We do not have to mention randomized dataset.
- LMAs / LT (= leaf structural density)
    - Leaf density is more relevant to LL compared to LMA, so using (LMAs/LT) is more straightforward. Leaves are now three-dimensional, which is also improvement of Osnas's PNAS.
- Effect of LMAs/LT to photosynthesis
    - The parameters $\alpha_s$ in Eq. 1 is the effect of leaf structural density to gross photosynthesis.
    - Based on effects of cell-wall thickness on mesophyll conductance, we expect $\alpha_s$ < 0; i.e., leaves with higher LMAs and/or higher LD have lower Aarea for a given LMAp.


```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=TRUE,
                      warning=FALSE,
                      message=FALSE)
```


```{r, echo = F, message = F}
library(loo)
library(rstan)
library(dplyr)
library(scales)
library(memisc)

theme_set(theme_light())

```


```{r, echo = F, message = F}

#load("./data/PA_latent_model_light_more_obs.rda")
#load("../data/PA_latent_model_light_more_obs.rda")
#
#load("./data/PA_latent_model_more_obs.rda")
#
#load("./data/PA_latent_model_site_more_obs.rda")
load("../data/PA_latent_model_light_site_more_obs.rda")


divergent <- get_sampler_params(res, inc_warmup=FALSE)[[1]][,'divergent__']
sum(divergent)
c_light <- c("#DCBCBC")
c_light_highlight <- c("#C79999")
c_mid <- c("#B97C7C")
c_mid_highlight <- c("#A25050")
c_dark <- c("#8F2727")
c_dark_highlight <- c("#7C0000")
params_cp80 <- as.data.frame(extract(res, permuted=FALSE))
params_cp80$divergent <- divergent
div_params_cp <- params_cp80[params_cp80$divergent == 1,]
nondiv_params_cp <- params_cp80[params_cp80$divergent == 0,]


par(mar = c(4, 4, 0.5, 0.5))
plot(nondiv_params_cp$`chain:1.z[4]`, nondiv_params_cp$`chain:1.z[7]`,
     col=c_dark, pch=16, cex=0.8)
points(div_params_cp$`chain:1.z[4]`, div_params_cp$`chain:1.z[7]`,
       col="orange", pch=16, cex=0.8)


my_logit <- function(x) {
  1/(1+exp(-x))
}

my_logit(-1)

#load("./data/PA_latent_model_light2_more_obs.rda")

pairs(res, pars = c("Z[1,1]", "Z[2,1]", "Z[3,1]", "lp__"))
pairs(res, pars = c("Z[1,2]", "Z[2,2]", "Z[3,2]", "lp__"))
pairs(res, pars = c("Z[1,3]", "Z[2,3]", "Z[3,3]", "lp__"))

traceplot(res, pars = "Z", ncol = 3)

traceplot(res, pars = c("p[1]",
                        "p[2]",
                        "lp__"
                        ))

traceplot(res, pars = "L_Omega")

summary_PA_LDL <- data.frame(summary(res)$summary)

summary_PA_LDL %>%
  round(3) %>%
  DT::datatable(.)

PA <- dat %>%
  as_data_frame %>%
  mutate(site2 = ifelse(site == "PNM", "DRY", "WET")) %>%
  mutate(sp_site_strata = paste(sp, site2, strata, sep = "_")) %>%
  mutate(site_strata = paste(site2, strata, sep = "_"))


P_vec <- paste("p[", 1:nrow(PA), "]", sep = "")
#mu2_vec <- paste("mu2[", 1:nrow(PA), "]", sep = "")
#Mu_vec <- paste("mu[", 1:nrow(PA), ",2]", sep = "")
temp_LMAp <- summary_PA_LDL[P_vec, "mean"] * PA$LMA
temp_LMAs <- PA$LMA - temp_LMAp
#temp_LL <- exp(summary_PA_LDL[mu2_vec, "mean"])


PA <- PA %>%
  mutate(LMAp = temp_LMAp) %>%
  mutate(LMAs = temp_LMAs) 


```


# Posterior distributions of $\alpha_s$ > 0


# Paramters 

```{r, echo = F, eval = T}

plot(res)

traceplot(res)

```


```{r, echo = F}

var_p <- function(data)cov(data$LMA, data$LMAp)
var_s <- function(data)cov(data$LMA, data$LMAs)
var_t <- function(data)var(data$LMA)

var_p <- function(data)var(data$LMAp)
var_s <- function(data)var(data$LMAs)
cov_ps <- function(data)cov(data$LMAp, data$LMAs)
var_t <- function(data)var(data$LMA)

var_dat <- PA %>%
  group_by(strata) %>%
  nest %>%
  mutate(var_LMAm = map_dbl(data, var_p)) %>%
  mutate(var_LMAs = map_dbl(data, var_s)) %>%
  mutate(cov_ms = 2 * map_dbl(data, cov_ps)) %>%
  mutate(var_total = map_dbl(data, var_t)) %>%
  dplyr::select(-data)

var_dat_s <- var_dat %>%
  mutate(LMAm_variance = var_LMAm / (var_LMAm + var_LMAs) * 100) %>%
  mutate(LMAs_variance = var_LMAs / (var_LMAm + var_LMAs) * 100) %>%
  gather(var, val, c(LMAm_variance, LMAs_variance))

var_dat_s2 <- var_dat %>%
  mutate(LMAm_variance = var_LMAm / (var_total) * 100) %>%
  mutate(LMAs_variance = var_LMAs / (var_total) * 100) %>%
  mutate(cov_LMAm_LMAs = cov_ms / (var_total) * 100) %>%
  gather(var, val, c(LMAm_variance, LMAs_variance, cov_LMAm_LMAs)) 


```



```{r, echo = F, eval = F}
# R values

#R values are based on posterior distributions of LMAs

par(mfrow = c(1,2))
apply(log_LMAs, 1, function(x)cor(x, log(PA$Aarea))) %>% hist(main = "LMAs vs Aarea, R")
apply(log_LMAs, 1, function(x)cor(x, log(PA$Aarea))^2) %>% hist(main = "LMAs vs Aarea, R2")

apply(log_LMAs, 1, function(x)cor(x, log(PA$Aarea + PA$Rarea))) %>% hist(main = "LMAs vs Aarea + Rarea, R")
apply(log_LMAs, 1, function(x)cor(x, log(PA$Aarea + PA$Rarea))^2) %>% hist(main = "LMAs vs Aarea + Rarea, R2")


apply(log_LDs, 1, function(x)cor(x, log(PA$Aarea))) %>% hist(main = "LMAs/LT vs Aarea, R")
apply(log_LDs, 1, function(x)cor(x, log(PA$Aarea))^2) %>% hist(main = "LMAs/LT vs Aarea, R2")

apply(log_LDs, 1, function(x)cor(x, log(PA$Aarea + PA$Rarea))) %>% hist(main = "LMAs/LT vs Aarea + Rarea, R")
apply(log_LDs, 1, function(x)cor(x, log(PA$Aarea + PA$Rarea))^2) %>% hist(main = "LMAs/LT vs Aarea + Rarea, R2")
```

```{r}

fit <-res

#fit %>% as.array %>% dim

params <- extract(fit)

m <- 4000

library(mvtnorm)

Z_i <- params$Z[,1:5,1:3]


plot(1:m, Z_i[,2,1], type = "l", col ="orange",ylim=c(0,0.8))
points(1:m, Z_i[,3,1], type = "l", col ="violet")

plot(1:m, Z_i[,2,2], type = "l", col ="orange",ylim=c(-1, 0.5))
points(1:m, Z_i[,3,2], type = "l", col ="violet")

plot(1:m, Z_i[,2,3], type = "l", col ="orange",ylim=c(-0.5,1.5))
points(1:m, Z_i[,3,3], type = "l", col ="violet")

plot(1:m, Z_i[,4,2], type = "l", col ="orange")
plot(1:m, Z_i[,5,2], type = "l", col ="orange")

lp_m <- params$lp__
p_m <- params$p
rho13 <- params$rho13

plot(1:m, rho13, type = "l", col ="orange")
plot(1:m, lp_m, type = "l", col ="orange")
plot(1:m, p_m[,1], type = "l", col ="orange")


Sigma_i <- params$Sigma[,1:3,1:3]
L_sigma_i <- params$L_sigma[,]

```

# para fix

```{r}
mcmc_res <- extract(fit) %>% as.data.frame
params1 <- as.data.frame(extract(fit, permuted=FALSE)[,1,])
params2 <- as.data.frame(extract(fit, permuted=FALSE)[,2,])
params3 <- as.data.frame(extract(fit, permuted=FALSE)[,3,])
params4 <- as.data.frame(extract(fit, permuted=FALSE)[,4,])


get_stancode(fit) %>% cat

re_am <- c(params1$`Z[2,1]`,
	 			 params2$`Z[3,1]`,
	 			 params3$`Z[3,1]`,
	 			 params4$`Z[3,1]`)

re_as <- c(params1$`Z[3,1]`,
	 			 params2$`Z[2,1]`,
	 			 params3$`Z[2,1]`,
	 			 params4$`Z[2,1]`)

re_bm <- c(params1$`Z[2,2]`,
	 			 params2$`Z[3,2]`,
         params3$`Z[3,2]`,
         params4$`Z[3,2]`)

re_bs <- c(params1$`Z[3,2]`,
           params2$`Z[2,2]`,
           params3$`Z[2,2]`,
           params4$`Z[2,2]`)

re_rs <- c(params1$`Z[3,3]`,
	 			 params2$`Z[2,3]`,
         params3$`Z[2,3]`,
         params4$`Z[2,3]`)

re_rm <- c(params1$`Z[2,3]`,
           params2$`Z[3,3]`,
           params3$`Z[3,3]`,
           params4$`Z[3,3]`)

re_p1 <- c(params1$`p[1]` ,
          1-params2$`p[1]` ,
          1-params3$`p[1]` ,
          1-params4$`p[1]` )

re_p0 <- c(1 - params1[,paste0("p[",1:130,"]")],
  params2[,paste0("p[",1:130,"]")],
  params3[,paste0("p[",1:130,"]")],
  1 - params4[,paste0("p[",1:130,"]")]) %>% unlist 

re_p <- array(NA, 
              dim = c(4000, 130))

mpc <- fit@sim$iter-fit@sim$warmup

tmp <- array(c(re_am,
               re_as,
               re_bm,
               re_bs,
               re_rm,
               re_rs
                  ),
             dim = c(mpc * 4, 6))

tmp2 <- tmp

#par(mfrow=c(5,5))
for (i in 1:130) {
  re_p0 <- c(params1[,paste0("p[",i,"]")],
    1-params2[,paste0("p[",i,"]")],
    1-params3[,paste0("p[",i,"]")],
    1 - params4[,paste0("p[",i,"]")]) 
 # hist(re_p0)
  tmp2 <- cbind(tmp2, re_p0)
}
par(mfrow=c(1,1))

amu <- array(NA,dim = c(mpc,fit@sim$chains,dim(tmp2)[2]))
for (cc in 1:fit@sim$chains) amu[,cc,] = tmp2[(1:mpc)+(cc-1)*mpc,]
mmu <- monitor(amu, print = F, warmup = 0)
rownames(mmu) <- c("am", "as", "bm", "bs", "rm", "rs", 
                   paste0("p",1:130))

mmu %>% DT::datatable(.)

mcmc_re <- data_frame(alpha_m = amu[,,1] %>% as.vector,
             alpha_s = amu[,,2] %>% as.vector,
             beta_m = amu[,,3] %>% as.vector,
             beta_s = amu[,,4] %>% as.vector,
             gamma_m = amu[,,5] %>% as.vector,
             gamma_s = amu[,,6] %>% as.vector
             ) %>%
  mutate(iter = rep(1:1000, 4)) %>%
  mutate(chain = rep(1:4, each = 1000) %>% as.factor)

mcmc_re2 <- mcmc_re %>%
  gather(para, val, 1:6)

ggplot(mcmc_re2, aes(x = iter, y = val, col = chain)) +
  geom_line() +
  facet_wrap(~para, nrow = 3)

```


```{r}
p_est<-NULL

LMAp_mcmc <- list()
LMAs_mcmc <- list()
for (i in 1:130) {
  re_p0 <- c(params1[,paste0("p[",i,"]")],
    1-params2[,paste0("p[",i,"]")],
    1-params3[,paste0("p[",i,"]")],
    1 - params4[,paste0("p[",i,"]")]) 
  p_est[i] <- median(re_p0)
  LMAp_mcmc[[i]] <- re_p0 * PA$LMA[i]
  LMAs_mcmc[[i]] <- (1 - re_p0) * PA$LMA[i]
}

names(p_est) <- paste0("p",1:130)

LMAps_cor <- NULL
for (i in 1:4000){
  LMAps_cor[i] <- cor(
  sapply(LMAp_mcmc, "[[", i) %>% log,
  sapply(LMAs_mcmc, "[[", i) %>% log)
}


Mu_mcmc <- extract(fit, pars = "Mu")[[1]] 

preLL <- Mu_mcmc[,,2] %>% apply(., 2, median) %>% exp

hist(LMAps_cor)
#mu2_vec <- paste("mu2[", 1:nrow(PA), "]", sep = "")
#Mu_vec <- paste("mu[", 1:nrow(PA), ",2]", sep = "")
temp_LMAp <- p_est * PA$LMA
temp_LMAs <- PA$LMA - temp_LMAp

PA <- PA %>%
  mutate(LMAp = temp_LMAp) %>%
  mutate(LMAs = temp_LMAs)  %>%
  mutate(preLL = preLL)

```


# Trait correlations

## A~area~, R~area~, LL

Open symbols indicate sun leaves and closed symbols indicate shade leaves. The results look nice and do not alter main conclusions.

```{r, echo = F, message = F}
PA2 <- PA %>%
  gather(LMA, val, c("LMA", "LMAp", "LMAs")) %>%
  gather(trait, val2, c("Aarea", "LL", "Rarea")) %>%
  gather(trait2, val3, c("Narea","Parea","cell_area","cell_vol"))


ggplot(PA2, aes(x = val, y = val2, fill = site2, col = site2, shape = strata)) +
  geom_point() +
  facet_grid(trait ~ LMA, scale = "free") +
  xlab("LMA and LD") +
  ylab("Traits") +
  scale_shape_manual(values = c("CAN" = 1, "UNDER" = 21)) +
  scale_fill_manual(values = c("DRY" = "#8B7355", "WET" = "#1874CD")) +
  scale_colour_manual(values = c("DRY" = "#8B7355", "WET" = "#1874CD")) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))




```

## Observed LL vs predicted LL

Note: Light and site effects were not included in this model.

```{r, echo = F, message = F}


my_breaks <- function(...){
  c(0.002, 0.005, 0.02, 0.05, 0.1, 0.2,  0.5, 1, 2, 5, 10, 20, 50, 100, 200, 500)
}


ggplot(PA, aes(x = LL, y = preLL, fill = site2, col = site2, shape = strata)) +
  geom_point() +
  xlab("obsLL") +
  ylab("preLL") +
  scale_shape_manual(values = c("CAN" = 1, "UNDER" = 21)) +
  scale_fill_manual(values = c("DRY" = "#8B7355", "WET" = "#1874CD")) +
  scale_colour_manual(values = c("DRY" = "#8B7355", "WET" = "#1874CD")) +
  geom_abline(aes(slope = 1, intercept = 0), 
              lty = 2, 
              lwd = 0.25, 
              col = "gray20") +
  scale_x_log10(breaks = my_breaks(), expand = c(0.1, 0),
                limits = c(min(PA$LL), max(PA$LL))) +
  scale_y_log10(breaks = my_breaks(), expand = c(0.1, 0), 
                limits = c(min(PA$LL), max(PA$LL))) +
  coord_fixed()
cor.test(log(PA$LL), log(PA$preLL))



```


## Cell~area~, Cell~vol~ (= cellulose density), N~area~, P~area~

```{r, echo = F, message = F}
ggplot(PA2, aes(x = val, y = val3, fill = site2, col = site2, shape = strata)) +
  geom_point() +
  facet_grid(trait2 ~ LMA, scale = "free") +
  xlab("LMA and LD") +
  ylab("Traits") +
  scale_shape_manual(values = c("CAN" = 1, "UNDER" = 21)) +
  scale_fill_manual(values = c("DRY" = "#8B7355", "WET" = "#1874CD")) +
  scale_colour_manual(values = c("DRY" = "#8B7355", "WET" = "#1874CD")) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))
```

## LMAp vs LMAs

```{r, echo = F, message = F}
ggplot(PA, aes(x = LMAp, y = LMAs, fill = site2, col = site2, shape = strata)) +
  geom_point() +
  xlab("LMAm") +
  ylab("LMAs") +
  scale_shape_manual(values = c("CAN" = 1, "UNDER" = 21)) +
  scale_fill_manual(values = c("DRY" = "#8B7355", "WET" = "#1874CD")) +
  scale_colour_manual(values = c("DRY" = "#8B7355", "WET" = "#1874CD")) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))
```

## Trait means

The results look nice and do not alter main conclusions.

```{r, echo = F, message = F}

ggplot(PA2 %>% filter(LMA != "LDs"), aes(x = site_strata, y = val,
                                         fill = site_strata,
                                         col = site_strata)) +
  geom_boxplot() +
  facet_grid(.~LMA, scale = "free") +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_fill_manual(values = c("DRY_CAN" = "white",
                               "WET_CAN" = "white",
                               "DRY_UNDER" = "#8B7355",
                               "WET_UNDER" = "#1874CD"
                               )) +
  scale_colour_manual(values = c("DRY_CAN" = "#8B7355",
                               "WET_CAN" = "#1874CD",
                               "DRY_UNDER" = "black",
                               "WET_UNDER" = "black"
                               ))


```


# Decomposition of LMA variation 

1) var(LMAm) & var(LMAs) & 2*cov(LMAm, LMAs), 2)relative contribution: var(LMAm)/(var(LMAm) + var(LMAs)) & var(LMAs) /(var(LMAm) + var(LMAs))

```{r, echo = F}

ggplot(var_dat_s2, aes(x = strata, y = val, fill = var)) +
  geom_col() +
  ylab("val (%)") 


ggplot(var_dat_s, aes(x = strata, y = val, fill = var)) +
  geom_bar(position = "fill",stat = "identity") +
  scale_y_continuous(labels = percent_format())

```


```{r, echo = F, message = F, eval = F}

cor.test(log(PA$LMA), log(PA$cell_area))
cor.test(log(PA$LMAs), log(PA$cell_area))

cor.test(log(PA$LD), log(PA$cell_vol))
cor.test(log(PA$LDs), log(PA$cell_vol))


cor.test(log(PA$LMA), log(PA$Narea))
cor.test(log(PA$LMAp), log(PA$Narea))

cor.test(log(PA$LMA), log(PA$Parea))
cor.test(log(PA$LMAp), log(PA$Parea))

apply(log_LMAp, 1, function(x)cor(x, log(PA$Narea))) %>% hist(main = "LMAp vs Narea, R")
apply(log_LMAp, 1, function(x)cor(x, log(PA$Parea))) %>% summary

apply(log_LMAp, 1, function(x)cor(x, log(PA$Narea))) %>% hist(main = "LMAp vs Narea, R")

apply(log_LMAp, 1, function(x)cor(x, log(PA$Rarea))) %>% quantile(0.025)
cor.test(log(PA$LMA), log(PA$Rarea))

apply(log_LMAs, 1, function(x)cor(x, log(PA$cell_area))) %>% hist(main = "LMAp vs Narea, R")

```


```{r, echo = F, message = F, eval = F}

ggplot(tmp3, aes(x = LD_shade, y = LD_sun)) +
  geom_point() +
  facet_grid(~site) +
  geom_abline(slope=1, lty = 2) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  coord_fixed()


ggplot(tmp3, aes(x = cell_vol_shade, y = cell_vol_sun)) +
  geom_point() +
  facet_grid(~site) +
  geom_abline(slope=1, lty = 2) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  coord_fixed()


ggplot(PA, aes(x = site_strata, y = cell_vol,
               fill = site_strata,
               col = site_strata)) +
  geom_boxplot() +
  #geom_violin() +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_fill_manual(values = c("DRY_CAN" = "white",
                               "WET_CAN" = "white",
                               "DRY_UNDER" = "#8B7355",
                               "WET_UNDER" = "#1874CD"
                               )) +
  scale_colour_manual(values = c("DRY_CAN" = "#8B7355",
                               "WET_CAN" = "#1874CD",
                               "DRY_UNDER" = "black",
                               "WET_UNDER" = "black"
                               ))

ggplot(PA, aes(x = site_strata, y = LD,
               fill = site_strata,
               col = site_strata)) +
  geom_boxplot() +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_fill_manual(values = c("DRY_CAN" = "white",
                               "WET_CAN" = "white",
                               "DRY_UNDER" = "#8B7355",
                               "WET_UNDER" = "#1874CD"
                               )) +
  scale_colour_manual(values = c("DRY_CAN" = "#8B7355",
                               "WET_CAN" = "#1874CD",
                               "DRY_UNDER" = "black",
                               "WET_UNDER" = "black"
                               ))



ggplot(PA, aes(x = site_strata, y = cell_area,
               fill = site_strata,
               col = site_strata)) +
  #geom_violin() +
  geom_boxplot() +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_fill_manual(values = c("DRY_CAN" = "white",
                               "WET_CAN" = "white",
                               "DRY_UNDER" = "#8B7355",
                               "WET_UNDER" = "#1874CD"
                               )) +
  scale_colour_manual(values = c("DRY_CAN" = "#8B7355",
                               "WET_CAN" = "#1874CD",
                               "DRY_UNDER" = "black",
                               "WET_UNDER" = "black"
                               ))


ggplot(PA, aes(x = LD, y = Aarea, fill = site, col = site, shape = strata)) +
  geom_point()  +
  facet_grid(site ~ strata)  +
  scale_shape_manual(values = c("CAN" = 1, "UNDER" = 21)) +
  scale_fill_manual(values = c("PNM" = "#8B7355", "PNSL" = "#1874CD")) +
  scale_colour_manual(values = c("PNM" = "#8B7355", "PNSL" = "#1874CD")) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))

ggplot(PA, aes(x = LD, y = Aarea, fill = site, col = site, shape = strata)) +
  geom_point()  +
  facet_grid( ~ strata)  +
  scale_shape_manual(values = c("CAN" = 1, "UNDER" = 21)) +
  scale_fill_manual(values = c("PNM" = "#8B7355", "PNSL" = "#1874CD")) +
  scale_colour_manual(values = c("PNM" = "#8B7355", "PNSL" = "#1874CD")) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))


ggplot(PA, aes(x = cell_vol, y = Aarea, fill = site, col = site, shape = strata)) +
  geom_point()  +
  facet_grid(site ~ strata)  +
  scale_shape_manual(values = c("CAN" = 1, "UNDER" = 21)) +
  scale_fill_manual(values = c("PNM" = "#8B7355", "PNSL" = "#1874CD")) +
  scale_colour_manual(values = c("PNM" = "#8B7355", "PNSL" = "#1874CD")) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))


ggplot(PA, aes(x = cell_vol, y = Aarea, fill = site, col = site, shape = strata)) +
  geom_point()  +
  facet_grid( ~ strata)  +
  scale_shape_manual(values = c("CAN" = 1, "UNDER" = 21)) +
  scale_fill_manual(values = c("PNM" = "#8B7355", "PNSL" = "#1874CD")) +
  scale_colour_manual(values = c("PNM" = "#8B7355", "PNSL" = "#1874CD")) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))


ggplot(PA, aes(x = cell_area, y = Aarea, fill = site, col = site, shape = strata)) +
  geom_point()  +
  facet_grid(site ~ strata)  +
  scale_shape_manual(values = c("CAN" = 1, "UNDER" = 21)) +
  scale_fill_manual(values = c("PNM" = "#8B7355", "PNSL" = "#1874CD")) +
  scale_colour_manual(values = c("PNM" = "#8B7355", "PNSL" = "#1874CD")) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))


 lm(log(Aarea) ~ log(LMAp) + log(LMAs/LT), PA) %>% summary


PA %>%
  filter(strata == "CAN") %>%
  filter(site2 == "DRY") %>%
  arrange(desc(cell_area)) %>%
  dplyr::select(sp, cell_area, LMAs)


```
