---
title: Panama results using MVN (Kikuzawa model)
author: Masatoshi Katabuchi
date: "`r format(Sys.time(), '%B %d, %Y')`"
---

More species than LT dataset

- alpha_0 LMAp^alpha_p 
- beta_0 {LMALMAs/(qA-R)}^0.5

Those update include:

- y = ax to y = ax^b^
    - Originally we were using y = ax, which is simple but not biologically correct and could produce some degrees of correlations even in randomized dataset.
    - The new model includes scaling parameters and now randomized dataset never produced correlations. We do not have to mention randomized dataset.
- LMAs / LT (= leaf structural density)
    - Leaf density is more relevant to LL compared to LMA, so using (LMAs/LT) is more straightforward. Leaves are now three-dimensional, which is also improvement of Osnas's PNAS.
- Effect of LMAs/LT to photosynthesis
    - The parameters $\alpha_s$ in Eq. 1 is the effect of leaf structural density to gross photosynthesis.
    - Based on effects of cell-wall thickness on mesophyll conductance, we expect $\alpha_s$ < 0; i.e., leaves with higher LMAs and/or higher LD have lower Aarea for a given LMAp.


```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=TRUE,
                      warning=FALSE,
                      message=FALSE)
```


```{r, echo = F, message = F}
library(loo)
library(rstan)
library(dplyr)
library(scales)
library(memisc)

theme_set(theme_light())

```


```{r, echo = F, message = F}

#load("./data/PA_latent_model_light_more_obs.rda")
load("../data/PA_latent_model_light2_more_obs.rda")
#
#load("./data/PA_latent_model_more_obs.rda")
#
#load("./data/PA_latent_model_site_more_obs.rda")
#load("../data/PA_latent_model_site_more_obs.rda")

#load("./data/PA_latent_model_light2_more_obs.rda")

pairs(res, pars = c("Z[1,1]", "Z[2,1]", "Z[3,1]", "lp__"))
pairs(res, pars = c("Z[1,2]", "Z[2,2]", "Z[3,2]", "lp__"))
pairs(res, pars = c("Z[1,3]", "Z[2,3]", "Z[3,3]", "lp__"))

traceplot(res, pars = "Z", ncol = 3)

traceplot(res, pars = c("p[1]",
                        "p[2]",
                        "lp__"
                        ))

traceplot(res, pars = "L_Omega")

summary_PA_LDL <- data.frame(summary(res)$summary)

summary_PA_LDL %>%
  round(3) %>%
  DT::datatable(.)

PA <- dat %>%
  as_data_frame %>%
  mutate(site2 = ifelse(site == "PNM", "DRY", "WET")) %>%
  mutate(sp_site_strata = paste(sp, site2, strata, sep = "_")) %>%
  mutate(site_strata = paste(site2, strata, sep = "_"))


P_vec <- paste("p[", 1:nrow(PA), "]", sep = "")
#mu2_vec <- paste("mu2[", 1:nrow(PA), "]", sep = "")
#Mu_vec <- paste("mu[", 1:nrow(PA), ",2]", sep = "")
temp_LMAp <- summary_PA_LDL[P_vec, "mean"] * PA$LMA
temp_LMAs <- PA$LMA - temp_LMAp
#temp_LL <- exp(summary_PA_LDL[mu2_vec, "mean"])


PA <- PA %>%
  mutate(LMAp = temp_LMAp) %>%
  mutate(LMAs = temp_LMAs) 


```


# Posterior distributions of $\alpha_s$ > 0


# Paramters 

```{r, echo = F, eval = T}

plot(res)

traceplot(res)

```


```{r, echo = F}

var_p <- function(data)cov(data$LMA, data$LMAp)
var_s <- function(data)cov(data$LMA, data$LMAs)
var_t <- function(data)var(data$LMA)

var_p <- function(data)var(data$LMAp)
var_s <- function(data)var(data$LMAs)
cov_ps <- function(data)cov(data$LMAp, data$LMAs)
var_t <- function(data)var(data$LMA)

var_dat <- PA %>%
  group_by(strata) %>%
  nest %>%
  mutate(var_LMAm = map_dbl(data, var_p)) %>%
  mutate(var_LMAs = map_dbl(data, var_s)) %>%
  mutate(cov_ms = 2 * map_dbl(data, cov_ps)) %>%
  mutate(var_total = map_dbl(data, var_t)) %>%
  dplyr::select(-data)

var_dat_s <- var_dat %>%
  mutate(LMAm_variance = var_LMAm / (var_LMAm + var_LMAs) * 100) %>%
  mutate(LMAs_variance = var_LMAs / (var_LMAm + var_LMAs) * 100) %>%
  gather(var, val, c(LMAm_variance, LMAs_variance))

var_dat_s2 <- var_dat %>%
  mutate(LMAm_variance = var_LMAm / (var_total) * 100) %>%
  mutate(LMAs_variance = var_LMAs / (var_total) * 100) %>%
  mutate(cov_LMAm_LMAs = cov_ms / (var_total) * 100) %>%
  gather(var, val, c(LMAm_variance, LMAs_variance, cov_LMAm_LMAs)) 


```



```{r, echo = F, eval = F}
# R values

#R values are based on posterior distributions of LMAs

par(mfrow = c(1,2))
apply(log_LMAs, 1, function(x)cor(x, log(PA$Aarea))) %>% hist(main = "LMAs vs Aarea, R")
apply(log_LMAs, 1, function(x)cor(x, log(PA$Aarea))^2) %>% hist(main = "LMAs vs Aarea, R2")

apply(log_LMAs, 1, function(x)cor(x, log(PA$Aarea + PA$Rarea))) %>% hist(main = "LMAs vs Aarea + Rarea, R")
apply(log_LMAs, 1, function(x)cor(x, log(PA$Aarea + PA$Rarea))^2) %>% hist(main = "LMAs vs Aarea + Rarea, R2")


apply(log_LDs, 1, function(x)cor(x, log(PA$Aarea))) %>% hist(main = "LMAs/LT vs Aarea, R")
apply(log_LDs, 1, function(x)cor(x, log(PA$Aarea))^2) %>% hist(main = "LMAs/LT vs Aarea, R2")


apply(log_LDs, 1, function(x)cor(x, log(PA$Aarea + PA$Rarea))) %>% hist(main = "LMAs/LT vs Aarea + Rarea, R")
apply(log_LDs, 1, function(x)cor(x, log(PA$Aarea + PA$Rarea))^2) %>% hist(main = "LMAs/LT vs Aarea + Rarea, R2")
```

```{r}

fit <-res

#fit %>% as.array %>% dim

params <- extract(fit)

m <- 4000

library(mvtnorm)

Z_i <- params$Z[,1:4,1:3]


plot(1:m, Z_i[,2,1], type = "l", col ="orange",ylim=c(0,0.8))
points(1:m, Z_i[,3,1], type = "l", col ="violet")

plot(1:m, Z_i[,2,2], type = "l", col ="orange",ylim=c(-1, 0.5))
points(1:m, Z_i[,3,2], type = "l", col ="violet")

plot(1:m, Z_i[,2,3], type = "l", col ="orange",ylim=c(-0.5,1.5))
points(1:m, Z_i[,3,3], type = "l", col ="violet")

plot(1:m, Z_i[,4,2], type = "l", col ="orange")

lp_m <- params$lp__
p_m <- params$p
rho13 <- params$rho13

plot(1:m, rho13, type = "l", col ="orange")
plot(1:m, lp_m, type = "l", col ="orange")
plot(1:m, p_m[,1], type = "l", col ="orange")


Sigma_i <- params$Sigma[,1:3,1:3]
L_sigma_i <- params$L_sigma[,]

```

# para fix

```{r}
mcmc_res <- extract(fit) %>% as.data.frame
params1 <- as.data.frame(extract(fit, permuted=FALSE)[,1,])
params2 <- as.data.frame(extract(fit, permuted=FALSE)[,2,])
params3 <- as.data.frame(extract(fit, permuted=FALSE)[,3,])
params4 <- as.data.frame(extract(fit, permuted=FALSE)[,4,])


get_stancode(fit) %>% cat

re_am <- c(params1$`Z[2,1]`,
	 			 params2$`Z[3,1]`,
	 			 params3$`Z[2,1]`,
	 			 params4$`Z[2,1]`)

re_as <- c(params1$`Z[3,1]`,
	 			 params2$`Z[2,1]`,
	 			 params3$`Z[3,1]`,
	 			 params4$`Z[3,1]`)

re_bm <- c(params1$`Z[2,2]`,
	 			 params2$`Z[3,2]`,
         params3$`Z[2,2]`,
         params4$`Z[2,2]`)

re_bs <- c(params1$`Z[3,2]`,
           params2$`Z[2,2]`,
           params3$`Z[3,2]`,
           params4$`Z[3,2]`)

re_rs <- c(params1$`Z[3,3]`,
	 			 params2$`Z[2,3]`,
         params3$`Z[3,3]`,
         params4$`Z[3,3]`)

re_rm <- c(params1$`Z[2,3]`,
           params2$`Z[3,3]`,
           params3$`Z[2,3]`,
           params4$`Z[2,3]`)

re_p1 <- c(params1$`p[1]` ,
          1-params2$`p[1]` ,
          params3$`p[1]` ,
          params4$`p[1]` )

re_p0 <- c(params1[,paste0("p[",1:130,"]")],
  1-params2[,paste0("p[",1:130,"]")],
  params3[,paste0("p[",1:130,"]")],
  params4[,paste0("p[",1:130,"]")]) %>% unlist 

re_p <- array(NA, 
              dim = c(4000, 130))

mpc <- fit@sim$iter-fit@sim$warmup

tmp <- array(c(re_am,
               re_as,
               re_bm,
               re_bs,
               re_rm,
               re_rs
                  ),
             dim = c(mpc * 4, 6))

tmp2 <- tmp

#par(mfrow=c(5,5))
for (i in 1:130) {
  re_p0 <- c(params1[,paste0("p[",i,"]")],
    1-params2[,paste0("p[",i,"]")],
    params3[,paste0("p[",i,"]")],
    params4[,paste0("p[",i,"]")]) 
 # hist(re_p0)
  tmp2 <- cbind(tmp2, re_p0)
}
par(mfrow=c(1,1))

amu <- array(NA,dim = c(mpc,fit@sim$chains,dim(tmp2)[2]))
for (cc in 1:fit@sim$chains) amu[,cc,] = tmp2[(1:mpc)+(cc-1)*mpc,]
mmu <- monitor(amu, print = F, warmup = 0)
rownames(mmu) <- c("am", "as", "bm", "bs", "rm", "rs", 
                   paste0("p",1:130))

mmu %>% DT::datatable(.)

mcmc_re <- data_frame(alpha_m = amu[,,1] %>% as.vector,
             alpha_s = amu[,,2] %>% as.vector,
             beta_m = amu[,,3] %>% as.vector,
             beta_s = amu[,,4] %>% as.vector,
             gamma_m = amu[,,5] %>% as.vector,
             gamma_s = amu[,,6] %>% as.vector
             ) %>%
  mutate(iter = rep(1:1000, 4)) %>%
  mutate(chain = rep(1:4, each = 1000) %>% as.factor)

mcmc_re2 <- mcmc_re %>%
  gather(para, val, 1:6)

ggplot(mcmc_re2, aes(x = iter, y = val, col = chain)) +
  geom_line() +
  facet_wrap(~para, nrow = 3)

```


```{r}
p_est<-NULL

LMAp_mcmc <- list()
LMAs_mcmc <- list()
for (i in 1:130) {
  re_p0 <- c(params1[,paste0("p[",i,"]")],
    1-params2[,paste0("p[",i,"]")],
    params3[,paste0("p[",i,"]")],
    params4[,paste0("p[",i,"]")]) 
  p_est[i] <- median(re_p0)
  LMAp_mcmc[[i]] <- re_p0 * PA$LMA[i]
  LMAs_mcmc[[i]] <- (1 - re_p0) * PA$LMA[i]
}

names(p_est) <- paste0("p",1:130)

LMAps_cor <- NULL
for (i in 1:4000){
  LMAps_cor[i] <- cor(
  sapply(LMAp_mcmc, "[[", i) %>% log,
  sapply(LMAs_mcmc, "[[", i) %>% log)
}


Mu_mcmc <- extract(fit, pars = "Mu")[[1]] 

preLL <- Mu_mcmc[,,2] %>% apply(., 2, median) %>% exp

hist(LMAps_cor)
#mu2_vec <- paste("mu2[", 1:nrow(PA), "]", sep = "")
#Mu_vec <- paste("mu[", 1:nrow(PA), ",2]", sep = "")
temp_LMAp <- p_est * PA$LMA
temp_LMAs <- PA$LMA - temp_LMAp

#PA <- PA %>%
#  mutate(LMAp = temp_LMAp) %>%
#  mutate(LMAs = temp_LMAs)  %>%
#  mutate(preLL = preLL)

Mu_mcmc <- extract(fit, pars = "Mu")[[1]]

preLL <- Mu_mcmc[,,2] %>% apply(., 2, mean) %>% exp
preA <- Mu_mcmc[,,1] %>% apply(., 2, mean) %>% exp
preR <- Mu_mcmc[,,3] %>% apply(., 2, mean) %>% exp

preLL_lo <- Mu_mcmc[,,2] %>% apply(., 2, function(x)quantile(x, 0.025)) %>% exp
preLL_up <- Mu_mcmc[,,2] %>% apply(., 2, function(x)quantile(x, 0.975)) %>% exp
preA_lo <- Mu_mcmc[,,1] %>% apply(., 2, function(x)quantile(x, 0.025)) %>% exp
preA_up <- Mu_mcmc[,,1] %>% apply(., 2, function(x)quantile(x, 0.975)) %>% exp
preR_lo <- Mu_mcmc[,,3] %>% apply(., 2, function(x)quantile(x, 0.025)) %>% exp
preR_up <- Mu_mcmc[,,3] %>% apply(., 2, function(x)quantile(x, 0.975)) %>% exp
preLL1 <- Mu_mcmc[1,,2] %>% exp

PA <- PA %>%
  mutate(LMAp = temp_LMAp) %>%
  mutate(LMAs = temp_LMAs)  %>%
  mutate(preLL = preLL) %>%
  mutate(preA = preA) %>%
  mutate(preR = preR) %>%
  mutate(preLL_lo = preLL_lo) %>%
  mutate(preLL_up = preLL_up) %>%
  mutate(preA_lo = preA_lo) %>%
  mutate(preA_up = preA_up) %>%
  mutate(preR_lo = preR_lo) %>%
  mutate(preR_up = preR_up) %>%
  mutate(preLL1 = preLL1)

write.csv(PA,"../data/PA_res.csv", row.names = F)
#PA <- read_csv("./data/PA_res.csv")


```


# Trait correlations

## A~area~, R~area~, LL

Open symbols indicate sun leaves and closed symbols indicate shade leaves. The results look nice and do not alter main conclusions.

```{r, echo = F, message = F}
PA2 <- PA %>%
  gather(LMA, val, c("LMA", "LMAp", "LMAs")) %>%
  gather(trait, val2, c("Aarea", "LL", "Rarea")) %>%
  gather(trait2, val3, c("Narea","Parea","cell_area","cell_vol"))


ggplot(PA2, aes(x = val, y = val2, fill = site2, col = site2, shape = strata)) +
  geom_point() +
  facet_grid(trait ~ LMA, scale = "free") +
  xlab("LMA and LD") +
  ylab("Traits") +
  scale_shape_manual(values = c("CAN" = 1, "UNDER" = 21)) +
  scale_fill_manual(values = c("DRY" = "#8B7355", "WET" = "#1874CD")) +
  scale_colour_manual(values = c("DRY" = "#8B7355", "WET" = "#1874CD")) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))




```

## Observed LL vs predicted LL

Note: Light and site effects were not included in this model.

```{r, echo = F, message = F}


my_breaks <- function(...){
  c(0.002, 0.005, 0.02, 0.05, 0.1, 0.2,  0.5, 1, 2, 5, 10, 20, 50, 100, 200, 500)
}


ggplot(PA, aes(x = LL, y = preLL, fill = site2, col = site2, shape = strata)) +
  geom_point() +
  xlab("obsLL") +
  ylab("preLL") +
  scale_shape_manual(values = c("CAN" = 1, "UNDER" = 21)) +
  scale_fill_manual(values = c("DRY" = "#8B7355", "WET" = "#1874CD")) +
  scale_colour_manual(values = c("DRY" = "#8B7355", "WET" = "#1874CD")) +
  geom_abline(aes(slope = 1, intercept = 0), 
              lty = 2, 
              lwd = 0.25, 
              col = "gray20") +
  scale_x_log10(breaks = my_breaks(), expand = c(0.1, 0),
                limits = c(min(PA$LL), max(PA$LL))) +
  scale_y_log10(breaks = my_breaks(), expand = c(0.1, 0), 
                limits = c(min(PA$LL), max(PA$LL))) +
  coord_fixed()
cor.test(log(PA$LL), log(PA$preLL))



```


## Cell~area~, Cell~vol~ (= cellulose density), N~area~, P~area~

```{r, echo = F, message = F}
ggplot(PA2, aes(x = val, y = val3, fill = site2, col = site2, shape = strata)) +
  geom_point() +
  facet_grid(trait2 ~ LMA, scale = "free") +
  xlab("LMA and LD") +
  ylab("Traits") +
  scale_shape_manual(values = c("CAN" = 1, "UNDER" = 21)) +
  scale_fill_manual(values = c("DRY" = "#8B7355", "WET" = "#1874CD")) +
  scale_colour_manual(values = c("DRY" = "#8B7355", "WET" = "#1874CD")) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))
```

## LMAp vs LMAs

```{r, echo = F, message = F}
ggplot(PA, aes(x = LMAp, y = LMAs, fill = site2, col = site2, shape = strata)) +
  geom_point() +
  xlab("LMAm") +
  ylab("LMAs") +
  scale_shape_manual(values = c("CAN" = 1, "UNDER" = 21)) +
  scale_fill_manual(values = c("DRY" = "#8B7355", "WET" = "#1874CD")) +
  scale_colour_manual(values = c("DRY" = "#8B7355", "WET" = "#1874CD")) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))
```

## Trait means

The results look nice and do not alter main conclusions.

```{r, echo = F, message = F}

ggplot(PA2 %>% filter(LMA != "LDs"), aes(x = site_strata, y = val,
                                         fill = site_strata,
                                         col = site_strata)) +
  geom_boxplot() +
  facet_grid(.~LMA, scale = "free") +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_fill_manual(values = c("DRY_CAN" = "white",
                               "WET_CAN" = "white",
                               "DRY_UNDER" = "#8B7355",
                               "WET_UNDER" = "#1874CD"
                               )) +
  scale_colour_manual(values = c("DRY_CAN" = "#8B7355",
                               "WET_CAN" = "#1874CD",
                               "DRY_UNDER" = "black",
                               "WET_UNDER" = "black"
                               ))


```


# Decomposition of LMA variation 

1) var(LMAm) & var(LMAs) & 2*cov(LMAm, LMAs), 2)relative contribution: var(LMAm)/(var(LMAm) + var(LMAs)) & var(LMAs) /(var(LMAm) + var(LMAs))

```{r, echo = F}

ggplot(var_dat_s2, aes(x = strata, y = val, fill = var)) +
  geom_col() +
  ylab("val (%)") 


ggplot(var_dat_s, aes(x = strata, y = val, fill = var)) +
  geom_bar(position = "fill",stat = "identity") +
  scale_y_continuous(labels = percent_format())

```


```{r, echo = F, message = F, eval = F}

cor.test(log(PA$LMA), log(PA$cell_area))
cor.test(log(PA$LMAs), log(PA$cell_area))

cor.test(log(PA$LD), log(PA$cell_vol))
cor.test(log(PA$LDs), log(PA$cell_vol))


cor.test(log(PA$LMA), log(PA$Narea))
cor.test(log(PA$LMAp), log(PA$Narea))

cor.test(log(PA$LMA), log(PA$Parea))
cor.test(log(PA$LMAp), log(PA$Parea))

apply(log_LMAp, 1, function(x)cor(x, log(PA$Narea))) %>% hist(main = "LMAp vs Narea, R")
apply(log_LMAp, 1, function(x)cor(x, log(PA$Parea))) %>% summary

apply(log_LMAp, 1, function(x)cor(x, log(PA$Narea))) %>% hist(main = "LMAp vs Narea, R")

apply(log_LMAp, 1, function(x)cor(x, log(PA$Rarea))) %>% quantile(0.025)
cor.test(log(PA$LMA), log(PA$Rarea))

apply(log_LMAs, 1, function(x)cor(x, log(PA$cell_area))) %>% hist(main = "LMAp vs Narea, R")

```


```{r, echo = F, message = F, eval = F}

ggplot(tmp3, aes(x = LD_shade, y = LD_sun)) +
  geom_point() +
  facet_grid(~site) +
  geom_abline(slope=1, lty = 2) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  coord_fixed()


ggplot(tmp3, aes(x = cell_vol_shade, y = cell_vol_sun)) +
  geom_point() +
  facet_grid(~site) +
  geom_abline(slope=1, lty = 2) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  coord_fixed()


ggplot(PA, aes(x = site_strata, y = cell_vol,
               fill = site_strata,
               col = site_strata)) +
  geom_boxplot() +
  #geom_violin() +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_fill_manual(values = c("DRY_CAN" = "white",
                               "WET_CAN" = "white",
                               "DRY_UNDER" = "#8B7355",
                               "WET_UNDER" = "#1874CD"
                               )) +
  scale_colour_manual(values = c("DRY_CAN" = "#8B7355",
                               "WET_CAN" = "#1874CD",
                               "DRY_UNDER" = "black",
                               "WET_UNDER" = "black"
                               ))

ggplot(PA, aes(x = site_strata, y = LD,
               fill = site_strata,
               col = site_strata)) +
  geom_boxplot() +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_fill_manual(values = c("DRY_CAN" = "white",
                               "WET_CAN" = "white",
                               "DRY_UNDER" = "#8B7355",
                               "WET_UNDER" = "#1874CD"
                               )) +
  scale_colour_manual(values = c("DRY_CAN" = "#8B7355",
                               "WET_CAN" = "#1874CD",
                               "DRY_UNDER" = "black",
                               "WET_UNDER" = "black"
                               ))



ggplot(PA, aes(x = site_strata, y = cell_area,
               fill = site_strata,
               col = site_strata)) +
  #geom_violin() +
  geom_boxplot() +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_fill_manual(values = c("DRY_CAN" = "white",
                               "WET_CAN" = "white",
                               "DRY_UNDER" = "#8B7355",
                               "WET_UNDER" = "#1874CD"
                               )) +
  scale_colour_manual(values = c("DRY_CAN" = "#8B7355",
                               "WET_CAN" = "#1874CD",
                               "DRY_UNDER" = "black",
                               "WET_UNDER" = "black"
                               ))


ggplot(PA, aes(x = LD, y = Aarea, fill = site, col = site, shape = strata)) +
  geom_point()  +
  facet_grid(site ~ strata)  +
  scale_shape_manual(values = c("CAN" = 1, "UNDER" = 21)) +
  scale_fill_manual(values = c("PNM" = "#8B7355", "PNSL" = "#1874CD")) +
  scale_colour_manual(values = c("PNM" = "#8B7355", "PNSL" = "#1874CD")) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))

ggplot(PA, aes(x = LD, y = Aarea, fill = site, col = site, shape = strata)) +
  geom_point()  +
  facet_grid( ~ strata)  +
  scale_shape_manual(values = c("CAN" = 1, "UNDER" = 21)) +
  scale_fill_manual(values = c("PNM" = "#8B7355", "PNSL" = "#1874CD")) +
  scale_colour_manual(values = c("PNM" = "#8B7355", "PNSL" = "#1874CD")) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))


ggplot(PA, aes(x = cell_vol, y = Aarea, fill = site, col = site, shape = strata)) +
  geom_point()  +
  facet_grid(site ~ strata)  +
  scale_shape_manual(values = c("CAN" = 1, "UNDER" = 21)) +
  scale_fill_manual(values = c("PNM" = "#8B7355", "PNSL" = "#1874CD")) +
  scale_colour_manual(values = c("PNM" = "#8B7355", "PNSL" = "#1874CD")) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))


ggplot(PA, aes(x = cell_vol, y = Aarea, fill = site, col = site, shape = strata)) +
  geom_point()  +
  facet_grid( ~ strata)  +
  scale_shape_manual(values = c("CAN" = 1, "UNDER" = 21)) +
  scale_fill_manual(values = c("PNM" = "#8B7355", "PNSL" = "#1874CD")) +
  scale_colour_manual(values = c("PNM" = "#8B7355", "PNSL" = "#1874CD")) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))


ggplot(PA, aes(x = cell_area, y = Aarea, fill = site, col = site, shape = strata)) +
  geom_point()  +
  facet_grid(site ~ strata)  +
  scale_shape_manual(values = c("CAN" = 1, "UNDER" = 21)) +
  scale_fill_manual(values = c("PNM" = "#8B7355", "PNSL" = "#1874CD")) +
  scale_colour_manual(values = c("PNM" = "#8B7355", "PNSL" = "#1874CD")) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))


 lm(log(Aarea) ~ log(LMAp) + log(LMAs/LT), PA) %>% summary


PA %>%
  filter(strata == "CAN") %>%
  filter(site2 == "DRY") %>%
  arrange(desc(cell_area)) %>%
  dplyr::select(sp, cell_area, LMAs)


```

# PA LES plot ----------------------------------------------------------------0
PA_dat1 <- PA_dat %>%
  filter(Trait == "LL" |
         Trait == "Aarea" |
         Trait == "Rarea")

lim_PA <- lim_func(PA_dat1 %>% filter(site_strata != "Rand"),
                   trait = "Val")
lim_PA2 <- lim_func(PA_dat1 %>% filter(site_strata != "Rand"),
                    trait = "Val2", LMA = FALSE)

lab1 <- data_frame(lab = paste("(", letters[1:9], ")", sep = ""),
                   Val2 = lim_PA2$max_val %>% rep(each = 3),
                   Val = lim_PA$min_val %>% rep(3),
                   Val_max = lim_PA$max_val %>% rep(3),
                   gr = "Sun-Dry",
                   r_vals = r_vals$r_vals$PA %>% unlist,
                   Trait2 = rep(lim_PA2$Trait2, each = 3),
                   LMA = rep(lim_PA$LMA, 3)
                   )

fills <- c("Sun-Dry" = settings$fills$sun_dry,
          "Sun-Wet" = settings$fills$sun_wet,
          "Shade-Dry" = settings$fills$shade_dry,
          "Shade-Wet" = settings$fills$shade_wet)

cols <- c("Sun-Dry" = settings$colors$sun_dry,
          "Sun-Wet" = settings$colors$sun_wet,
          "Shade-Dry" = settings$colors$shade_dry,
          "Shade-Wet" = settings$colors$shade_wet)

PA_plot <- scatter_plt(PA_dat1)

print(PA_plot)

my_ggsave("./figs/PA.png", PA_plot)

```

## Observed LL vs predicted LL

Looks OK but each group has less steep slope. Our model tend to under-and over-esimate LL for long- and short-LL leaves, respectively.

```{r, echo = F, message = F}


LL_dat <- PA %>%
  mutate(site_strata = factor(site_strata,
          levels = c("WET_CAN", "DRY_CAN", "WET_UNDER", "DRY_UNDER"))) %>%
  mutate(gr = factor(site_strata,
    labels = c("Sun-Wet",
               "Sun-Dry",
               "Shade-Wet",
               "Shade-Dry"
                      )))

labLL <- data_frame(
                   LL = Inf,
                   #Val = min(PA$LL),
                   #Val_max = max(PA$LL),
                   Val = min(PA$preLL_lo),
                   Val_max = max(PA$preLL_up),
                   gr = "Sun-Dry",
                   r_vals = r_vals$r_vals$PA_LL$preLL_LL
                   )


LL_plot2 <- ggplot(LL_dat, aes(x = LL, y = preLL,
                              fill = gr, col = gr)) +
  geom_errorbar(aes(ymin = preLL_lo, ymax = preLL_up), alpha = 0.8, col = "gray") +
  geom_point(shape = 21) +
  scale_fill_manual(values = fills) +
  scale_colour_manual(values = cols) +
  scale_x_log10(breaks = my_breaks(), expand = c(0.1, 0),
                limits = c(min(labLL$Val), max(labLL$Val_max))) +
  scale_y_log10(breaks = my_breaks(), expand = c(0.1, 0),
                limits = c(min(labLL$Val), max(labLL$Val_max))) +
  ylab("Predicted LL (months)") +
  xlab("Observed LL (months)") +
  geom_abline(aes(slope = 1, intercept = 0),
              lty = 2,
              lwd = 0.25,
              col = "gray20") +
  geom_text(data = labLL, aes(label = r_vals, x = Val_max, y = 2.5),
            colour = "black",
            hjust = 1,
            vjust= 0,
            parse = TRUE,
            show.legend = FALSE) +
  coord_fixed() +
  theme_LES() +
  theme(
    legend.position = c(0.2, 0.85),
    axis.title.x = element_text(margin = margin(t = 1,
                                                b = 1,
                                                l = 0,
                                                r = 0)),
    axis.title.y = element_text(margin = margin(t = 1,
                                                b = 1,
                                                l = 1,
                                                r = 1))
        )

print(LL_plot2)

my_ggsave("../figs/LL_plot.png", LL_plot2,
          width = 6.7,
          height = 6.7)

cor.test(log(PA$LL), log(PA$preLL))

```

## A, R

```{r}

labA <- data_frame(
                   LL = Inf,
                   Val = min(PA$Aarea),
                   Val_max = max(PA$Aarea),
                   gr = "Sun-Dry",
                   r_vals = r_vals$r_vals$PA_LL$preA_Aarea
                   )

A_plot2 <- ggplot(LL_dat, aes(x = Aarea, y = preA,
                              fill = gr, col = gr)) +
  geom_errorbar(aes(ymin = preA_lo, ymax = preA_up), alpha = 0.8, col = "gray") +
  geom_point(shape = 21) +
  scale_fill_manual(values = fills) +
  scale_colour_manual(values = cols) +
  scale_x_log10(breaks = my_breaks(), expand = c(0.1, 0),
                limits = c(min(labA$Val), max(labA$Val_max))) +
  scale_y_log10(breaks = my_breaks(), expand = c(0.1, 0),
                limits = c(min(labA$Val), max(labA$Val_max))) +
  ylab("Predicted A (months)") +
  xlab("Observed A (months)") +
  geom_abline(aes(slope = 1, intercept = 0),
              lty = 2,
              lwd = 0.25,
              col = "gray20") +
  geom_text(data = labA, aes(label = r_vals, x = Val_max, y = 2.5),
            colour = "black",
            hjust = 1,
            vjust= 0,
            parse = TRUE,
            show.legend = FALSE) +
  coord_fixed() +
  theme_LES() +
  theme(
    legend.position = c(0.2, 0.85),
    axis.title.x = element_text(margin = margin(t = 1,
                                                b = 1,
                                                l = 0,
                                                r = 0)),
    axis.title.y = element_text(margin = margin(t = 1,
                                                b = 1,
                                                l = 1,
                                                r = 1))
        )

print(A_plot2)

my_ggsave("../figs/A_plot.png", A_plot2,
          width = 6.7,
          height = 6.7)

labR <- data_frame(
         LL = Inf,
         Val = min(PA$Rarea),
         Val_max = max(PA$Rarea),
         gr = "Sun-Dry",
         r_vals = r_vals$r_vals$PA_LL$preR_Rarea
         )

R_plot2 <- ggplot(LL_dat, aes(x = Rarea, y = preR,
                              fill = gr, col = gr)) +
  geom_errorbar(aes(ymin = preR_lo, ymax = preR_up), alpha = 0.8, col = "gray") +
  geom_point(shape = 21) +
  scale_fill_manual(values = fills) +
  scale_colour_manual(values = cols) +
  scale_x_log10(breaks = my_breaks(), expand = c(0.1, 0),
                limits = c(min(labR$Val), max(labR$Val_max))) +
  scale_y_log10(breaks = my_breaks(), expand = c(0.1, 0),
                limits = c(min(labR$Val), max(labR$Val_max))) +
  ylab("Predicted R (months)") +
  xlab("Observed R (months)") +
  geom_abline(aes(slope = 1, intercept = 0),
              lty = 2,
              lwd = 0.25,
              col = "gray20") +
  geom_text(data = labR, aes(label = r_vals, x = Val_max, y = 2.5),
            colour = "black",
            hjust = 1,
            vjust= 0,
            parse = TRUE,
            show.legend = FALSE) +
  coord_fixed() +
  theme_LES() +
  theme(
    legend.position = c(0.2, 0.85),
    axis.title.x = element_text(margin = margin(t = 1,
                                                b = 1,
                                                l = 0,
                                                r = 0)),
    axis.title.y = element_text(margin = margin(t = 1,
                                                b = 1,
                                                l = 1,
                                                r = 1))
        )

print(R_plot2)

my_ggsave("../figs/R_plot.png", R_plot2,
          width = 6.7,
          height = 6.7)

```

## Cell~area~, Cell~vol~ (= cellulose density), N~area~, P~area~

```{r, echo = F, message = F}


PA_dat2 <- PA_dat %>%
  filter(Trait == "Narea" |
         Trait == "Parea" |
         Trait == "cell_area")


lim_PA <- lim_func(PA_dat2 %>% filter(site_strata != "Rand"),
                   trait = "Val")
lim_PA2 <- lim_func(PA_dat2 %>% filter(site_strata != "Rand"),
                    trait = "Val2", LMA = FALSE)
lab1 <- data_frame(lab = paste("(", letters[1:9], ")", sep = ""),
                   Val2 = lim_PA2$max_val %>% rep(each = 3),
                   Val = lim_PA$min_val %>% rep(3),
                   Val_max = lim_PA$max_val %>% rep(3),
                   gr = "Sun-Dry",
                   r_vals = r_vals$r_vals$PA_NP %>% unlist,
                   Trait2 = rep(lim_PA2$Trait2, each = 3),
                   LMA = rep(lim_PA$LMA, 3)
                   )

PA_NPC_plot <- scatter_plt(PA_dat2)

print(PA_NPC_plot)

my_ggsave("../figs/PA_NPC.png", PA_NPC_plot)

```

## LMAp vs LMAs

```{r, echo = F, message = F}
ggplot(PA, aes(x = LMAp, y = LMAs, fill = site2, col = site2, shape = strata)) +
  geom_point() +
  xlab("LMAm") +
  ylab("LMAs") +
  scale_shape_manual(values = c("CAN" = 1, "UNDER" = 21)) +
  scale_fill_manual(values = c("DRY" = "#8B7355", "WET" = "#1874CD")) +
  scale_colour_manual(values = c("DRY" = "#8B7355", "WET" = "#1874CD")) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))
```

## Trait means

The results look OK.

```{r, echo = F, message = F, eval = T}

ggplot(PA2, aes(x = site_strata, y = val,
                                         fill = site_strata,
                                         col = site_strata)) +
  geom_boxplot() +
  facet_grid(.~LMA, scale = "free") +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_fill_manual(values = c("DRY_CAN" = "white",
                               "WET_CAN" = "white",
                               "DRY_UNDER" = "#8B7355",
                               "WET_UNDER" = "#1874CD"
                               )) +
  scale_colour_manual(values = c("DRY_CAN" = "#8B7355",
                               "WET_CAN" = "#1874CD",
                               "DRY_UNDER" = "black",
                               "WET_UNDER" = "black"
                               ))

```


## Decomposition of LMA variation

```{r}

var_p <- function(data)cov(data$LMA, data$LMAp)
var_s <- function(data)cov(data$LMA, data$LMAs)
var_t <- function(data)var(data$LMA)

var_p <- function(data)var(data$LMAp)
var_s <- function(data)var(data$LMAs)
cov_ps <- function(data)cov(data$LMAp, data$LMAs)
var_t <- function(data)var(data$LMA)

var_dat <- PA %>%
  group_by(strata) %>%
  nest %>%
  mutate(var_LMAm = map_dbl(data, var_p)) %>%
  mutate(var_LMAs = map_dbl(data, var_s)) %>%
  mutate(cov_ms = 2 * map_dbl(data, cov_ps)) %>%
  mutate(var_total = map_dbl(data, var_t)) %>%
  dplyr::select(-data)

var_dat_s <- var_dat %>%
  mutate(LMAm_variance = var_LMAm / (var_LMAm + var_LMAs) * 100) %>%
  mutate(LMAs_variance = var_LMAs / (var_LMAm + var_LMAs) * 100) %>%
  gather(var, val, c(LMAm_variance, LMAs_variance))

var_dat_s2 <- var_dat %>%
  mutate(LMAm_variance = var_LMAm / (var_total) * 100) %>%
  mutate(LMAs_variance = var_LMAs / (var_total) * 100) %>%
  mutate(cov_LMAm_LMAs = cov_ms / (var_total) * 100) %>%
  gather(var, val, c(LMAm_variance, LMAs_variance, cov_LMAm_LMAs))

```

1) var(LMAm) & var(LMAs) & 2*cov(LMAm, LMAs), 2)relative contribution: var(LMAm)/(var(LMAm) + var(LMAs)) & var(LMAs) /(var(LMAm) + var(LMAs))

```{r, echo = F, eval = T}

ggplot(var_dat_s2, aes(x = strata, y = val, fill = var)) +
  geom_col() +
  ylab("val (%)")


ggplot(var_dat_s, aes(x = strata, y = val, fill = var)) +
  geom_bar(position = "fill",stat = "identity") +
  scale_y_continuous(labels = percent_format())

```

# Randomized dataset

```{r}

load("../data/PA_latent_replusive_model_light2_more_rand.rda")

res <- rand_res$model[[1]]
tmp <- rand_res$data[[1]]

PA <- dat %>%
  as_data_frame %>%
  mutate(site2 = ifelse(site == "PNM", "DRY", "WET")) %>%
  mutate(sp_site_strata = paste(sp, site2, strata, sep = "_")) %>%
  mutate(site_strata = paste(site2, strata, sep = "_")) %>%
  mutate(LMA = tmp$LMA) %>%
  mutate(Aarea = tmp$A) %>%
  mutate(LL = tmp$LL) %>%
  mutate(Rarea = tmp$R)

```

## There are divergent transitions (fatal errors)

This means the model cannot reliable posterior distributions.

```{r}
pairs(res, pars = c("Z[1,1]","Z[1,2]"))
```

## No label problems?

It seems that MCMC tried to reduce the residuals for the expected LL and R~area~, which results in the failure for estimating A~area~.

```{r}


fit <- res
params <- extract(fit)
m <- 4000
Z_i <- params$Z[,1:3,1:3]

tmp <- data_frame(iter = 1:m,
           alpha_m = Z_i[,2,1],
           alpha_s = Z_i[,3,1]
           ) %>%
  gather(para, val, 2:3)

tmp_text <- data_frame(x = c(0, 1000, 2000, 3000),
                       y = 0.7,
                       lab = c("chain1->", "chain2->", "chain3->", "chain4->"))

ggplot(tmp, aes(x = iter, y = val)) +
  geom_line(aes(col = para)) +
  geom_vline(xintercept = c(0, 1000,2000,3000)) +
  geom_text(data = tmp_text, aes(x = x, y =y , label = lab))

traceplot(res, par = "Z", ncol = 3)

summary_PA <- data.frame(summary(res)$summary)

summary_PA %>%
  round(3) %>%
  DT::datatable(.)

```

add mean LMAm and LMAs

```{r}

mcmc_res <- extract(fit) %>% as.data.frame

p_est <- mcmc_res[,paste0("p.",  1:130)] %>% apply(.,2,mean)

Mu_mcmc <- extract(fit, pars = "Mu")[[1]]
preLL <- Mu_mcmc[,,2] %>% apply(., 2, mean) %>% exp
preA <- Mu_mcmc[,,1] %>% apply(., 2, mean) %>% exp
preR <- Mu_mcmc[,,3] %>% apply(., 2, mean) %>% exp

temp_LMAp <- p_est * PA$LMA
temp_LMAs <- PA$LMA - temp_LMAp

PA <- PA %>%
  mutate(LMAp = temp_LMAp) %>%
  mutate(LMAs = temp_LMAs)  %>%
  mutate(preLL = preLL) %>%
  mutate(preR = preR) %>%
  mutate(preA = preA)

plot(preR ~ Rarea, log = "xy", PA)
plot(preA ~ Aarea, log = "xy", PA)
plot(preLL ~ LL, log = "xy", PA)

plot(LL ~ preLL, PA, log= "xy")

```

## LMAm and LMAs vs N~area~, P~area~ and LL

```{r, echo = F, message = F}

setwd("..")

PA_dat <- PA %>%
  gather(LMA, Val,
         c(LMA, LMAs, LMAp)) %>%
  gather(Trait, Val2, c(Aarea, Rarea, LL,
                        Parea, Narea, cell_area)) %>%
  mutate(LMA = factor(LMA,
    labels = c("LMA", "LMAm", "LMAs"))) %>%
  mutate(LMA = factor(LMA,
    labels = c("LMA~(~g~m^{-2})", "LMAm~(~g~m^{-2})", "LMAs~(~g~m^{-2})"))) %>%
  mutate(site_strata = factor(site_strata,
          levels = c("WET_CAN", "DRY_CAN", "WET_UNDER", "DRY_UNDER"))) %>%
  mutate(Trait = factor(Trait,
    levels = c("Aarea", "Rarea", "LL", "Narea", "Parea", "cell_area"))) %>%
  mutate(Trait2 = factor(Trait,
    labels = c("italic(A)[area]~(~mu~mol~m^{-2}~s^{-1})",
               "italic(R)[area]~(~mu~mol~m^{-2}~s^{-1})",
               "LL~(months)",
               "italic(N)[area]~(~g~m^{-2})",
               "italic(P)[area]~(~g~m^{-2})",
               "italic(CL)[area]~(~g~m^{-2})"
               ))) %>%
  mutate(gr = factor(site_strata,
    labels = c("Sun-Wet",
               "Sun-Dry",
               "Shade-Wet",
               "Shade-Dry"
                      ))) %>%
  arrange(gr)

# PA LES plot ----------------------------------------------------------------0
PA_dat1 <- PA_dat %>%
  filter(Trait == "LL" |
         Trait == "Aarea" |
         Trait == "Rarea")

lim_PA <- lim_func(PA_dat1 %>% filter(site_strata != "Rand"),
                   trait = "Val")
lim_PA2 <- lim_func(PA_dat1 %>% filter(site_strata != "Rand"),
                    trait = "Val2", LMA = FALSE)

lab1 <- data_frame(lab = paste("(", letters[1:9], ")", sep = ""),
                   Val2 = lim_PA2$max_val %>% rep(each = 3),
                   Val = lim_PA$min_val %>% rep(3),
                   Val_max = lim_PA$max_val %>% rep(3),
                   gr = "Sun-Dry",
                   r_vals = "",
                   Trait2 = rep(lim_PA2$Trait2, each = 3),
                   LMA = rep(lim_PA$LMA, 3)
                   )

fills <- c("Sun-Dry" = settings$fills$sun_dry,
          "Sun-Wet" = settings$fills$sun_wet,
          "Shade-Dry" = settings$fills$shade_dry,
          "Shade-Wet" = settings$fills$shade_wet)

cols <- c("Sun-Dry" = settings$colors$sun_dry,
          "Sun-Wet" = settings$colors$sun_wet,
          "Shade-Dry" = settings$colors$shade_dry,
          "Shade-Wet" = settings$colors$shade_wet)

PA_plot <- scatter_plt(PA_dat1)

print(PA_plot)

my_ggsave("./figs/PA_rand.png", PA_plot)

```

## Observed LL vs predicted LL


```{r, echo = F, message = F}


LL_dat <- PA %>%
  mutate(site_strata = factor(site_strata,
          levels = c("WET_CAN", "DRY_CAN", "WET_UNDER", "DRY_UNDER"))) %>%
  mutate(gr = factor(site_strata,
    labels = c("Sun-Wet",
               "Sun-Dry",
               "Shade-Wet",
               "Shade-Dry"
                      )))

labLL <- data_frame(
                   LL = Inf,
                   Val = min(PA$LL),
                   Val_max = max(PA$LL),
                   gr = "Sun-Dry",
                   r_vals = ""
                   )

LL_plot <- ggplot(LL_dat, aes(x = preLL, y = LL,
                              fill = gr, col = gr)) +
  geom_point(shape = 21) +
  scale_fill_manual(values = fills) +
  scale_colour_manual(values = cols) +
  scale_x_log10(breaks = my_breaks(), expand = c(0.1, 0),
                limits = c(min(labLL$Val), max(labLL$Val_max))) +
  scale_y_log10(breaks = my_breaks(), expand = c(0.1, 0),
                limits = c(min(labLL$Val), max(labLL$Val_max))) +
  xlab("Predicted LL (months)") +
  ylab("Observed LL (months)") +
  geom_abline(aes(slope = 1, intercept = 0),
              lty = 2,
              lwd = 0.25,
              col = "gray20") +
  geom_text(data = labLL, aes(label = r_vals, x = Val_max, y = 2.5),
            colour = "black",
            hjust = 1,
            vjust= 0,
            parse = TRUE,
            show.legend = FALSE) +
  coord_fixed() +
  theme_LES() +
  theme(
    legend.position = c(0.2, 0.85),
    axis.title.x = element_text(margin = margin(t = 1,
                                                b = 1,
                                                l = 0,
                                                r = 0)),
    axis.title.y = element_text(margin = margin(t = 1,
                                                b = 1,
                                                l = 1,
                                                r = 1))
        )

print(LL_plot)

my_ggsave("../figs/LL_plot_rand.png", LL_plot,
          width = 6.7,
          height = 6.7)

cor.test(log(LL_dat$preLL), log(LL_dat$LL))

```

# Model code

```{r}
get_stancode(res) %>% cat
```

# R info

```{r}
sessionInfo()
```
