---
title: Panama Kikuzawa model
author: Masatoshi Katabuchi
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
---

$$
\begin{aligned}
LMAm_i = f_i \times LMA_i \\
LMAs_i = (1-f_i) \times LMA_i
\end{aligned}
$$

$$
\begin{aligned}
A_{area\ i} = \alpha_0 LMAm_{i}^{\alpha_m} LMAs_{i}^{\alpha_s} \\
LL_{i} = \beta_0 LMA_{i}^{0.5} LMAs_{i}^{0.5} (\theta_L A_{area\ i}-R_{area\ i})^{-0.5} \\
R_{area\ i} = \gamma_0 LMAm_{i}^{\gamma_m} LMAs_{i}^{\gamma_s}
\end{aligned}
$$

where, $\theta_L$ is 1 for sun leaves and $0 < \theta_L < 1$ for shade leaves.


```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=TRUE,
                      warning=FALSE,
                      message=FALSE)
```


```{r, echo = F, message = F}
library(loo)
library(rstan)
library(scales)
library(memisc)
library(tidyverse)
library(stringr)
library(lazyeval)
library(cowplot)
library(ggrepel)
library(ggthemes)

```


# Check divergent transitions


```{r, echo = F, message = F}
setwd("..")
getwd()

load("./data/PA_latent_kiku_more_obs.rda")

PA <- dat %>%
  as_data_frame %>%
  mutate(site2 = ifelse(site == "PNM", "DRY", "WET")) %>%
  mutate(sp_site_strata = paste(sp, site2, strata, sep = "_")) %>%
  mutate(site_strata = paste(site2, strata, sep = "_"))

```

There is no divergent transitions (i.e., no fatal error).

```{r}
divergent <- get_sampler_params(res, inc_warmup=FALSE)[[1]][,'divergent__']
sum(divergent)

pairs(res, pars = c("beta[2]", "beta[3]"))
```




# Label problem

Only chain 3 of MCMC shows $\alpha_m > \alpha_s$. So I need to exchange $\alpha_m$ and $\alpha_s$ for the chain 1, 2 and 4.

```{r}

fit <-res
params <- extract(fit)
m <- 4000
Z_i <- params$beta[,1:7]

tmp <- data_frame(iter = 1:m,
           alpha_m = Z_i[,2],
           alpha_s = Z_i[,3]
           ) %>%
  gather(para, val, 2:3)

tmp_text <- data_frame(x = c(0, 1000, 2000, 3000),
                       y = 0.7,
                       lab = c("chain1->", "chain2->", "chain3->", "chain4->"))

ggplot(tmp, aes(x = iter, y = val)) +
  geom_line(aes(col = para)) +
  geom_vline(xintercept = c(0, 1000,2000,3000)) +
  geom_text(data = tmp_text, aes(x = x, y =y , label = lab))
```

Same for $\beta_m$, $\beta_m$, $\gamma_m$, $\gamma_s$ and $f_i$

```{r}

tmp <- data_frame(iter = 1:m,
           r_m = Z_i[,6],
           r_s = Z_i[,7]
           ) %>%
  gather(para, val, 2:3)

ggplot(tmp, aes(x = iter, y = val)) +
  geom_line(aes(col = para)) +
  geom_vline(xintercept = c(0, 1000,2000,3000)) +
  geom_text(data = tmp_text, aes(x = x, y = 0.5 , label = lab))

p_m <- params$p

tmp <- data_frame(iter = 1:m,
           f_1 = p_m[,1],
           para = "f_1"
           )
ggplot(tmp, aes(x = iter, y = f_1)) +
  geom_line(aes(col = para)) +
  geom_vline(xintercept = c(0, 1000,2000,3000)) +
  geom_text(data = tmp_text, aes(x = x, y = 1 , label = lab))

```

Thus, posterior distributions looks bad.

```{r}
summary_PA_LDL <- data.frame(summary(res)$summary)

summary_PA_LDL %>%
  round(3) %>%
  DT::datatable(.)

```


# Main results from here


# Correlation between LMAp and LMAs at each MCMC step

```{r}
p_est<-NULL
mcmc_res <- extract(fit) %>% as.data.frame
params1 <- as.data.frame(extract(fit, permuted=FALSE)[,1,])
params2 <- as.data.frame(extract(fit, permuted=FALSE)[,2,])
params3 <- as.data.frame(extract(fit, permuted=FALSE)[,3,])
params4 <- as.data.frame(extract(fit, permuted=FALSE)[,4,])

LMAp_mcmc <- list()
LMAs_mcmc <- list()
for (i in 1:130) {
  re_p0 <- c(params1[,paste0("p[",i,"]")],
    params2[,paste0("p[",i,"]")],
    params3[,paste0("p[",i,"]")],
    params4[,paste0("p[",i,"]")])
  p_est[i] <- median(re_p0)
  LMAp_mcmc[[i]] <- re_p0 * PA$LMA[i]
  LMAs_mcmc[[i]] <- (1 - re_p0) * PA$LMA[i]
}

names(p_est) <- paste0("p",1:130)

LMAps_cor <- NULL
for (i in 1:4000){
  LMAps_cor[i] <- cor(
  sapply(LMAp_mcmc, "[[", i) %>% log,
  sapply(LMAs_mcmc, "[[", i) %>% log)
}

Mu_mcmc <- extract(fit, pars = "Mu")[[1]]

preLL <- Mu_mcmc[,,2] %>% apply(., 2, median) %>% exp

hist(LMAps_cor)
#mu2_vec <- paste("mu2[", 1:nrow(PA), "]", sep = "")
#Mu_vec <- paste("mu[", 1:nrow(PA), ",2]", sep = "")
temp_LMAp <- p_est * PA$LMA
temp_LMAs <- PA$LMA - temp_LMAp

PA <- PA %>%
  mutate(LMAp = temp_LMAp) %>%
  mutate(LMAs = temp_LMAs)  %>%
  mutate(preLL = preLL)

```


# Trait correlations

## A~area~, R~area~, LL

Open symbols indicate sun leaves and closed symbols indicate shade leaves.

```{r, echo = F, message = F}

setwd("..")
settings <- yaml::yaml.load_file("./settings.yml")
r_vals <- yaml::yaml.load_file("r_val.yml")
p_letters <- yaml::yaml.load_file("letters.yml")
source("fig_theme.r")


PA_dat <- PA %>%
  gather(LMA, Val,
         c(LMA, LMAs, LMAp)) %>%
  gather(Trait, Val2, c(Aarea, Rarea, LL,
                        Parea, Narea, cell_area)) %>%
  mutate(LMA = factor(LMA,
    labels = c("LMA", "LMAm", "LMAs"))) %>%
  mutate(LMA = factor(LMA,
    labels = c("LMA~(~g~m^{-2})", "LMAm~(~g~m^{-2})", "LMAs~(~g~m^{-2})"))) %>%
  mutate(site_strata = factor(site_strata,
          levels = c("WET_CAN", "DRY_CAN", "WET_UNDER", "DRY_UNDER"))) %>%
  mutate(Trait = factor(Trait,
    levels = c("Aarea", "Rarea", "LL", "Narea", "Parea", "cell_area"))) %>%
  mutate(Trait2 = factor(Trait,
    labels = c("italic(A)[area]~(~mu~mol~m^{-2}~s^{-1})",
               "italic(R)[area]~(~mu~mol~m^{-2}~s^{-1})",
               "LL~(months)",
               "italic(N)[area]~(~g~m^{-2})",
               "italic(P)[area]~(~g~m^{-2})",
               "italic(CL)[area]~(~g~m^{-2})"
               ))) %>%
  mutate(gr = factor(site_strata,
    labels = c("Sun-Wet",
               "Sun-Dry",
               "Shade-Wet",
               "Shade-Dry"
                      ))) %>%
  arrange(gr)

# PA LES plot ----------------------------------------------------------------0
PA_dat1 <- PA_dat %>%
  filter(Trait == "LL" |
         Trait == "Aarea" |
         Trait == "Rarea")

lim_PA <- lim_func(PA_dat1 %>% filter(site_strata != "Rand"),
                   trait = "Val")
lim_PA2 <- lim_func(PA_dat1 %>% filter(site_strata != "Rand"),
                    trait = "Val2", LMA = FALSE)

lab1 <- data_frame(lab = paste("(", letters[1:9], ")", sep = ""),
                   Val2 = lim_PA2$max_val %>% rep(each = 3),
                   Val = lim_PA$min_val %>% rep(3),
                   Val_max = lim_PA$max_val %>% rep(3),
                   gr = "Sun-Dry",
                   r_vals = r_vals$r_vals$PA %>% unlist,
                   Trait2 = rep(lim_PA2$Trait2, each = 3),
                   LMA = rep(lim_PA$LMA, 3)
                   )

fills <- c("Sun-Dry" = settings$fills$sun_dry,
          "Sun-Wet" = settings$fills$sun_wet,
          "Shade-Dry" = settings$fills$shade_dry,
          "Shade-Wet" = settings$fills$shade_wet)

cols <- c("Sun-Dry" = settings$colors$sun_dry,
          "Sun-Wet" = settings$colors$sun_wet,
          "Shade-Dry" = settings$colors$shade_dry,
          "Shade-Wet" = settings$colors$shade_wet)

PA_plot <- scatter_plt(PA_dat1)

print(PA_plot)

my_ggsave("./figs/PA_kiku.png", PA_plot)

```

## Observed LL vs predicted LL

Looks OK but each group has less steep slope. Our model tend to under-and over-esimate LL for long- and short-LL leaves, respectively.

```{r, echo = F, message = F}


LL_dat <- PA %>%
  mutate(site_strata = factor(site_strata,
          levels = c("WET_CAN", "DRY_CAN", "WET_UNDER", "DRY_UNDER"))) %>%
  mutate(gr = factor(site_strata,
    labels = c("Sun-Wet",
               "Sun-Dry",
               "Shade-Wet",
               "Shade-Dry"
                      )))

labLL <- data_frame(
                   LL = Inf,
                   Val = min(PA$LL),
                   Val_max = max(PA$LL),
                   gr = "Sun-Dry",
                   r_vals = r_vals$r_vals$PA_LL$preLL_LL
                   )

LL_plot <- ggplot(LL_dat, aes(x = preLL, y = LL,
                              fill = gr, col = gr)) +
  geom_point(shape = 21) +
  scale_fill_manual(values = fills) +
  scale_colour_manual(values = cols) +
  scale_x_log10(breaks = my_breaks(), expand = c(0.1, 0),
                limits = c(min(labLL$Val), max(labLL$Val_max))) +
  scale_y_log10(breaks = my_breaks(), expand = c(0.1, 0),
                limits = c(min(labLL$Val), max(labLL$Val_max))) +
  xlab("Predicted LL (months)") +
  ylab("Observed LL (months)") +
  geom_abline(aes(slope = 1, intercept = 0),
              lty = 2,
              lwd = 0.25,
              col = "gray20") +
  geom_text(data = labLL, aes(label = r_vals, x = Val_max, y = 2.5),
            colour = "black",
            hjust = 1,
            vjust= 0,
            parse = TRUE,
            show.legend = FALSE) +
  coord_fixed() +
  theme_LES() +
  theme(
    legend.position = c(0.2, 0.85),
    axis.title.x = element_text(margin = margin(t = 1,
                                                b = 1,
                                                l = 0,
                                                r = 0)),
    axis.title.y = element_text(margin = margin(t = 1,
                                                b = 1,
                                                l = 1,
                                                r = 1))
        )

print(LL_plot)

my_ggsave("../figs/LL_plot_kiku.png", LL_plot,
          width = 6.7,
          height = 6.7)

cor.test(log(PA$LL), log(PA$preLL))

```


## Cell~area~, Cell~vol~ (= cellulose density), N~area~, P~area~

```{r, echo = F, message = F}


PA_dat2 <- PA_dat %>%
  filter(Trait == "Narea" |
         Trait == "Parea" |
         Trait == "cell_area")


lim_PA <- lim_func(PA_dat2 %>% filter(site_strata != "Rand"),
                   trait = "Val")
lim_PA2 <- lim_func(PA_dat2 %>% filter(site_strata != "Rand"),
                    trait = "Val2", LMA = FALSE)

lab1 <- data_frame(lab = paste("(", letters[1:9], ")", sep = ""),
                   Val2 = lim_PA2$max_val %>% rep(each = 3),
                   Val = lim_PA$min_val %>% rep(3),
                   Val_max = lim_PA$max_val %>% rep(3),
                   gr = "Sun-Dry",
                   r_vals = r_vals$r_vals$PA_NP %>% unlist,
                   Trait2 = rep(lim_PA2$Trait2, each = 3),
                   LMA = rep(lim_PA$LMA, 3)
                   )

PA_NPC_plot <- scatter_plt(PA_dat2)

print(PA_NPC_plot)

my_ggsave("../figs/PA_NPC_kiku.png", PA_NPC_plot)

```

## LMAp vs LMAs

```{r, echo = F, message = F}
ggplot(PA, aes(x = LMAp, y = LMAs, fill = site2, col = site2, shape = strata)) +
  geom_point() +
  xlab("LMAm") +
  ylab("LMAs") +
  scale_shape_manual(values = c("CAN" = 1, "UNDER" = 21)) +
  scale_fill_manual(values = c("DRY" = "#8B7355", "WET" = "#1874CD")) +
  scale_colour_manual(values = c("DRY" = "#8B7355", "WET" = "#1874CD")) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))
```

## Trait means

The results look OK.

```{r, echo = F, message = F, eval = F}

ggplot(PA2 %>% filter(LMA != "LDs"), aes(x = site_strata, y = val,
                                         fill = site_strata,
                                         col = site_strata)) +
  geom_boxplot() +
  facet_grid(.~LMA, scale = "free") +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_fill_manual(values = c("DRY_CAN" = "white",
                               "WET_CAN" = "white",
                               "DRY_UNDER" = "#8B7355",
                               "WET_UNDER" = "#1874CD"
                               )) +
  scale_colour_manual(values = c("DRY_CAN" = "#8B7355",
                               "WET_CAN" = "#1874CD",
                               "DRY_UNDER" = "black",
                               "WET_UNDER" = "black"
                               ))

```


# Decomposition of LMA variation

```{r}

var_p <- function(data)cov(data$LMA, data$LMAp)
var_s <- function(data)cov(data$LMA, data$LMAs)
var_t <- function(data)var(data$LMA)

var_p <- function(data)var(data$LMAp)
var_s <- function(data)var(data$LMAs)
cov_ps <- function(data)cov(data$LMAp, data$LMAs)
var_t <- function(data)var(data$LMA)

var_dat <- PA %>%
  group_by(strata) %>%
  nest %>%
  mutate(var_LMAm = map_dbl(data, var_p)) %>%
  mutate(var_LMAs = map_dbl(data, var_s)) %>%
  mutate(cov_ms = 2 * map_dbl(data, cov_ps)) %>%
  mutate(var_total = map_dbl(data, var_t)) %>%
  dplyr::select(-data)

var_dat_s <- var_dat %>%
  mutate(LMAm_variance = var_LMAm / (var_LMAm + var_LMAs) * 100) %>%
  mutate(LMAs_variance = var_LMAs / (var_LMAm + var_LMAs) * 100) %>%
  gather(var, val, c(LMAm_variance, LMAs_variance))

var_dat_s2 <- var_dat %>%
  mutate(LMAm_variance = var_LMAm / (var_total) * 100) %>%
  mutate(LMAs_variance = var_LMAs / (var_total) * 100) %>%
  mutate(cov_LMAm_LMAs = cov_ms / (var_total) * 100) %>%
  gather(var, val, c(LMAm_variance, LMAs_variance, cov_LMAm_LMAs))

```

1) var(LMAm) & var(LMAs) & 2*cov(LMAm, LMAs), 2)relative contribution: var(LMAm)/(var(LMAm) + var(LMAs)) & var(LMAs) /(var(LMAm) + var(LMAs))

```{r, echo = F, eval = T}

ggplot(var_dat_s2, aes(x = strata, y = val, fill = var)) +
  geom_col() +
  ylab("val (%)")


ggplot(var_dat_s, aes(x = strata, y = val, fill = var)) +
  geom_bar(position = "fill",stat = "identity") +
  scale_y_continuous(labels = percent_format())

```



# Randomized dataset

```{r}

load("../data/PA_latent_kiku_more_rand.rda")
#load("./data/PA_latent_model_light_site_flat_more_rand.rda")

res <- rand_res$model[[1]]
tmp <- rand_res$data[[1]]

PA <- dat %>%
  as_data_frame %>%
  mutate(site2 = ifelse(site == "PNM", "DRY", "WET")) %>%
  mutate(sp_site_strata = paste(sp, site2, strata, sep = "_")) %>%
  mutate(site_strata = paste(site2, strata, sep = "_")) %>%
  mutate(LMA = tmp$LMA) %>%
  mutate(Aarea = tmp$A) %>%
  mutate(LL = tmp$LL) %>%
  mutate(Rarea = tmp$R)


```

## There are divergent transitions (fatal errors)

```{r, eval = T}
pairs(res, pars = c("beta[1]", "beta[2]"))
```


```{r}

traceplot(res, par = "beta")

summary_PA <- data.frame(summary(res)$summary)

summary_PA %>%
  round(3) %>%
  DT::datatable(.)

```

add mean LMAm and LMAs

```{r}

mcmc_res <- extract(fit) %>% as.data.frame

p_est <- mcmc_res[,paste0("p.",  1:130)] %>% apply(.,2,mean)

Mu_mcmc <- extract(fit, pars = "Mu")[[1]]
preLL <- Mu_mcmc[,,2] %>% apply(., 2, mean) %>% exp

temp_LMAp <- p_est * PA$LMA
temp_LMAs <- PA$LMA - temp_LMAp

PA <- PA %>%
  mutate(LMAp = temp_LMAp) %>%
  mutate(LMAs = temp_LMAs)  %>%
  mutate(preLL = preLL)

```

## LMAm and LMAs vs N~area~, P~area~ and LL

```{r, echo = F, message = F}

setwd("..")

PA_dat <- PA %>%
  gather(LMA, Val,
         c(LMA, LMAs, LMAp)) %>%
  gather(Trait, Val2, c(Aarea, Rarea, LL,
                        Parea, Narea, cell_area)) %>%
  mutate(LMA = factor(LMA,
    labels = c("LMA", "LMAm", "LMAs"))) %>%
  mutate(LMA = factor(LMA,
    labels = c("LMA~(~g~m^{-2})", "LMAm~(~g~m^{-2})", "LMAs~(~g~m^{-2})"))) %>%
  mutate(site_strata = factor(site_strata,
          levels = c("WET_CAN", "DRY_CAN", "WET_UNDER", "DRY_UNDER"))) %>%
  mutate(Trait = factor(Trait,
    levels = c("Aarea", "Rarea", "LL", "Narea", "Parea", "cell_area"))) %>%
  mutate(Trait2 = factor(Trait,
    labels = c("italic(A)[area]~(~mu~mol~m^{-2}~s^{-1})",
               "italic(R)[area]~(~mu~mol~m^{-2}~s^{-1})",
               "LL~(months)",
               "italic(N)[area]~(~g~m^{-2})",
               "italic(P)[area]~(~g~m^{-2})",
               "italic(CL)[area]~(~g~m^{-2})"
               ))) %>%
  mutate(gr = factor(site_strata,
    labels = c("Sun-Wet",
               "Sun-Dry",
               "Shade-Wet",
               "Shade-Dry"
                      ))) %>%
  arrange(gr)

# PA LES plot ----------------------------------------------------------------0
PA_dat1 <- PA_dat %>%
  filter(Trait == "LL" |
         Trait == "Aarea" |
         Trait == "Rarea")

lim_PA <- lim_func(PA_dat1 %>% filter(site_strata != "Rand"),
                   trait = "Val")
lim_PA2 <- lim_func(PA_dat1 %>% filter(site_strata != "Rand"),
                    trait = "Val2", LMA = FALSE)

lab1 <- data_frame(lab = paste("(", letters[1:9], ")", sep = ""),
                   Val2 = lim_PA2$max_val %>% rep(each = 3),
                   Val = lim_PA$min_val %>% rep(3),
                   Val_max = lim_PA$max_val %>% rep(3),
                   gr = "Sun-Dry",
                   r_vals = r_vals$r_vals$PA %>% unlist,
                   Trait2 = rep(lim_PA2$Trait2, each = 3),
                   LMA = rep(lim_PA$LMA, 3)
                   )

fills <- c("Sun-Dry" = settings$fills$sun_dry,
          "Sun-Wet" = settings$fills$sun_wet,
          "Shade-Dry" = settings$fills$shade_dry,
          "Shade-Wet" = settings$fills$shade_wet)

cols <- c("Sun-Dry" = settings$colors$sun_dry,
          "Sun-Wet" = settings$colors$sun_wet,
          "Shade-Dry" = settings$colors$shade_dry,
          "Shade-Wet" = settings$colors$shade_wet)

PA_plot <- scatter_plt(PA_dat1)

print(PA_plot)

my_ggsave("./figs/PA_rand_kiku.png", PA_plot)

```

## Observed LL vs predicted LL


```{r, echo = F, message = F}


LL_dat <- PA %>%
  mutate(site_strata = factor(site_strata,
          levels = c("WET_CAN", "DRY_CAN", "WET_UNDER", "DRY_UNDER"))) %>%
  mutate(gr = factor(site_strata,
    labels = c("Sun-Wet",
               "Sun-Dry",
               "Shade-Wet",
               "Shade-Dry"
                      )))

labLL <- data_frame(
                   LL = Inf,
                   Val = min(PA$LL),
                   Val_max = max(PA$LL),
                   gr = "Sun-Dry",
                   r_vals = r_vals$r_vals$PA_LL$preLL_LL
                   )

LL_plot <- ggplot(LL_dat, aes(x = preLL, y = LL,
                              fill = gr, col = gr)) +
  geom_point(shape = 21) +
  scale_fill_manual(values = fills) +
  scale_colour_manual(values = cols) +
  scale_x_log10(breaks = my_breaks(), expand = c(0.1, 0),
                limits = c(min(labLL$Val), max(labLL$Val_max))) +
  scale_y_log10(breaks = my_breaks(), expand = c(0.1, 0),
                limits = c(min(labLL$Val), max(labLL$Val_max))) +
  xlab("Predicted LL (months)") +
  ylab("Observed LL (months)") +
  geom_abline(aes(slope = 1, intercept = 0),
              lty = 2,
              lwd = 0.25,
              col = "gray20") +
  geom_text(data = labLL, aes(label = r_vals, x = Val_max, y = 2.5),
            colour = "black",
            hjust = 1,
            vjust= 0,
            parse = TRUE,
            show.legend = FALSE) +
  coord_fixed() +
  theme_LES() +
  theme(
    legend.position = c(0.2, 0.85),
    axis.title.x = element_text(margin = margin(t = 1,
                                                b = 1,
                                                l = 0,
                                                r = 0)),
    axis.title.y = element_text(margin = margin(t = 1,
                                                b = 1,
                                                l = 1,
                                                r = 1))
        )

print(LL_plot)


my_ggsave("../figs/LL_plot_rand_kiku.png", LL_plot,
          width = 6.7,
          height = 6.7)

```

# Model code  

```{r}
get_stancode(res) %>% cat
```

# R info

```{r}
sessionInfo()
```
