---
title: Panama results
author: Masatoshi Katabuchi
date: "`r format(Sys.time(), '%B %d, %Y')`"
---

Note: posterior median or mean is not good values to visualize our results because $LMAm_i = LMA_i - LMAs_i$ but not sure if $mean(LMAp_i) = LMA_i - mean(LMAs_i)$ is true.

$$
LMAm = f_i * LMA_i \\
LMAs = (1-f_i) * LMA_i
$$

$$
A_{area\ i} = \alpha_0 LMAm_{i}^{\alpha_m} LMAs_{i}^{\alpha_s} \\
LL_{i} = \beta_0 LMAm_{i}^{\beta_m} LMAs_{i}^{\beta_s} exp(\beta_L Light_i) exp(\beta_d Dry_i) \\
R_{area\ i} = \gamma_0 LMAm_{i}^{\gamma_m} LMAs_{i}^{\gamma_s}
$$

where, the dummy variable “light” is 1 for sun leaves and 0 for shade leaves, and the dummy variable "dry" is 1 for dry sites and 0 for wet sites.

$$
Z =
\begin{bmatrix}
\alpha_0 &\beta_0 &\gamma_0 \\
\alpha_m &\beta_m &\gamma_m \\
\alpha_s &\beta_s &\gamma_s \\
0 &\beta_L &0 \\
0 &\beta_d &0 \\
\end{bmatrix}
$$

$$
X =
\begin{bmatrix}
1 &logLMA_i & logLMAs_{i} &Light_i &Dry_i
\end{bmatrix}
$$

$$
Mu = X \times Z
$$

Some parameters are exchangeable. For example, $[f_i, \ \alpha_m, \ \alpha_s]$ = [0.2, 0.5, 1.0] and [0.8, 1.0, 0.5] return the identical expected mean of $A~area~$. When the posterior distributions from different chains were symmetric (as above), I switched labels so that $\alpha_m$ > $\alpha_s$ (LMAm is associated with metabolic rates by our definition).

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=TRUE,
                      warning=FALSE,
                      message=FALSE)
```


```{r, echo = F, message = F}
library(loo)
library(rstan)
library(scales)
library(memisc)
library(tidyverse)
library(stringr)
library(lazyeval)
library(cowplot)
library(ggrepel)
library(ggthemes)

```


# Check divergent transitions


```{r, echo = F, message = F}
setwd("..")
getwd()

load("./data/PA_latent_model_light_site_flat_more_obs.rda")

PA <- dat %>%
  as_data_frame %>%
  mutate(site2 = ifelse(site == "PNM", "DRY", "WET")) %>%
  mutate(sp_site_strata = paste(sp, site2, strata, sep = "_")) %>%
  mutate(site_strata = paste(site2, strata, sep = "_"))

```

There is no divergent transitions (i.e., no fatal error).

```{r}
divergent <- get_sampler_params(res, inc_warmup=FALSE)[[1]][,'divergent__']
sum(divergent)
```




# Label problem

Only chain 3 of MCMC shows $\alpha_m > \alpha_s$. So I need to exchange $\alpha_m$ and $\alpha_s$ for the chain 1, 2 and 4.

```{r}

fit <-res
params <- extract(fit)
m <- 4000
Z_i <- params$Z[,1:5,1:3]

tmp <- data_frame(iter = 1:m,
           alpha_m = Z_i[,2,1],
           alpha_s = Z_i[,3,1]
           ) %>%
  gather(para, val, 2:3)

tmp_text <- data_frame(x = c(0, 1000, 2000, 3000),
                       y = 0.7,
                       lab = c("chain1->", "chain2->", "chain3->", "chain4->"))

ggplot(tmp, aes(x = iter, y = val)) +
  geom_line(aes(col = para)) +
  geom_vline(xintercept = c(0, 1000,2000,3000)) +
  geom_text(data = tmp_text, aes(x = x, y =y , label = lab))
```

Same for $\beta_m$, $\beta_m$, $\gamma_m$, $\gamma_s$ and $f_i$

```{r}

tmp <- data_frame(iter = 1:m,
           beta_m = Z_i[,2,2],
           beta_s = Z_i[,3,2]
           ) %>%
  gather(para, val, 2:3)

ggplot(tmp, aes(x = iter, y = val)) +
  geom_line(aes(col = para)) +
  geom_vline(xintercept = c(0, 1000,2000,3000)) +
  geom_text(data = tmp_text, aes(x = x, y = 0.5 , label = lab))


tmp <- data_frame(iter = 1:m,
           gamma_m = Z_i[,2,3],
           gamma_s = Z_i[,3,3]
           ) %>%
  gather(para, val, 2:3)

ggplot(tmp, aes(x = iter, y = val)) +
  geom_line(aes(col = para)) +
  geom_vline(xintercept = c(0, 1000,2000,3000)) +
  geom_text(data = tmp_text, aes(x = x, y = 1 , label = lab))


p_m <- params$p

tmp <- data_frame(iter = 1:m,
           f_1 = p_m[,1],
           para = "f_1"
           )
ggplot(tmp, aes(x = iter, y = f_1)) +
  geom_line(aes(col = para)) +
  geom_vline(xintercept = c(0, 1000,2000,3000)) +
  geom_text(data = tmp_text, aes(x = x, y = 1 , label = lab))


```

Thus, posterior distributions looks bad.

```{r}
summary_PA_LDL <- data.frame(summary(res)$summary)

summary_PA_LDL %>%
  round(3) %>%
  DT::datatable(.)

```

# Fixing labels manually

$\alpha_m$ <- $\alpha_m$ of chain 3 or $\alpha_s$ of chain 1, 2, and 4.

$\alpha_s$ <- $\alpha_s$ of chain 3 or $\alpha_m$ of chain 1, 2, and 4.


```{r}
mcmc_res <- extract(fit) %>% as.data.frame
params1 <- as.data.frame(extract(fit, permuted=FALSE)[,1,])
params2 <- as.data.frame(extract(fit, permuted=FALSE)[,2,])
params3 <- as.data.frame(extract(fit, permuted=FALSE)[,3,])
params4 <- as.data.frame(extract(fit, permuted=FALSE)[,4,])

#get_stancode(fit) %>% cat

re_am <- c(params1$`Z[3,1]`,
	 			 params2$`Z[3,1]`,
	 			 params3$`Z[2,1]`,
	 			 params4$`Z[3,1]`)

re_as <- c(params1$`Z[2,1]`,
	 			 params2$`Z[2,1]`,
	 			 params3$`Z[3,1]`,
	 			 params4$`Z[2,1]`)

re_bm <- c(params1$`Z[3,2]`,
	 			 params2$`Z[3,2]`,
         params3$`Z[2,2]`,
         params4$`Z[3,2]`)

re_bs <- c(params1$`Z[2,2]`,
           params2$`Z[2,2]`,
           params3$`Z[3,2]`,
           params4$`Z[2,2]`)

re_rs <- c(params1$`Z[2,3]`,
	 			 params2$`Z[2,3]`,
         params3$`Z[3,3]`,
         params4$`Z[2,3]`)

re_rm <- c(params1$`Z[3,3]`,
           params2$`Z[3,3]`,
           params3$`Z[2,3]`,
           params4$`Z[3,3]`)

re_p1 <- c(1-params1$`p[1]` ,
          1-params2$`p[1]` ,
          params3$`p[1]` ,
          1-params4$`p[1]` )

re_p0 <- c(1 - params1[,paste0("p[",1:130,"]")],
  1 - params2[,paste0("p[",1:130,"]")],
  params3[,paste0("p[",1:130,"]")],
  1 - params4[,paste0("p[",1:130,"]")]) %>% unlist

re_p <- array(NA,
              dim = c(4000, 130))

mpc <- fit@sim$iter-fit@sim$warmup

tmp <- array(c(re_am,
               re_as,
               re_bm,
               re_bs,
               re_rm,
               re_rs
                  ),
             dim = c(mpc * 4, 6))

tmp2 <- tmp

#par(mfrow=c(5,5))
for (i in 1:130) {
  re_p0 <- c(1-params1[,paste0("p[",i,"]")],
    1-params2[,paste0("p[",i,"]")],
    params3[,paste0("p[",i,"]")],
    1 - params4[,paste0("p[",i,"]")])
 # hist(re_p0)
  tmp2 <- cbind(tmp2, re_p0)
}
#par(mfrow=c(1,1))

amu <- array(NA,dim = c(mpc,fit@sim$chains,dim(tmp2)[2]))
for (cc in 1:fit@sim$chains) amu[,cc,] = tmp2[(1:mpc)+(cc-1)*mpc,]
mmu <- monitor(amu, print = F, warmup = 0)
rownames(mmu) <- c("am", "as", "bm", "bs", "rm", "rs",
                   paste0("p",1:130))


mcmc_re <- data_frame(alpha_m = amu[,,1] %>% as.vector,
             alpha_s = amu[,,2] %>% as.vector,
             beta_m = amu[,,3] %>% as.vector,
             beta_s = amu[,,4] %>% as.vector,
             gamma_m = amu[,,5] %>% as.vector,
             gamma_s = amu[,,6] %>% as.vector
             ) %>%
  mutate(iter = rep(1:1000, 4)) %>%
  mutate(chain = rep(1:4, each = 1000) %>% as.factor)

mcmc_re2 <- mcmc_re %>%
  gather(para, val, 1:6)

```

New parameters look good.

```{r}
mmu %>% DT::datatable(.)

ggplot(mcmc_re2, aes(x = iter, y = val, col = chain)) +
  geom_line() +
  facet_wrap(~para, nrow = 3)
```

# Main results from here


# Correlation between LMAp and LMAs at each MCMC step

```{r}
p_est<-NULL

LMAp_mcmc <- list()
LMAs_mcmc <- list()
for (i in 1:130) {
  re_p0 <- c(1-params1[,paste0("p[",i,"]")],
    1-params2[,paste0("p[",i,"]")],
    params3[,paste0("p[",i,"]")],
    1 - params4[,paste0("p[",i,"]")])
  p_est[i] <- median(re_p0)
  LMAp_mcmc[[i]] <- re_p0 * PA$LMA[i]
  LMAs_mcmc[[i]] <- (1 - re_p0) * PA$LMA[i]
}

names(p_est) <- paste0("p",1:130)

p_est <- params3[500,paste0("p[",1:130,"]")] %>% unlist


p_est <- params3[500,paste0("p[",1:130,"]")] %>% unlist

p_est2 <- params3[,paste0("p[",1:130,"]")] %>% apply(.,2, mean)
p_est3 <- params3[,paste0("p[",1:130,"]")] %>% apply(.,2, median)

plot(p_est2, p_est3, ylab = "p median", xlab = "p mean")
abline(a=0,b=1)

temp <- data_frame(p_est2, p_est3) %>%
  gather(p, val, 1:2)
ggplot(temp, aes(val, col = p)) +
  geom_density()

LMAps_cor <- NULL
for (i in 1:4000){
  LMAps_cor[i] <- cor(
  sapply(LMAp_mcmc, "[[", i) %>% log,
  sapply(LMAs_mcmc, "[[", i) %>% log)
}

Mu_mcmc <- extract(fit, pars = "Mu")[[1]]

preLL <- Mu_mcmc[,,2] %>% apply(., 2, median) %>% exp
preLL1 <- Mu_mcmc[500,,2] %>% exp

hist(LMAps_cor)
#mu2_vec <- paste("mu2[", 1:nrow(PA), "]", sep = "")
#Mu_vec <- paste("mu[", 1:nrow(PA), ",2]", sep = "")
temp_LMAp <- p_est * PA$LMA
temp_LMAs <- PA$LMA - temp_LMAp

PA <- PA %>%
  mutate(LMAp = temp_LMAp) %>%
  mutate(LMAs = temp_LMAs)  %>%
  mutate(preLL = preLL) %>%
  mutate(preLL1 = preLL1)

```


# Trait correlations

## A~area~, R~area~, LL

Open symbols indicate sun leaves and closed symbols indicate shade leaves.

```{r, echo = F, message = F}

setwd("..")
settings <- yaml::yaml.load_file("./settings.yml")
r_vals <- yaml::yaml.load_file("r_val.yml")
p_letters <- yaml::yaml.load_file("letters.yml")
source("fig_theme.r")


PA_dat <- PA %>%
  gather(LMA, Val,
         c(LMA, LMAs, LMAp)) %>%
  gather(Trait, Val2, c(Aarea, Rarea, LL,
                        Parea, Narea, cell_area)) %>%
  mutate(LMA = factor(LMA,
    labels = c("LMA", "LMAm", "LMAs"))) %>%
  mutate(LMA = factor(LMA,
    labels = c("LMA~(~g~m^{-2})", "LMAm~(~g~m^{-2})", "LMAs~(~g~m^{-2})"))) %>%
  mutate(site_strata = factor(site_strata,
          levels = c("WET_CAN", "DRY_CAN", "WET_UNDER", "DRY_UNDER"))) %>%
  mutate(Trait = factor(Trait,
    levels = c("Aarea", "Rarea", "LL", "Narea", "Parea", "cell_area"))) %>%
  mutate(Trait2 = factor(Trait,
    labels = c("italic(A)[area]~(~mu~mol~m^{-2}~s^{-1})",
               "italic(R)[area]~(~mu~mol~m^{-2}~s^{-1})",
               "LL~(months)",
               "italic(N)[area]~(~g~m^{-2})",
               "italic(P)[area]~(~g~m^{-2})",
               "italic(CL)[area]~(~g~m^{-2})"
               ))) %>%
  mutate(gr = factor(site_strata,
    labels = c("Sun-Wet",
               "Sun-Dry",
               "Shade-Wet",
               "Shade-Dry"
                      ))) %>%
  arrange(gr)

# PA LES plot ----------------------------------------------------------------0
PA_dat1 <- PA_dat %>%
  filter(Trait == "LL" |
         Trait == "Aarea" |
         Trait == "Rarea")

lim_PA <- lim_func(PA_dat1 %>% filter(site_strata != "Rand"),
                   trait = "Val")
lim_PA2 <- lim_func(PA_dat1 %>% filter(site_strata != "Rand"),
                    trait = "Val2", LMA = FALSE)

lab1 <- data_frame(lab = paste("(", letters[1:9], ")", sep = ""),
                   Val2 = lim_PA2$max_val %>% rep(each = 3),
                   Val = lim_PA$min_val %>% rep(3),
                   Val_max = lim_PA$max_val %>% rep(3),
                   gr = "Sun-Dry",
                   r_vals = r_vals$r_vals$PA %>% unlist,
                   Trait2 = rep(lim_PA2$Trait2, each = 3),
                   LMA = rep(lim_PA$LMA, 3)
                   )

fills <- c("Sun-Dry" = settings$fills$sun_dry,
          "Sun-Wet" = settings$fills$sun_wet,
          "Shade-Dry" = settings$fills$shade_dry,
          "Shade-Wet" = settings$fills$shade_wet)

cols <- c("Sun-Dry" = settings$colors$sun_dry,
          "Sun-Wet" = settings$colors$sun_wet,
          "Shade-Dry" = settings$colors$shade_dry,
          "Shade-Wet" = settings$colors$shade_wet)

PA_plot <- scatter_plt(PA_dat1)

print(PA_plot)

my_ggsave("./figs/PA.png", PA_plot)

```

## Observed LL vs predicted LL

Looks OK but each group has less steep slope. Our model tend to under-and over-esimate LL for long- and short-LL leaves, respectively.

```{r, echo = F, message = F}


LL_dat <- PA %>%
  mutate(site_strata = factor(site_strata,
          levels = c("WET_CAN", "DRY_CAN", "WET_UNDER", "DRY_UNDER"))) %>%
  mutate(gr = factor(site_strata,
    labels = c("Sun-Wet",
               "Sun-Dry",
               "Shade-Wet",
               "Shade-Dry"
                      )))

labLL <- data_frame(
                   LL = Inf,
                   Val = min(PA$LL),
                   Val_max = max(PA$LL),
                   gr = "Sun-Dry",
                   r_vals = r_vals$r_vals$PA_LL$preLL_LL
                   )

LL_plot <- ggplot(LL_dat, aes(x = preLL, y = LL,
                              fill = gr, col = gr)) +
  geom_point(shape = 21) +
  scale_fill_manual(values = fills) +
  scale_colour_manual(values = cols) +
  scale_x_log10(breaks = my_breaks(), expand = c(0.1, 0),
                limits = c(min(labLL$Val), max(labLL$Val_max))) +
  scale_y_log10(breaks = my_breaks(), expand = c(0.1, 0),
                limits = c(min(labLL$Val), max(labLL$Val_max))) +
  xlab("Predicted LL (months)") +
  ylab("Observed LL (months)") +
  geom_abline(aes(slope = 1, intercept = 0),
              lty = 2,
              lwd = 0.25,
              col = "gray20") +
  geom_text(data = labLL, aes(label = r_vals, x = Val_max, y = 2.5),
            colour = "black",
            hjust = 1,
            vjust= 0,
            parse = TRUE,
            show.legend = FALSE) +
  coord_fixed() +
  theme_LES() +
  theme(
    legend.position = c(0.2, 0.85),
    axis.title.x = element_text(margin = margin(t = 1,
                                                b = 1,
                                                l = 0,
                                                r = 0)),
    axis.title.y = element_text(margin = margin(t = 1,
                                                b = 1,
                                                l = 1,
                                                r = 1))
        )

print(LL_plot)

LL_plot2 <- ggplot(LL_dat, aes(x = preLL1, y = LL,
                              fill = gr, col = gr)) +
  geom_point(shape = 21) +
  scale_fill_manual(values = fills) +
  scale_colour_manual(values = cols) +
  scale_x_log10(breaks = my_breaks(), expand = c(0.1, 0),
                limits = c(min(labLL$Val), max(labLL$Val_max))) +
  scale_y_log10(breaks = my_breaks(), expand = c(0.1, 0),
                limits = c(min(labLL$Val), max(labLL$Val_max))) +
  xlab("Predicted LL (months)") +
  ylab("Observed LL (months)") +
  geom_abline(aes(slope = 1, intercept = 0),
              lty = 2,
              lwd = 0.25,
              col = "gray20") +
  geom_text(data = labLL, aes(label = r_vals, x = Val_max, y = 2.5),
            colour = "black",
            hjust = 1,
            vjust= 0,
            parse = TRUE,
            show.legend = FALSE) +
  coord_fixed() +
  theme_LES() +
  theme(
    legend.position = c(0.2, 0.85),
    axis.title.x = element_text(margin = margin(t = 1,
                                                b = 1,
                                                l = 0,
                                                r = 0)),
    axis.title.y = element_text(margin = margin(t = 1,
                                                b = 1,
                                                l = 1,
                                                r = 1))
        )

print(LL_plot2)

my_ggsave("../figs/LL_plot.png", LL_plot,
          width = 6.7,
          height = 6.7)

cor.test(log(PA$LL), log(PA$preLL1))

```


## Cell~area~, Cell~vol~ (= cellulose density), N~area~, P~area~

```{r, echo = F, message = F}


PA_dat2 <- PA_dat %>%
  filter(Trait == "Narea" |
         Trait == "Parea" |
         Trait == "cell_area")


lim_PA <- lim_func(PA_dat2 %>% filter(site_strata != "Rand"),
                   trait = "Val")
lim_PA2 <- lim_func(PA_dat2 %>% filter(site_strata != "Rand"),
                    trait = "Val2", LMA = FALSE)

lab1 <- data_frame(lab = paste("(", letters[1:9], ")", sep = ""),
                   Val2 = lim_PA2$max_val %>% rep(each = 3),
                   Val = lim_PA$min_val %>% rep(3),
                   Val_max = lim_PA$max_val %>% rep(3),
                   gr = "Sun-Dry",
                   r_vals = r_vals$r_vals$PA_NP %>% unlist,
                   Trait2 = rep(lim_PA2$Trait2, each = 3),
                   LMA = rep(lim_PA$LMA, 3)
                   )

PA_NPC_plot <- scatter_plt(PA_dat2)

print(PA_NPC_plot)

my_ggsave("../figs/PA_NPC.png", PA_NPC_plot)

```

## LMAp vs LMAs

```{r, echo = F, message = F}
ggplot(PA, aes(x = LMAp, y = LMAs, fill = site2, col = site2, shape = strata)) +
  geom_point() +
  xlab("LMAm") +
  ylab("LMAs") +
  scale_shape_manual(values = c("CAN" = 1, "UNDER" = 21)) +
  scale_fill_manual(values = c("DRY" = "#8B7355", "WET" = "#1874CD")) +
  scale_colour_manual(values = c("DRY" = "#8B7355", "WET" = "#1874CD")) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))
```

## Trait means

The results look OK.

```{r, echo = F, message = F, eval =F}

ggplot(PA2 %>% filter(LMA != "LDs"), aes(x = site_strata, y = val,
                                         fill = site_strata,
                                         col = site_strata)) +
  geom_boxplot() +
  facet_grid(.~LMA, scale = "free") +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_fill_manual(values = c("DRY_CAN" = "white",
                               "WET_CAN" = "white",
                               "DRY_UNDER" = "#8B7355",
                               "WET_UNDER" = "#1874CD"
                               )) +
  scale_colour_manual(values = c("DRY_CAN" = "#8B7355",
                               "WET_CAN" = "#1874CD",
                               "DRY_UNDER" = "black",
                               "WET_UNDER" = "black"
                               ))

```


# Decomposition of LMA variation

```{r}

var_p <- function(data)cov(data$LMA, data$LMAp)
var_s <- function(data)cov(data$LMA, data$LMAs)
var_t <- function(data)var(data$LMA)

var_p <- function(data)var(data$LMAp)
var_s <- function(data)var(data$LMAs)
cov_ps <- function(data)cov(data$LMAp, data$LMAs)
var_t <- function(data)var(data$LMA)

var_dat <- PA %>%
  group_by(strata) %>%
  nest %>%
  mutate(var_LMAm = map_dbl(data, var_p)) %>%
  mutate(var_LMAs = map_dbl(data, var_s)) %>%
  mutate(cov_ms = 2 * map_dbl(data, cov_ps)) %>%
  mutate(var_total = map_dbl(data, var_t)) %>%
  dplyr::select(-data)

var_dat_s <- var_dat %>%
  mutate(LMAm_variance = var_LMAm / (var_LMAm + var_LMAs) * 100) %>%
  mutate(LMAs_variance = var_LMAs / (var_LMAm + var_LMAs) * 100) %>%
  gather(var, val, c(LMAm_variance, LMAs_variance))

var_dat_s2 <- var_dat %>%
  mutate(LMAm_variance = var_LMAm / (var_total) * 100) %>%
  mutate(LMAs_variance = var_LMAs / (var_total) * 100) %>%
  mutate(cov_LMAm_LMAs = cov_ms / (var_total) * 100) %>%
  gather(var, val, c(LMAm_variance, LMAs_variance, cov_LMAm_LMAs))

```

1) var(LMAm) & var(LMAs) & 2*cov(LMAm, LMAs), 2)relative contribution: var(LMAm)/(var(LMAm) + var(LMAs)) & var(LMAs) /(var(LMAm) + var(LMAs))

```{r, echo = F, eval = T}

ggplot(var_dat_s2, aes(x = strata, y = val, fill = var)) +
  geom_col() +
  ylab("val (%)")


ggplot(var_dat_s, aes(x = strata, y = val, fill = var)) +
  geom_bar(position = "fill",stat = "identity") +
  scale_y_continuous(labels = percent_format())

```
