`r page_break("")`

# Figs

```{r GL-rand, echo = F, dpi = 300, fig.width = 6, fig.height = 6, fig.cap = "Leaf-trait relationships in the randomized GLOPNET dataset with no functional correlations involving LMA, *A*~area~, *R*~area~ and LL. Symbols show 1 of 9 randomized datasets, in which all leaf traits were randomized among all leaves in the GLOPNET dataset we analyzed."}

img <- readPNG("../figs/GL_rand2.png", native = TRUE, info = TRUE)
grid.raster(img, interpolate=F)

```

`r page_break("")`

**Just delete or replace by a more meaningful figure** Because the model didn't convergent at all.

```{r LL-rand, echo = F, dpi = 300, fig.width = 6, fig.height = 6, fig.cap = "Leaf-trait relationships in the randomized Panama dataset. Symbols show 1 of 9 randomized datasets, in which all leaf traits were randomized among all leaves in the Panama dataset we analyzed."}

img <- readPNG("../figs/PA_rand.png", native = TRUE, info = TRUE)
grid.raster(img, interpolate=F)

```

`r page_break("")`

**Just delete or replace by a more meaningful figure** Because the model didn't convergent at all.

```{r PA-rand, echo = F, dpi = 300, fig.width = 6, fig.height = 6, fig.cap = "Observed vs. predicted leaf lifespan (LL) in the Optimal LL Model (Eq. 8). Predicted LL values are posterior means. The dashed line indicates the 1:1 relationship. Symbols show 1 of 10 randomized datasets, in which all leaf traits were randomized among all leaves in the Panama dataset we analyzed."}

img <- readPNG("../figs/LL_rand.png", native = TRUE, info = TRUE)
grid.raster(img, interpolate=F)

```


```{r GL-NP, echo = F, dpi = 300, fig.width = 6, fig.height = 6, fig.cap = "Measured traits related to photosynthesis and metabolism traits (nitrogen and phosphorus per-unit leaf area; *N*~area~ and *P*~area~) are strongly correlated with estimates (posterior medians) of the metabolic LMA component (LMAm) in the GLOPNET dataset. LMAm yields slightly stronger correlations and more consistent relationships compared to LMA and LMAs; e.g., evergreen and deciduous leaves align along a single relationship in  panel b, but not in panels a or c."}

img <- readPNG("../figs/GL_NP.png", native = TRUE, info = TRUE)
grid.raster(img, interpolate=F)

```



`r page_break()`

```{r LMAms, echo = F, dpi = 300, fig.width = 6, fig.height = 6, fig.cap = "Posterior medians of LMAp vs LMAs in the (a) GLOPNET and (b) Panama datasets. The *r* values were calucalted for each MCMC step and the medians [95% credible interval] are shown. Correlations between LMAm and LMAs are significantly negative for Panama, but the small absoulte *r* value indicate that a single axis could not accurately represent the two-dimensional space. Symbols as in Main Text Figs. 2-3."}

img <- readPNG("../figs/LMAms.png", native = TRUE, info = TRUE)
grid.raster(img, interpolate=F)
```


`r page_break()`

```{r box-frac, echo = F, dpi = 300, fig.width = 6, fig.height = 6, fig.cap = "Boxplots comparing posterior medians of the latent variable f (the fraction of total LMA comprised by LMAm) across deciduous (D) and evergreen (E) leaves in GLOPNET and Panama, and across sites (wet and dry) and canopy strata (sun and shade) in Panama. Note that LMAm = f × LMA, and LMAs = (1 – f) × LMA. (a) Deciduous and evergreen leaves in the GLOPNET dataset; (b) deciduous and evergreen leaves for Panama species for which both sun and shade leaves were available; (c) leaves for Panama species for which both sun and shade leaves were available; and (d) all leaves for Panama. At the dry Panama site, the increase in LMA from shade to sun (Fig. 4) is due to roughly equal proportional increases in LMAm and LMAs (because f is similar between sun and shade), whereas at the wet Panama site, the increase in LMA from shade to sun (Fig. 4) is due primarily to increased LMAm (because f is greater in sun than shade). GLOPNET results are for the Potential LL Model (Eq. 5), and Panama results are for the Optimal LL Model with site effects (Eq. 13 and Supporting Information, Section 3.1). The center line in each box indicates the median, upper and lower box edges indicate the interquartile range, whiskers show 1.5 times the interquartile range, and points are outliers.Groups sharing the same letters are not significantly different (P > 0.05; t-tests)."}

img <- readPNG("../figs/box_frac2.png", native = TRUE, info = TRUE)
grid.raster(img, interpolate=F)
```

`r page_break()`

```{r boxplt, echo = F, dpi = 300, fig.width = 6, fig.height = 6, fig.cap = "Boxplots comparing leaf mass per area (LMA) and posterior medians of photosynthetic and structural LMA components (LMAm and LMAs, respectively) across sites (wet and dry) and canopy strata (sun and shade) in the Panama dataset). The results shown here include all leaves in the Panama dataset, whereas Fig. 6 in the main text only includes Panema species for which both sun and shade leaves were available."}

img <- readPNG("../figs/box_SI.png", native = TRUE, info = TRUE)
grid.raster(img, interpolate=F)
```

`r page_break()`

**Keepng this figure for now for understanding patterns**

```{r frac-pair, echo = F, dpi = 300, fig.width = 6, fig.height = 6, fig.cap = "Posterior medians of the latent variable f (the fraction of total LMA comprised by LMAm) of leaves for Panama species for which both sun and shade leaves were available. Same species are connected"}


img <- readPNG("../figs/frac_paired.png", native = TRUE, info = TRUE)
grid.raster(img, interpolate=F)
```

`r page_break()`

**Keepng this figure for now for understanding patterns**

```{r LMAm-pair, echo = F, dpi = 300, fig.width = 6, fig.height = 6, fig.cap = "Posterior medians of LMAm of leaves for Panama species for which both sun and shade leaves were available. Same species are connected"}

img <- readPNG("../figs/LMAm_paired.png", native = TRUE, info = TRUE)
grid.raster(img, interpolate=F)
```

# References
