---
title: "MVN for LMAp and LMAs"
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=TRUE,
                      warning=FALSE,
                      message=FALSE,
                   eval = T)
```

```{r caching, include=FALSE}
library(methods)
library(knitr)
library(kableExtra)
```

Consider a multivariate multiple regression model:

$$
\begin{eqnarray}
\left[
\begin{array}{ccc}
y_{11} & y_{12} & y_{13} \\
\vdots & \ddots & \vdots \\
y_{n1} & y_{n2} & y_{n3} \\
\end{array}
\right]
=
\left[
\begin{array}{ccc}
x_{11} & x_{12} & x_{13} \\
\vdots & \ddots & \vdots \\
x_{n1} & x_{n2} & x_{n3} \\
\end{array}
\right]
\left[
\begin{array}{ccc}
\beta_{11} & \beta_{12} & \beta_{13} \\
\beta_{21} & \beta_{22} & \beta_{23} \\
\beta_{31} & \beta_{32} & \beta_{33} \\
\end{array}
\right]
+
\left[
\begin{array}{ccc}
\epsilon_{11} & \epsilon_{12} & \epsilon_{13} \\
\vdots & \ddots & \vdots \\
\epsilon_{n1} & \epsilon_{n2} & \epsilon_{n3} \\
\end{array}
\right]
\end{eqnarray}
$$


Where $y_{i1}$ is Aarea,  $y_{i2}$ is LL, $y_{i3}$ is Rarea, $x_{i1}$ is an intercept, $x_{i2}$ is LMAp, and $x_{i3}$ is LMAs. In our model, however, $y_{i1}$ is sum of Aarea and Rarea. Thus, the above eq. can be rewritten as:

$$
\begin{eqnarray}
\left[
\begin{array}{ccc}
y_{11} + y_{13} & y_{12} & y_{13} \\
\vdots & \ddots & \vdots \\
y_{n1} + y_{n3} & y_{n2} & y_{n3} \\
\end{array}
\right]
=
\left[
\begin{array}{ccc}
x_{11} & x_{12} & x_{13} \\
\vdots & \ddots & \vdots \\
x_{n1} & x_{n2} & x_{n3} \\
\end{array}
\right]
\left[
\begin{array}{ccc}
\beta_{11} + \beta_{13} & \beta_{12} & \beta_{13} \\
\beta_{21} + \beta_{23} & \beta_{22} & \beta_{23} \\
\beta_{31} + \beta_{33} & \beta_{32} & \beta_{33} \\
\end{array}
\right]
+
\left[
\begin{array}{ccc}
\epsilon_{11} & \epsilon_{12} & \epsilon_{13} \\
\vdots & \ddots & \vdots \\
\epsilon_{n1} & \epsilon_{n2} & \epsilon_{n3} \\
\end{array}
\right]
\end{eqnarray}
$$

but what we are assuming is:

$$
\begin{eqnarray}
\left[
\begin{array}{ccc}
y_{11} + y_{12} & y_{12} & y_{13} \\
\vdots & \ddots & \vdots \\
y_{n1} + y_{n3} & y_{n2} & y_{n3} \\
\end{array}
\right]
=
\left[
\begin{array}{ccc}
x_{11} & x_{12} & x_{13} \\
\vdots & \ddots & \vdots \\
x_{n1} & x_{n2} & x_{n3} \\
\end{array}
\right]
\left[
\begin{array}{ccc}
\beta_{11} + \beta_{13} & \beta_{12} & \beta_{13} \\
\beta_{21} + \beta_{23} & 0 & \beta_{23} \\
0 & \beta_{32} & \beta_{33} \\
\end{array}
\right]
+
\left[
\begin{array}{ccc}
\epsilon_{11} & \epsilon_{12} & \epsilon_{13} \\
\vdots & \ddots & \vdots \\
\epsilon_{n1} & \epsilon_{n2} & \epsilon_{n3} \\
\end{array}
\right]
\end{eqnarray}
$$

In our definition, LMAs does not play role for Aarea, thus $\beta_{22} = 0$ is fine but the above model further produces an assumption of $\beta_{31} = -\beta_{33}$. Thus, negative covariance between LMAs and net and/or gross photosynthetic rates will produce positive covariance between LMAs and Rdark and vice versa.


**What should we do?**

1. We can assume that $y_{i1}$ is Aarea instead of Aarea + Rarea. Aarea is gross Amax - daily respiration which could still contain information for Rdark.
1. We can use three univariate regressions instead of using a single multivariate regression to get rid of the assumption of $\beta_{31} = -\beta_{33}$. As far as I ran some simulation, a univariate normal distribution has better performance to estimate parameters of interests.
    1. As long as we use a univariate normal distribution, Aarea or Aarea + Rarea seems not a big deal.
