---
fontsize: 12pt
geometry: margin=1in
link-citations: true
csl: ecology.csl
bibliography: [LMA.bib]
crossref:
  fig-title: Fig.
  fig-prefix: Fig.
  eq-prefix: Equation
format:
  pdf:
    toc: false
    keep-tex: true
    latex-clean: false
    pdf-engine: xelatex
    number-sections: true  # Enable section numbering
    include-in-header:
      text: |
        \usepackage[default]{sourcesanspro}
        \usepackage{sourcecodepro}
        \usepackage{lineno}
        \linenumbers
        \linespread{1.2}
  html:
    theme: spacelab #readable #sandstone #spacelab #flatly
    toc: true
    toc-depth: 2
    toc-title: Contents
    self-contained: true
    smooth-scroll: true
    highlight-style: github
  docx:
    toc: false
    number-sections: true
    highlight-style: github
    html-math-method: katex
    reference-doc: my_template.docx
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(knitr)
library(kableExtra)
library(flextable)
library(here)
# here::i_am("ms/LMAms_main.qmd")
source(here("R/vpart.R"))
source(here("R/render.R"))
var_val <- yaml::yaml.load_file(here("yml/var.yml"))
lma_val <- yaml::yaml.load_file(here("yml/lma.yml"))
para_val <- yaml::yaml.load_file(here("yml/para.yml"))
```

```{r, include=FALSE}
# Make sure the order is correct
s_fig("box_inter")
s_fig("box_frac_de")
s_fig("box_frac_pa")
s_fig("mass_prop_sim")
s_fig("mass_prop_comp")
s_fig("gl_point_np2")
s_fig("pa_point_npc_par")
s_fig("LMAm_LMAs")

# because \stepcounter doesn't work for docx
eq("LMA")
eq("Aarea")
eq("Rarea")
eq("LL")
eq("mass")
eq("var")

appendix("prior")
appendix("model-fit")
appendix("rand")
appendix("mass-prop")
appendix("mass-prop-sim")
appendix("vpart")
appendix("stan")

```

**Decomposing leaf mass into metabolic and structural components explains divergent patterns of trait variation within and among plant species**
\
\
Masatoshi Katabuchi^1,2\*^, Kaoru Kitajima^2,3,4^, S. Joseph Wright^4^, Sunshine A. Van Bael^4,5^, Jeanne L. D. Osnas^2^ and Jeremy W. Lichstein^2^
\
\
^1^ Xishuangbanna Tropical Botanical Garden, Chinese Academy of Sciences, Menglun, Yunnan 666303 China

^2^ Department of Biology, University of Florida, Gainesville, FL 32611, USA

^3^ Graduate School of Agriculture, Kyoto University, Kitashirakawa Oiwake-Cho, Kyoto 606-8502 Japan

^4^ Smithsonian Tropical Research Institute, 9100 Panama City Pl., Washington, DC 20521

^5^ Department of Ecology and Evolutionary Biology, Tulane University, New Orleans, LA 70118 USA

**\*Corresponding Author**: E-mail: <katabuchi@xtbg.ac.cn>; <mattocci27@gmail.com>

Masatoshi Katabuchi:
<https://orcid.org/0000-0001-9900-9029>

Kaoru Kitajima:
<https://orcid.org/0000-0001-6822-8536>

Sunshine A. Van Bael:
<https://orcid.org/0000-0001-7317-3533>

S. Joseph Wright:
<https://orcid.org/0000-0003-4260-5676>

Jeremy W. Lichstein:
<https://orcid.org/0000-0001-5553-6142>
\
\
**Keywords:**
Bayesian inference,
functional diversity,
global vegetation models,
interspecific variation,
intraspecific variability,
latent variables,
optimal leaf lifespan

**Abstract**

Across the global flora, interspecific variation in photosynthetic and metabolic rates depends more strongly on leaf area than leaf mass.
In contrast, intraspecific variation in these rates is strongly mass-dependent.
These contrasting patterns suggest that the causes of variation in leaf mass per area (LMA) may be fundamentally different within vs. among species.
In order to explain these contrasting patterns, We developed a statistical modeling framework to decompose LMA into two conceptual components – metabolic LMAm (which determines photosynthetic capacity and dark respiration) and structural LMAs (which determines leaf toughness and potential leaf lifespan) - using leaf trait data from tropical forests in Panama and a global leaf-trait database.
Decomposing LMA into LMAm and LMAs improves predictions of leaf trait variation (photosynthesis, respiration, and lifespan).
We show that strong area-dependence of metabolic traits across species can result from multiple factors, including high LMAs variance and/or a slow increase in photosynthetic capacity with increasing LMAm.
In contrast, strong mass-dependence of metabolic traits within species results from LMAm increasing from shady to sunny conditions.
LMAm and LMAs were nearly independent of each other in both global and Panama datasets.
Our results explain different patterns in variation in LMA between across- versus within-species.
Additionally, our analysis highlights that leaf functional variation is multi-dimensional, suggesting that biogeochemical models should treat metabolic and structural leaf components separately.

# Introduction

Leaf functional traits play an important role in ecological and physiological tradeoffs [@Wright2004a; @Reich2014; @Onoda2017] and in carbon and nutrient cycling [e.g., @Huntingford2017; @Tcherkez2017].
Thus, understanding the causes and consequences of leaf trait variation is an important goal in ecology, plant physiology, and biogeochemistry [@Bonan2002; @Poorter2009].
Different leaf assemblages exhibit markedly different patterns of trait variation.
For example, across global species, whole-leaf values of traits related to photosynthesis and metabolism (e.g., the photosynthetic capacity, respiration rate, or nutrient content of entire leaves) tend to increase more strongly with leaf area than leaf mass [@Osnas2013],
whereas the opposite pattern (strong mass-dependence of these same traits) is observed within species across light gradients [@Osnas2018; @Osada2001].
Functional groups (e.g., deciduous vs. evergreen angiosperms) also differ from each other in terms of how photosynthetic and metabolic traits variation depend on leaf mass and area [@Lusk2008; @Osnas2018].
These divergent patterns suggest the presence of multiple drivers of trait variation.

Strong interspecific correlations among leaf mass per area (LMA), leaf lifespan (LL), and mass-normalized leaf traits related to photosynthesis and metabolism (hereafter, 'metabolic traits') have been interpreted as evidence for a single dominant axis of leaf functional variation [@Wright2004a].
However, the interpretation of these strong correlations, which can result from mass-normalization of area-dependent traits [@Lloyd2013; @Osnas2013], is controversial.
On the one hand, mass-normalization has been justified based on economic principles alone [@Westoby2013], because leaf mass provides a simple index of investment that underlies the leaf economics spectrum (LES), ranging from cheap, low-LMA leaves with a fast rate of return per-unit investment to expensive, high-LMA leaves with a slow rate of return [@Wright2004a].
On the other hand, because the lifetime return on investment depends on LL [@Westoby2000; @Falster2012], traits should be normalized by their annualized construction costs (or mass/LL), which should be used as a normalizer in leaf economics analyses [@Osnas2013].
In global analyses, normalizing leaf traits by mass/LL yields similarly weak correlations as area-normalization, because leaf area is roughly proportional to mass/LL across global species [@Osnas2013].
Area-dependence of metabolic traits and the inconsistent correlation strengths obtained from different normalizers [@Lloyd2013; @Osnas2013] do not invalidate mass-normalization or the LES [@Westoby2013].
These considerations do, however, lead us to question the evidence for a single dominant axis of leaf functional diversity, which has important implications for how trait variation is represented in ecosystem models [@Bonan2002; @Sakschewski2016].
On the one hand, mass-normalization has been justified based on economic principles alone [@Kikuzawa1991; @Westoby2013], because leaf mass provides a simple index of investment that underlies the leaf economics spectrum (LES), ranging from cheap, low-LMA leaves with a fast rate of return per-unit investment to expensive, high-LMA leaves with a slow rate of return [@Wright2004a].

To understand why the same data can be interpreted as either supporting or opposing the existence of a single dominant axis of leaf trait variation, consider the conceptual model proposed by @Osnas2018, in which LMA is comprised of two additive components:
metabolic LMA (LMAm) -- the mass per area of chloroplasts and other metabolically active leaf components that contribute directly to photosynthesis and respiration -- and structural LMA (LMAs) -- the mass per area of structural leaf components that contribute to toughness and durability [@Kitajima2012; @Kitajima2016; @Onoda2017].
Suppose these two LMA components are independent axes of functional variation, so that LMAm and LMAs are uncorrelated across species ([@fig-hypo]a).
These two independent axes can be translated into either a two-dimensional trait space (if metabolic traits are area-normalized; [@fig-hypo]b) or a one-dimensional trait space (if metabolic traits are mass-normalized; [@fig-hypo]c).
While both [@fig-hypo]b and [@fig-hypo]c are both 'correct' representations of the same data, they lead to different perceptions about the dimensionality of leaf functional variation.
If LMAm, LMAs, and/or other axes are largely independent and have distinct functional consequences, then it would not be possible to accurately represent functional variation with a single axis.

The hypothetical example in [@fig-hypo] shows how mass-normalization can, in principle, make a two-dimensional trait space appear one-dimensional, but the dimensionality of functional variation in real leaf assemblages remains an open question.
One way to better understand the dimensionality of leaf trait variation is to compare models with different numbers of dimensions, and to ask if models with multiple dimensions provide improved statistical fits and conceptual insights compared to a single axis.
For example, the two-dimensional 'LMAm-LMAs' model proposed by @Osnas2018 could be compared to a one-dimensional LMA model in terms of their capacities to explain variation in other traits.
However, implementing the 'LMAm-LMAs' model is challenging, because although certain leaf mass components can be neatly classified as 'metabolic' or 'structural' [@Poorter2009; @Osnas2018], other leaf mass components cannot.
For example, thick cell walls contribute to structural toughness [@Onoda2015], but at least some cell wall mass is required for the biomechanical support that enables photosynthesis.
Thus, partitioning LMA into different functional components requires novel empirical or modeling approaches.

In this paper, we present a statistical modeling framework to partition LMA into metabolic and structural components: LMAm and LMAs.
Our models assume that the influence of LMAm on metabolic rates is complex and non-linear, reflecting the intricate nature of leaf mass components, which resist simplistic 'metabolic' or 'structural' categorization.
We develop and test this two-dimensional trait space model using leaf trait data from two tropical forest sites (sun and shade leaves from wet and dry sites in Panama) and the GLOPNET global leaf traits database [@Wright2004a].
We use the model to address the following questions:
(1) Are measured leaf traits (including photosynthetic capacity, dark respiration rate, LL, and concentrations of nutrients and cellulose) better predicted by a one-dimensional (total LMA) or two-dimensional (LMAm-LMAs) trait space model?
(2) What are the relative contributions of LMAm and LMAs to total LMA variance in different leaf assemblages (Panama sun leaves, Panama shade leaves, and the global flora)?
(3) Do LMAm and LMAs differ between evergreen and deciduous species, and between sun and shade leaves? If so, how? and
(4) How do the answers to the preceding questions inform our understanding of empirical patterns of trait variation (e.g., relationships among measured leaf traits)?

# Material and Methods

## Overview

We considered multiple approaches to modeling datasets that include observations of leaf mass per area (LMA), leaf lifespan (LL), net photosynthetic capacity (*A*~max~), and dark respiration rate (*R*~dark~).
These traits comprise four of the six traits in the global leaf economics spectrum (LES) analysis of @Wright2004a.
For simplicity, we did not include the other two LES traits – leaf nitrogen (N) and phosphorus (P) concentrations – in our modeling framework.
Instead, we reserved observations of leaf N and P concentrations (and cellulose content, when available) for independent model tests.
We considered a simple one-dimensional model that predicts LL, *A*~max~, and *R*~dark~ from LMA alone, as well as two-dimensional models that predict LL, *A*~max~, and *R*~dark~ from two additive LMA components: metabolic and structural leaf mass per area (LMAm and LMAs, respectively).
We formulate the models in terms of area-normalized *A*~max~ and *R*~dark~ (*A*~area~ and *R*~area~, respectively), as these trait forms share the same denominator as LMA and its components.

To implement the two-dimensional models, we developed a statistical modeling framework to partition LMA into additive LMAm and LMAs components (Appendices `r appendix_num("prior")`-`r appendix_num("model-fit")`).
By using power-law relationships, our models assume that LMAm the contribution of LMAm to metabolic rates is not direct or linear, which reflects that leaf mass components still cannot be neatly classified as 'metabolic' or 'structural'.
We fit the models to two datasets: the GLOPNET global leaf traits dataset [@Wright2004a], which primarily represents interspecific variation; and the Panama dataset described by @Osnas2018, which includes traits for both sun and shade leaves at wet and dry tropical forest sites.
Because LMAm and LMAs are modeled (rather than observed), the two-dimensional models require one parameter per analysis unit (a species in GLOPNET or a species $\times$ canopy-position in the Panama dataset) to partition observed LMA into LMAm and LMAs.
Given the large number of parameters in these models, we performed tests with randomized data to evaluate whether our two-dimensional models were prone to creating patterns from noise, which could lead to spurious conclusions.
These tests suggested that our two-dimensional modeling approach revealed meaningful patterns in the observed trait data (`r appendix("rand")`).

To further understand the intricacies of LMA to photosynthetic capacity (`r appendix("mass-prop")`), we simulated datasets varying in LMAm and LMAs (`r appendix("mass-prop-sim")`).
These simulations allowed us to explore the controversial mass versus area normalization debate in leaf trait analysis.
Additionally, we performed variance partitioning to determine the contributions of LMAm and LMAs to total LMA variance, using both simulated and actual datasets (`r appendix("vpart")`).

All statistical analyses were conducted in R version 4.2.1 [@RCoreTeam2022] using the R package *targets* version 0.12.0 for workflow management [@Landau2021]

## Modeling leaf lifespan, photosynthetic capacity, and dark respiration rate

We considered five types of models, ranging from simple models with LMA as the sole predictor, to more complex models in which LMA was partitioned into LMAm and LMAs.
In all models, the unit of analysis is a 'leaf sample', defined as a species in the GLOPNET dataset or a species $\times$ canopy position in the Panama dataset (the datasets are described below).

First, we considered a simple set of models with LMA as the sole predictor for *A*~area~, *R*~area~, and LL according to power-law relationships (e.g., $A_{\mathrm{area}} = a\mathrm{LMA}^b$).

Next, we considered models in which LMA is partitioned into additive metabolic and structural components for leaf sample *i*:

$$
\begin{aligned}
  &\mathrm{LMA}_{i} =\mathrm{LMAm}_{i} + \mathrm{LMAs}_{i} \\
  &\mathrm{LMAm}_{i} = f_{i} \mathrm{LMA}_{i} \\
  &\mathrm{LMAs}_{i} = (1 - f_{i})  \mathrm{LMA}_{i}
\end{aligned}
$$ {#eq-LMA}

where the *f~i~* values -- the fractions of LMA comprised of LMAm for each sample *i* --  are estimated as latent variables in our modeling framework (see details below).
We assumed that the observed values of *A*~area~, *R*~area~ and LL were related to the unobserved values LMAm and LMAs according to multivariate power laws:

$$
\mathrm{E}[A_{\mathrm{area} \, i}]
= \alpha_0\mathrm{LMAm}_{i}^{\alpha_m}\mathrm{LMAs}_i^{\alpha_s}  =  \alpha_0 (f_i \mathrm{LMA}_{i})^{\alpha_m} \bigl\{(1-f_i) \mathrm{LMA}_{i}\bigr\}^{\alpha_s}
$$ {#eq-Aarea}

$$
\mathrm{E}[R_{\mathrm{area} \, i}]
= \gamma_0\mathrm{LMAm}_{i}^{\gamma_m} \mathrm{LMAs}_{i}^{\gamma_s}
= \gamma_0 (f_i \mathrm{LMA}_{i})^{\gamma_m} \bigl\{(1-f_i)\mathrm{LMA}_{i}\bigr\}^{\gamma_s}
$$ {#eq-Rarea}

$$
\mathrm{E}[\mathrm{LL}_i] = \beta_0\mathrm{LMAm}_{i}^{\beta_m} \mathrm{LMAs}_{i}^{\beta_s}  = \beta_0 (f_i \mathrm{LMA}_{i})^{\beta_m} \bigl\{(1-f_i) \mathrm{LMA}_{i}\bigr\}^{\beta_s} \qquad(4a)
$$

$$
\mathrm{E[LL}_i] = \beta_0\mathrm{LMAm}_{i}^{\beta_m} \mathrm{LMAs}_{i}^{\beta_s} \mathrm{exp}(\theta \mathrm{Light}_i) \qquad(4\mathrm{b})
$$

where E[$\cdot$] indicates expected value;
the $\alpha$, $\beta$, and $\gamma$ terms are fitted parameters;
and the logarithms of *A*~area~, *R*~area~, and LL are assumed to have a multivariate normal distribution (`r appendix("prior")`).
In preliminary analyses, we also considered alternative (non-power-law) forms, but none of these performed better than the power-law forms and are not discussed further.

We also considered models in which LMAs in `r eq("LL")` was replaced by leaf structural density (LMAs/LT, where LT is leaf thickness), which is based on the observation that cellulose density is a good predictor for LL in some species assemblages [@Kitajima2012; @Kitajima2013].
These models with leaf thickness yielded qualitatively similar results as those without leaf thickness, but they did not perform as well in the model evaluation (see below).

For example, the parameter set of [$\alpha_m$, $\alpha_s$, $f_i$] with values [0.3, 0.6, 0.2] produces the same model output as the set [0.6, 0.3, 0.8] in `r eq("Aarea")`.
To ensure that the model was identifiable, we imposed two broad assumptions:
(i) *A*~area~ depends more strongly on metabolic leaf mass (LMAm: parameter $\alpha_m$ in `r eq("Aarea")`) than structural leaf mass (LMAs: parameter $\alpha_s$), and
(ii) LL depends more strongly on LMAs ($\beta_s$ in `r eq("LL")` than LMAm ($\beta_m$).
The first assumption was implemented in different model versions either by setting $\alpha_s$ = 0 or by imposing the constraint $\alpha_m$ > $\alpha_s$.
Similarly, the second assumption was implemented either by setting $\beta_m$ = 0 or by imposing the constraint  $\beta_s$ > $\beta_m$.
See Table 1 Appendix S2 for more detail.

Finally, we considered a functional form that accounts for the effects of light availability on LL.
This light-dependent model is motivated by optimal LL theory, which predicts decreasing LL with increasing light availability [@Kikuzawa1991], and also by the often-observed 'LMA counter-gradient', whereby LL and LMA positively covary across species but negatively covary within species across light gradients [@Lusk2008; @Russo2016; @Osnas2018].
Mechanistically modeling how light affects LL via leaf carbon balance [@Xu2017] is beyond the scope of our study.
Instead, we introduced light effects in a simple way by modifying `r eq("LL")`a  to `r eq_num("LL")`b  where the dummy variable Light~*i*~ is set to 1 for sun leaves and 0 for shade leaves, and exp($\theta$) is the sun:shade LL ratio.

## Datasets

We fit the models described above using observations of LMA (g m^-2^), *A*~area~ (mol s^-1^ m^-2^), *R*~area~ (mol s^-1^ m^-2^) and LL (months) from the GLOPNET global leaf traits database [@Wright2004a] and from two tropical forest sites in Panama: Parque Natural Metropolitano (PNM, “dry site”) and Bosque Protector San Lorenzo (SL, “wet site”).
The GLOPNET data primarily represents interspecific variation (the dataset reports 2,548 species $\times$ site combinations, with 2,021 unique species and only 350 species occurring at more than one site) and only reports data for sun leaves if data for both sun and shade leaves are available [@Wright2004a].
In contrast, the Panama dataset represents both inter- and intraspecific variation, including leaves sampled at two canopy positions ("sun": full sun at the top of the canopy; and "shade": well shaded, sampled within 2 m of the forest floor) from trees within reach of canopy cranes.
The dry PNM site is a semi-deciduous coastal Pacific forest with a 5-month dry season from December-April and 1740 mm of annual rainfall [@Wright2003].
The wet SL site is an evergreen Caribbean coastal forest with 3100 mm of annual rainfall [@Wright2003].
Additional details of the Panama dataset are described in @Osnas2018.

We restricted our analysis to database records for which all four traits (LMA, *A*~area~, *R*~area~, and LL; each typically averaged over multiple leaves) were available.
After excluding database records that lacked one or more of the four traits, 198 samples for 198 unique species were available for GLOPNET, and 130 samples for 104 unique species were available for Panama (dry and wet sites combined; 26 species sampled in both sun and shade; no species with all four traits available at both sites).
In addition to LMA, *A*~area~, *R*~area~, and LL, the model based on structural leaf density also requires observations of leaf thickness, which was not available in GLOPNET but was available for 106/130 Panama samples.
Both the GLOPNET and Panama datasets include additional traits that we used to interpret model results, but which were not used to fit models.
These traits include nitrogen and phosphorus content per leaf unit area (*N*~area~ and *P*~area~; g m^-2^) and leaf habit (deciduous or evergreen) in both datasets, and cellulose content per unit area (CL~area~; g m^-2^) in the Panama dataset.

## Model estimation and evaluation

We modeled *A*~area~ and *R*~area~ using Eqs. `r eq_num("Aarea")` and `r eq_num("Rarea")`, respectively, for both GLOPNET and Panama.
To model LL for GLOPNET, we used `r eq("LL")`a, because GLOPNET does not report canopy position (and prioritizes data for sun leaves, as described above).
To model LL for Panama, we also used `r eq("LL")`b (light effects model), which was motivated by the negative intraspecific LL-LMA relationship observed in Panama [@Xu2017; @Osnas2018] and elsewhere [@Lusk2008; @Russo2016].

Posterior distributions of all parameters were estimated using the Hamiltonian Monte Carlo algorithm (HMC) implemented in Stan [@Carpenter2017].
We used non-informative or weakly informative prior distributions [@Lemoine2019].
Prior distributions for the latent variables *f~i~* (which are used to partition LMA into LMAm and LMAs according to `r eq("LMA")` were non-informative uniform(0, 1) distributions, (i.e., LMA was partitioned based on patterns in the data).
See `r appendix("prior")` for more details.
The Stan code use to fit models is available from Github at: [https://github.com/mattocci27/LMAms](https://github.com/mattocci27/LMAms).
Convergence of the posterior distribution was assessed with the Gelman-Rubin statistic with a convergence threshold of 1.05 for all parameters [@Gelman2013].

Considering the complexity of our two-dimensional modeling approach, which involves numerous latent parameters, we validated our estimates of LMAm and LMAs through randomized tests, detailed in `r appendix("rand")`.
These tests support our assumption that the patterns observed in the GLOPNET and Panama datasets are meaningful and allow for a substantial exploration of our questions.

To compare the performance of different models, we used Pareto-smoothed importance sampling leave-one-out cross-validation (PSIS-LOO; [@Vehtari2017]).
PSIS-LOO is an accurate and reliable approximation to standard leave-one-out cross-validation (LOO), which is a robust method for comparing models with different numbers of parameters [@Vehtari2017].
We used PSIS-LOO to calculate the LOO Information Criterion (LOOIC) for each model.
A model with lower LOOIC is considered to have better predictive accuracy [@Vehtari2017].

# Results

**1. Decomposing LMA into metabolic and structural components leads to improved predictions of *A*~area~, *R*~area~ and LL.**
For both the GLOPNET and Panama datasets, two-dimensional models (LMAm-LMAs) performed better in cross validation than one-dimensional (LMA) models (Table 1).
Thus, predictions of *A*~area~, *R*~area~ and LL were improved by partitioning LMA into separate metabolic and structural components (see details below and Tables 1-2).
Each of these traits (*A*~area~, *R*~area~ and LL) had a strong positive relationship with either LMAm or LMAs (but not both), which was always stronger than the corresponding relationship with total LMA (Figs. [-@fig-gl_point]-[-@fig-pa_point]).

In the best GLOPNET model (lowest LOOIC), *A*~area~ increased with LMAm and decreased with LMAs, *R*~area~ increased with LMAm and had little dependence on LMAs, and LL increased with LMAs and was independent of LMAm (Tables 1-2; @fig-gl_point).

In the best Panama model, *A*~area~ increased with LMAm and was independent of LMAs, *R*~area~ increased with LMAm and had little dependence on LMAs, and LL increased with LMAs and was independent of LMAm (Tables 1-2; @fig-pa_point).
The best Panama model also included the effect of light (sun vs. shade) on LL (Tables 1-2;  @fig-ll_point).
The light effect was not tested for GLOPNET because GLOPNET does not report canopy position.

**2. Nearly all leaf dark respiration is associated with metabolic leaf tissue mass.**
According to our model results, metabolic leaf mass (LMAm) accounted for nearly all leaf dark respiration; i.e., the estimated dark respiration rate per-unit structural mass ($\gamma_s$) was close to zero for both GLOPNET and Panama (Table 2).
Thus, although building costs appear similar for different leaf chemical components and tissues [@Williams1989; @Villar2001], our results suggest that leaf mass associated with structural function (i.e., toughness and LL) has roughly zero maintenance cost.

**3. Evergreen leaves have greater LMAs than deciduous leaves, and sun leaves have both greater LMAm and LMAs than shade leaves.**
Leaf habit (evergreen vs. deciduous) explained very little variation in LMAm but substantial variation in LMAs for both GLOPNET and Panama ([@fig-vpart]a-b).
Despite the relatively large residuals in LMAs [@fig-vpart]a-b), the higher total LMA in evergreen compared to deciduous leaves was primarily due to differences in LMAs (Figs. [-@fig-box_de], `r s_fig_num("box_inter")`, and `r s_fig_num("box_frac_de")`).
In the Panama dataset, light (sun vs. shade) explained much more variation in LMAm than LMAs, whereas site (wet vs. dry) explained much more variation in LMAs than LMAm ([@fig-vpart]c).
Panama wet-site leaves had higher total LMA than dry-site leaves due to greater LMAs when we compared the same canopy strata (Figs. [-@fig-box_pa] and `r s_fig_num("box_inter")`).
In contrast, Panama sun leaves had higher total LMA than shade leaves primarily due to greater LMAm, with a smaller but significant contribution from LMAs (Figs. [-@fig-box_pa] and `r s_fig_num("box_inter")`).
Thus, LMAs comprised a larger fraction of total LMA for shade leaves than for sun leaves (`r s_fig("box_frac_de")`).

**4. LMA variance components affect the area- vs. mass-dependence of leaf photosynthetic capacity (*A*~max~).**
Analysis of simulated data showed that the percent of LMA variation due to LMAs variation (`r eq("var")`) strongly affected the area- vs. mass-dependence of *A*~max~ (i.e., the degree to which whole-leaf *A*~max~ depends on leaf area or mass; @Osnas2013; @Osnas2018).
Generally, mass-dependence of *A*~max~ decreased (area-dependence increased) with increasing LMAs variance (@fig-mass_prop).
Other factors, including the sensitivity of *A*~max~ per-unit leaf area (*A*~area~) to variation in LMAm and LMAs ($\alpha_m$ and $\alpha_s$ in `r eq("Aarea")`) and covariance between LMAm and LMAs, also affect the mass-dependence of *A*~max~ (@fig-mass_prop, Figs. `r s_fig_num("mass_prop_sim")`-`r s_fig_num("mass_prop_comp")`).
For example, LMAs variance was relatively low in GLOPNET (`r var_val$GL |> round(0)`% of total LMA variance).
However, variation in *A*~max~ in GLOPNET was strongly area-dependent.
This is because the relatively weak dependence of *A*~area~ on LMAm and the negative dependence of *A*~area~ on LMAs (Table 2) led to a weak overall dependence of *A*~area~ on LMA.
Thus, mass-dependence of *A*~max~ in GLOPNET was similar or weaker than in the Panama data, despite LMAs variance comprising a larger fraction of LMA variance in Panama (`r var_val$sun_intra |> round(0)`% and `r var_val$shade_intra |> floor()`% of total LMA variance for Panama sun and shade leaves, respectively).

**5. Nitrogen and phosphorus per-unit leaf area are strongly correlated with LMAm, and cellulose per-unit leaf area is strongly correlated with total LMA.**
In the GLOPNET dataset, *N*~area~ and *P*~area~ had strong positive correlations with LMAm, but only weak correlations with LMAs (`r s_fig("gl_point_np2")`).
Similarly, in the Panama dataset, *N*~area~ and *P*~area~ had strong positive correlations with LMAm but not with LMAs (@fig-pa_npc).
Cellulose content per-unit leaf area (CL~area~), which was only available for the Panama dataset, was most strongly correlated with total LMA, but also showed a clear relationship with LMAs that was consistent across canopy positions and sites (@fig-pa_npc).
Partial regression plots showed that the higher *N*~area~, *P*~area~, and CL~area~ in sun leaves (compared to shade leaves; @fig-pa_npc) was due to sun leaves having higher LMAm than shade leaves (`r s_fig("pa_point_npc_par")`).

# Discussion

Our analysis of the GLOPNET global dataset and a dataset including sun and shade leaves from wet and dry Panama sites demonstrates that decomposing LMA variation into separate metabolic and structural components (LMAm and LMAs, respectively) leads to improved predictions of photosynthetic capacity (*A*~max~), dark respiration rate (*R*~dark~), and leaf lifespan (LL) relative to a simpler one-dimensional model based on total LMA.
Furthermore, LMAm and LMAs showed clear relationships with independent model evaluation data (nitrogen, phosphorus, and cellulose content per-unit leaf area).
In our analysis, LMAm and LMAs are model-inferred (rather than observed) traits.
Nevertheless, several observations suggest that we can derive meaningful insights from these modeled LMA components.
Firstly, our analysis revealed several patterns that are consistent with previous analyses of measured traits (summarized below).
Secondly, although our modeling approach requires many parameters to partition LMA into LMAm and LMAs (one latent variable per leaf sample), which could lead to over-fitting or spurious results, our model performed poorly (often failing to converge) when applied to randomized data.
Thus, the model structure is not prone to creating signals from noise.
Together, these observations lend credibility to our model-inferred estimates of LMAm and LMAs.
We first summarize and interpret our model results in the context of previous studies.
We then discuss the new insights derived from our analysis and implications for vegetation models, including the land components of Earth system models.

## Modeling *A*~max~, *R*~dark~, and LL as functions of metabolic and structural LMA

*A*~max~ per-unit leaf area (*A*~area~) increased with LMAm and was either unrelated to LMAs (Panama) or decreased with LMAs (GLOPNET; Figs. [-@fig-gl_point]-[-@fig-pa_point] and Table 2).
The decline in GLOPNET *A*~area~ with increasing LMAs may be due to decreasing mesophyll conductance, which decreases with increasing cell wall thickness [@Evans2009; @Terashima2011; @Onoda2017].
This trend may be less apparent in our Panama analysis due to the relatively small LMA range (`r round(lma_val$PA$min)`-`r round(lma_val$PA$max)` g m^2^ in Panama, compared to `r round(lma_val$GL$min)`-`r round(lma_val$GL$max)` g m^2^ in GLOPNET) or may be obscured by pooling sun and shade leaves in a single analysis.
In any case, an increase in *A*~area~ with LMAm but not LMAs, as observed for both GLOPNET and Panama, is consistent with the view that LMA is a composite trait comprised of metabolically active and inactive components [@Poorter2009; @Osnas2018; @Lichstein2021].

*R*~dark~ per-unit leaf area (*R*~area~) increased with LMAm and was unrelated to LMAs (Figs. [-@fig-gl_point]-[-@fig-pa_point] and Table 2), consistent with the view that LMAs is comprised of non-metabolic leaf mass components, such as cellulose, lignin, and other fibres [@Onoda2011; @Kitajima2012; @Kitajima2016].
Median values of LMAs, as a percent of total LMA, were ~60% for evergreen species and ~40% for deciduous species (`r s_fig("box_frac_de")`), and 15-40% for shade leaves and 50-70% for sun leaves (`r s_fig("box_frac_pa")`).
These estimates align with measurements of cell wall (fibre) content from a global dataset (14–77%, median 47% of total leaf mass, with higher values for shade than sun leaves; Onoda et al. 2011) and a Panama study site (Barro Colorado Island, BCI: ~50% mean across species, with higher values for shade leaves; Kitajima et al. 2016).

LL increased with LMAs but was unrelated to LMAm (Figs. [-@fig-gl_point]-[-@fig-pa_point] and Tables 1-2), consistent with well-established relationships between leaf fibre content, toughness, and LL [@Onoda2011; @Onoda2017; @Kitajima2012; @Kitajima2016].
For a given LMAs, shade leaves had longer LL than sun leaves (see [@fig-pa_point]i and parameter $\theta$ in Table 2).
Thus, partitioning LMA into LMAm and LMAs is not sufficient to explain the widely observed counter-gradient, whereby LL increases with increasing LMA across species but decreases with increasing LMA within species from shade to sun [@Lusk2008; @Russo2016; @Osnas2018].
The failure of LMAs to provide robust predictions of LL across canopy positions may reflect factors unlreated to leaf toughness - e.g., how optimal LL depends on light availability [@Kikuzawa1991] - as well as variation in leaf toughness for a given value of LMAs.
In a global meta-analysis, leaf toughness was strongly related to toughness per-unit leaf tissue density, but only weakly related to leaf thickness or density [@Onoda2011].
Accordingly, on BCI, Panama, the difference between shade and sun leaves was greater for cellulose content (% of leaf mass) than cell wall content (which includes cellulose, as well as other fibres that provide less structural toughness).
Thus, just as not all LMA is equal, it is almost certainly true that not all LMAs is equal, with some LMAs components (e.g., cellulose) imparting greater toughness and longevity than others [e.g., lignin in secondary cell walls of vascular tissue, @Kitajima2016].

In summary, the dependencies of *A*~area~, *R*~area~, and LL on model-inferred values of LMAm and LMAs are broadly consistent with previous analyses of measured traits.
Our results support the hypothesis that LMA can be partitioned into metabolic and structural components, the latter having a maintenance respiration rate indistinguishable from zero.
Like total LMA, LMAs is likely a composite trait, including cuticle and fibres with different functions.
A more refined modeling approach that links different structural components to specific functions (e.g., mesophyll cell wall thickness that imparts overall lamina toughness vs. fibres in vascular tissue that contribute to cavitation resistance) could potentially provide a mechanistic explanation for LL variation across canopy positions as well as ecophysiological tradeoffs across edaphic-climate gradients.

## Implications for understanding LL variation

Optimal LL theory [@Kikuzawa1991] predicts that LL should increase with toughness (which increases building costs and decreases the rate of photosynthetic decline with leaf age) and decrease with initial (pre-aging) net assimilation rate.
Consistent with these predictions, our results indicate an increase in LL with increasing LMAs and decreasing light.
However, our results do not support a decrease in LL with increasing LMAm (which, along with the environmental conditions, should determine net assimilation rate).
For example, within each Panama site (wet and dry), *A*~area~ of sun leaves varied by over a factor of four and was strongly determined by LMAm ([@fig-pa_point]b).
Yet, the best-performing models excluded effects of LMAm on LL (Table 1).
Modeling LL as a function of LMAs, rather than total LMA, partially explains the sun-shade LL difference (compare Figs. [-@fig-pa_point]g and [-@fig-pa_point]i), and explicit consideration of fibre content and toughness may provide a more complete understanding of LL variation with light (see details above).
Thus, while optimal LL theory provides important insights regarding the coordination of structural toughness and net assimilation rate, toughness itself may provide a sufficient proximate explanation for LL variation.

## Implications for understanding relationships between photosynthetic capacity and LMA

@Osnas2018 suggested that in leaf assemblages where LMA variation is largely due to variation in structural leaf mass components that contribute to toughness but not photosynthetic capacity, metabolic traits (including *A*~max~) should be primarily area-dependent rather than mass-dependent (parameter *b* in `r eq("mass")`, with *b* > 0.5 indicating greatermass- than area-dependence).
Our simulations show that all else being equal, area-dependence of *A*~max~ increases (mass-dependence decreases) as LMAs variance increases ([@fig-mass_prop]), as hypothesized by @Osnas2018.
However, our analysis also revealed complexities that were not anticipated by @Osnas2018.
For example, if *A*~area~ only weakly increases with LMAm and/or decreases with LMAs (as in GLOPNET; Table 2), then *A*~max~ can be strongly area-dependent (weakly mass-dependent) even if LMAs accounts for relatively little LMA variance ([@fig-mass_prop]).
We conclude that area- vs. mass-dependence depends not only LMAs variance, but also on how metabolic traits vary with different LMA components.

Several factors could contribute to the less-than-proportional increase in *A*~area~ with LMAm ($\alpha_p$ < 1 for both GLOPNET and Panama; Table 2).
High LMAm may lead to low mesophyll conductance, e.g., due to a longer CO~2~ diffusion pathway [@Evans2009].
Similarly, high LMAm may lead to a high level of chloroplast self-shading within a leaf [@Terashima2011].
Both of these factors would decrease the marginal benefit of additional units of LMAm, and this marginal decrease may be especially strong in GLOPNET due to its large LMAm range (`r para_val$GL$LMAm_min |> round(1)` - `r para_val$GL$LMAm_max |> round()` g m^2^) relative to Panama (`r para_val$PA$LMAm_min |> round(1)` - `r para_val$PA$LMAm_max |> round()` g m^2^).

Finally, while much of the model-inferred LMAm may be directly or indirectly related to photosynthesis, 'LMAm' may also include metabolic leaf mass unrelated to photosynthesis and/or non-metabolic leaf mass that is statistically assigned (by default) to LMAm simply because it does not affect LL.
For example, non-structural carbohydrates, which are ignored in our model, can account for 5-8% of leaf mass in deciduous species [@Wurth2005] and up to 20% of leaf mass in evergreen species [@Martinez-Vilalta2016].
Although scaling parameters in our model allow us to include non-metabolic leaf mass in LMAm, a more refined model that accounts for more than two LMA components may reveal stronger relationships between metabolic traits (*A*~area~ and *R*~area~) and LMAm.


In addition to providing insights as to why metabolic traits tend to be area-dependent across species, our analysis also helps explain why these same traits tend to be mass-dependent within species across light levels [@Osnas2018].
LMA increased from shade to sun primarily due to an increase in LMAm (Figs. [-@fig-box_pa] and `r s_fig_num("box_inter")`), which should lead to strong mass-dependence of *A*~max~ and *R*~dark~ ([@fig-mass_prop]) and likely reflects an increase in the size and number of palisade mesophyll cells with increasing light levels [@Onoda2008; @Terashima2011].
In contrast to our interspecific analysis within sun and shade leaves ([@fig-mass_prop]), estimates of intraspecific mass-dependence of metabolic traits across light levels are around 0.8-1.0 for the same Panama dataset [@Osnas2018].

Although values of *b* greater than 1 are theoretically possible (which implies an exponential increase in area-normalized traits with increasing LMA), two factors likely constrain the observed values of *b*.
Firstly, at least some of the LMA increase from shade to sun is due to LMAs (Figs. [-@fig-box_pa] and `r s_fig_num("box_inter")`), which may reflect the additional LMAs needed to protect and support the high LMAm of sun leaves and/or additional LMAs required to withstand hydraulic or other stress in the upper canopy.
Secondly, as discussed above, *A*~area~ and *R*~area~ increase at a less-than-proportional rate with increasing LMAm, which in turn moderates the LMA scaling exponent (*b*).

## Implications for terrestrial ecosystem models

For over two decades, low-dimensional trait tradeoffs have been explored as a means to represent plant functional diversity in the vegetation components of biogeochemical models [@Moorcroft2001; @Bonan2002; @Verheijen2013], with the ultimate goal of adding ecological and physiological realism to coupled Earth System Models [ESMs, @Moorcroft2006; @Wullschleger2014].
Recently, there has been a proliferation of interest in this topic [@Scheiter2013; @Fisher2015; @Sakschewski2015; @Sakschewski2016; @DantasdePaula2021] along with demographic approaches capable of representing the dynamics of diverse communities in ESMs [@Fisher2018].
The leaf economics spectrum (LES) has been interpreted as a single dominant axis of global leaf functional variation [@Wright2004a], and this simple framework has been adopted in some of these dynamic vegetation models.
Our results demonstrate the value of considering metabolic and structural leaf mass separately, and indicate weak if any interspecific correlation between these LMA components (`r s_fig("LMAm_LMAs")`).
We therefore conclude that a one-dimensional representation of leaf functional variation is too simplistic.
Consistent with our conclusion that leaf function is multi-dimensional, analyses of leaf nitrogen indicate significant independent contributions of metabolic (e.g., rubisco) and structural leaf nitrogen components [@Dong2017; @Dong2022], with over 20% of total leaf nitrogen in high-LMA leaves occurring in cell walls [@Onoda2017].
Similarly, partitioning leaf phosphorus reveals substantial, LMA-dependent metabolic, structural, and nucleic acid fractions [@Hidaka2011].
We suggest that future biogeochemical model developments consider multiple axes of leaf functional diversity, and that models with coupled carbon-nutrient cycles consider how LMA, nitrogen, and phosphorus are allocated to different functions

# Acknowledgments

We thank Jonathan Dushoff for statistical advice and Martijn Slot for helpful comments that improved the paper.
Mirna Samaniego and Milton Garcia provided indispensable assistance in data collection.
We thank the Andrew W. Mellon Foundation for supporting this work.
MK was supported by
a Xishuangbanna State Rainforest Talent Support Program,
a CAS President's International Fellowship Initiative (2020FYB0003),
a ZiHui (Wisdom) Yunnan Program (202203AM140026), and
a Postdoctoral Fellowship for Research Abroad from the Japan Society for the Promotion of Science (25-504).

# Conflict of Interest

The authors declare no conflicts of interest.

# Author contributions

MK, JWL and JLDO conceived of the study;
KK provided guidance on leaf physiology and interpreting trait data;
KK, SJW and SAVB contributed data;
MK devised the analytical approach and performed analyses;
MK and JWL wrote the first draft of the manuscript, and all authors contributed to revisions.

# Data availability statement

Data, codes and computing environments to reproduce this manuscript will be archived on Zenodo at https://doi.org/xxx and also available on Github at https://github.com/mattocci27/LMAms.

# References

<div id="refs"></div>

\newpage

# Tables

Table 1.
List of models fit to GLOPNET and Panama data ordered from highest to lowest cross-validation performance.
Lower approximate leave-one-out cross-validation information criterion (LOOIC) indicates better predictive accuracy.
For GLOPNET, we considered two model forms were considered: *A*~area~, *R*~area~, and LL were modeled as power laws of LMA (one-dimensional model) or both LMAm and LMAs (two-dimensional model; see Eqs. `r eq_num("Aarea")`, `r eq_num("Rarea")`, and `r eq_num("LL")`a).
For the Panama data, which includes canopy position (sun vs. shade) for each leaf sample, we considered the same two model forms as for GLOPNET and three additional light-dependent forms for LL (as in `r eq_num("LL")`b), LMA and light (one-dimensional); LMAm, LMAs, and light (two-dimensional); and LMAm, leaf structural density (LSD = LMAs/LT), and light (two-dimensional).
We combined the two-dimensional model forms with four types of parameter constraints:
(i) $\alpha_s$ = 0 and $\beta_m$ = 0,
(ii) $\beta_m$ = 0,
(iii) $\alpha_s$ = 0, and
(iv) $\alpha_m$ > $\alpha_s$ and $\beta_s$ > $\beta_m$.
Those constraints were based on the following hypotheses:
(1) *A*~area~ depends more strongly on metabolic leaf mass (LMAm: parameter $\alpha_m$ in `r eq("Aarea")`) than structural leaf mass (LMAs: parameter $\alpha_s$), and
(2) LL depends more strongly on LMAs ($\beta_s$ in `r eq("LL")` than LMAm ($\beta_m$).
See the method and `r appendix("model-fit")` for more details.

```{r lootab, eval=TRUE, echo=FALSE}
loo_tbl <- read_csv(here("data/model_selection.csv"))  |>
  # arrange(Data) |>
  dplyr::select(Model, Constraints, N, LOOIC)

if ("html" == knitr::opts_knit$get("rmarkdown.pandoc.to")) {
  knitr::kable(loo_tbl, format = "html", escape = FALSE) |>
    kable_classic(full_width = FALSE) |>
    pack_rows("GLOPNET", 1, 5) |>
    pack_rows("Panama", 6, 23)
} else if ("docx" == knitr::opts_knit$get("rmarkdown.pandoc.to")) {
  tmp <- c("**GLOPNET**", rep("", 4), "**Panama**", rep("", 17))
  tmp2 <- cbind(tmp, loo_tbl)
  colnames(tmp2)[1] <- ""
  tmp2 |>
    kable(format = "pipe", escape = FALSE)
} else if ("latex" == knitr::opts_knit$get("rmarkdown.pandoc.to")) {
  loo_tbl |>
    kable(format = "latex", escape = TRUE, booktabs = TRUE) |>
    kable_classic(full_width = FALSE) |>
    pack_rows("GLOPNET", 1, 5) |>
    pack_rows("Panama", 6, 23)
}
```

\newpage

Table 2.
Posterior medians [95% credible interval] of parameters for the best GLOPNET and Panama models.
Bold values are significantly different from zero based on the 95% credible interval.

```{r, echo=FALSE, eval=TRUE}
para_tbl <- read_csv(here("data/para_tbl.csv"))

if ("html" == knitr::opts_knit$get("rmarkdown.pandoc.to")) {
  knitr::kable(para_tbl, format = "html", escape = FALSE) |>
    kable_classic(full_width = FALSE)
} else if ("docx" == knitr::opts_knit$get("rmarkdown.pandoc.to")) {
  para_tbl |>
    kable(format = "pipe", escape = FALSE) |>
    kable_classic(full_width = FALSE)
} else if ("latex" == knitr::opts_knit$get("rmarkdown.pandoc.to")) {
  para_tbl |>
    kable(format = "pipe", escape = FALSE) |>
    kable_classic(full_width = FALSE)
}
```

\newpage

![
Simulated two-dimensional functional space that can result in either a two- or one-dimensional trait space, depending on how metabolic traits are normalized.
(a) Hypothetical independent variation in two leaf mass per area (LMA) components: metabolic leaf mass per area (LMAm, which largely determines per-area values of photosynthesis, respiration, and nutrient concentrations) and structural leaf mass per area (LMAs, which determines leaf toughness but has little effect on metabolic traits).
Panels b and c represent normalization by leaf area and leaf mass, respectively.
(b) Two-dimensional relationship between photosynthetic capacity (*A*~max~) per-unit leaf area and total LMA (equal to LMAm + LMAs).
(c) One-dimensional relationship between *A*~max~ per-unit leaf mass and total LMA.
Methods used to simulate the data are explained in Appendix S3.
](`r here::here("figs/hypo.png")`){#fig-hypo}


![
Observed and estimated leaf-trait relationships in the global GLOPNET dataset.
Estimates are from the best GLOPNET model (Table 1).
Leaf life span (LL), net photosynthetic rate per unit leaf area (*A*~area~), and dark respiration rate per unit leaf area (*R*~area~) are plotted against observed LMA, posterior medians of LMAm and LMAs.
Pearson correlation coefficients (*r*) for LMA (left column) and posterior medians of Pearson correlation coefficients ($\bar{r}$) or partial correlation coefficients ($\bar{\rho}$) for LMAm (middle column) and LMAs (right column) are shown.
Note that LL was modeled by LMAs alone due to improved model performance with this parameter constraint (Table 1).
Therefore, ($\bar{r}$) is reported for LL (single predictor variable), whereas ($\bar{\rho}$) is reported for *A*~area~ and *R*~area~ (multiple predictor variables).
](`r here::here("figs/gl_point.png")`){#fig-gl_point}

\newpage

![
Observed and estimated leaf-trait relationships in the Panama dataset.
Estimates are from the best Panama model (Table 1), which included effects of light on LL.
Details as for [@fig-gl_point].
Note that *A*~area~ and LL were modeled by LMAm and LMAs alone, respectively, due to improved model performance with these parameter constraints (Table 1).
The results shown here include all leaves in the Panama dataset.
The observed separation of LL between sun and shade leaves is accounted for in the model predictions [@fig-ll_point].
](`r here::here("figs/pa_point.png")`){#fig-pa_point}

\newpage

![
Relationship between leaf lifespan (LL) and LMAs in the Panama dataset, after accounting for the effects of light (suv vs. shade leaves; see `r eq("LL")`b).
The dashed line indicates the 1:1 relationship expected for residuals on the log-scale.
The posterior medians of the partial correlation coefficient ($\bar{\rho}$) is shown.
The results shown here include all leaves in the Panama dataset.
](`r here::here("figs/pa_point_par_ll.png")`){#fig-ll_point}

\newpage

![
Variance partitioning on LMA components between and within leaf habits (evergreen vs. deciduous) for the GLOPNET dataset and the Panama dataset, and between and within sites (wet vs. dry) and light (sun vs. shade) for the Panama dataset.
To isolate the effects of intraspecific variation, the Panama results shown here only include species for which both sun and shade leaves were available.
](`r here::here("figs/vpart_intra.png")`){#fig-vpart}

\newpage

![
Boxplots comparing leaf mass per area (LMA) and posterior medians of photosynthetic and structural LMA components (LMAm and LMAs, respectively) across deciduous (Dev) and evergreen (Eve) leaves in the GLOPNET dataset (a) and in the Panama dataset (b).
The center line in each box indicates the median, upper and lower box edges indicate the interquartile range, whiskers show 1.5 times the interquartile range, and points are outliers.
Groups sharing the same letters are not significantly different (P > 0.05; t-tests).
The Panama results only include species for which both sun and shade leaves were available.
Qualitatively similar results were obtained when all Panama species were included (`r s_fig("box_inter")`).
Note that the vertical axis is on a log~10~ scale.
](`r here::here("figs/box_de.png")`){#fig-box_de}

\newpage

![
Boxplots comparing leaf mass per area (LMA) and posterior medians of metabolic and structural LMA components (LMAm and LMAs, respectively) across sites (wet and dry) and canopy strata (sun and shade) in the Panama dataset.
These results only include species for which both sun and shade leaves were available.
Qualitatively similar results were obtained when all Panama species were included (`r s_fig("box_inter")`).
Details as for [@fig-box_de].
](`r here::here("figs/box_pa.png")`){#fig-box_pa}

\newpage

![
Relationships between mass dependency (*b* in `r eq("mass")`) and LMAs variance (relative to total LMA variance; `r eq("var")`) for the global GLPNET dataset, sun leaves in Panama, and shade leaves in Panama.
Solid lines indicate simulated medians and shaded regions indicate 95% confidence intervals.
Each point indicates the estimated values from the empirical data and represents interspecific variation (e.g., across species within a canopy position in the Panama dataset).
The y-axis values indicates if photosynthetic capacity (*A*~max~) is is primarily mass-dependent (*b* > 0.5) or primarily area-dependent (0.5 > *b* > 0), with *b* near 0 indicating  purely area-dependendence [@Osnas2018].
If *b* > 1, then *A*~max~ increases exponentially with LMA, which is not consistent with observed relationships [@Osnas2018].
](`r here::here("figs/mass_prop_mv.png")`){#fig-mass_prop}

\newpage

![
Measured traits in the Panama dataset related to photosynthesis and metabolism (nitrogen and phosphorus per-unit leaf area; *N*~area~ and *P*~area~) are better correlated with estimates (posterior medians) of the metabolic LMA component (LMAm) than the structural component (LMAs), whereas the opposite pattern occurs for a measured structural trait (cellulose per-unit leaf area; CL~area~).
*N*~area~, *P*~area~, and CL~area~ data were not used to fit the models, and are presented here as independent support for the model results.
Pearson correlation coefficients (*r*) for LMA and the posterior medians of Pearson correlation coefficients ($\bar{r}$) for LMAm and LMAs are shown.
Analogous results were obtained for *N*~area~ and *P*~area~ for GLOPNET (`r s_fig("gl_point_np2")`).
The results shown here include all leaves for which *N*~area~, *P*~area~  and CL~area~ are available.
](`r here::here("figs/pa_point_npc.png")`){#fig-pa_npc}
