---
title: "Katabuchi et al. Appendix S1-x: Decomposing leaf mass into metabolic and structural components explains divergent patterns of trait variation within and among plant species"
fontsize: 12pt
geometry: margin=1in
link-citations: yes
csl: templates/ecology.csl
bibliography: [LMA.bib]
output:
  bookdown::pdf_document2:
    latex_engine: pdflatex
    fig_caption: yes
    keep_tex: yes
    toc: no
    number_sections: no
    template: templates/eisvogel2.tex
    includes:
        in_header: templates/format.sty
    pandoc_args:
      - "--filter"
      - "pandoc-crossref"
  bookdown::word_document2:
       fig_width: 6
       reference_docx: templates/rmd_style.docx
       pandoc_args:
         - "--filter"
         - "pandoc-crossref"
  bookdown::html_document2:
    fig_width: 6
    fig_caption: yes
    number_sections: no
    pandoc_args:
        - "--filter"
        - "pandoc-crossref"
        - "--mathml"
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r caching, include=FALSE}
#library(methods)
library(knitr)
library(kableExtra)
#source("rmd_func.r")
```

# Randomization

hoge

\newpage

![
Pearson correlation coefficients for posterior medians of LMAm vs LMAs in the (a) GLOPNET and (b) Panama datasets. The non-significant *r* values indicate that a single axis could not accurately represent the two-dimensional space. Symbols as in Main Text Figs. 1-2.
](../figs/ps_point.png){#fig:LMAms}

\newpage

![
Boxplots comparing leaf mass per area (LMA), photosynthetic leaf mass per area (LMAp; posterior means), and structural leaf mass per area (LMAs; posterior means) across sites (wet and dry) and canopy strata (sun and shade) in Panama.
The results shown here include all leaves in the Panama dataset, whereas Fig. 5 in the main text only includes Panama species for which both sun and shade leaves were available.
Boxplot symbols as in Fig. 5.
Groups sharing the same letters are not significantly different (P > 0.05; t-tests).
](../figs/box_frac.png){#fig:PAplt}

\newpage


![
Measured traits related to photosynthesis and metabolism (nitrogen and phosphorus per-unit leaf area; Narea and Parea) are positively correlated with LMA and with estimates (posterior means) of the photosynthetic and structural LMA components (LMAp and LMAs, respectively) in the GLOPNET dataset.
LMAp yields stronger correlations and more consistent relationships compared to LMA and LMAs; e.g., evergreen and deciduous leaves align along a single relationship in panel b, but not in panels a or c.
](../figs/gl_point_np2.png){#fig:glnp}
