---
fontsize: 12pt
geometry: margin=1in
link-citations: yes
csl: templates/apa.csl
bibliography: [LMA.bib]
crossref:
  fig-title: Fig.
  fig-prefix: Fig.
  eq-prefix: Eq.
format:
  html:
    theme: spacelab #readable #sandstone #spacelab #flatly
    toc: true
    toc-depth: 2
    toc-title: Contents
    self-contained: true
    smooth-scroll: true
    highlight-style: github
  docx:
    toc: false
    number-sections: false
    highlight-style: github
    html-math-method: katex
  pdf:
    toc: false
    keep-tex: true
    latex-clean: false
    pdf-engine: pdflatex
    template: templates/eisvogel2.tex
    include-in-header:
      text: |
        \usepackage{xr}
        \externaldocument{LMAms_SI}
        \usepackage{lineno}
        \linenumbers
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(knitr)
library(kableExtra)
library(here)
# here::i_am("ms/LMAms_main.qmd")
# source(here("templates/rmd_func.R"))
# here("R/vpart.R") |> message()
source(here("R/vpart.R"))
# read_csv(here("data/vpart_inter.csv"))
var_val <- yaml::yaml.load_file(here("yml/var.yml"))
```

**Running title**: Metabolic and structural leaf mass

**Decomposing leaf mass into metabolic and structural components explains divergent patterns of trait variation within and among plant species**

Masatoshi Katabuchi^1,2,6^, Kaoru Kitajima^2,3,4^, S. Joseph Wright^4^, Sunshine A. Van Bael^4,5^, Jeanne L. D. Osnas^2^ and Jeremy W. Lichstein^2^

^1^ Xishuangbanna Tropical Botanical Garden, Chinese Academy of Sciences, Menglun, Yunnan 666303 China

^2^ Department of Biology, University of Florida, Gainesville, FL 32611, USA

^3^ Graduate School of Agriculture, Kyoto University, Kitashirakawa Oiwake-Cho, Kyoto 606-8502 Japan

^4^ Smithsonian Tropical Research Institute, 9100 Panama City Pl., Washington, DC 20521

^5^ Department of Ecology and Evolutionary Biology, Tulane University, New Orleans, LA 70118 USA

^6^ **Corresponding Author**: E-mail: mattocci27@gmail.com


- Across the global flora, photosynthetic and metabolic rates depend more strongly on leaf area than leaf mass.
In contrast, intraspecific variation in these rates is strongly mass-dependent.
These contrasting patterns suggest that the causes of variation in leaf mass per area (LMA) may be fundamentally different within vs. among species.

- We developed statistical modeling framework to decompose LMA into two conceptual components – metabolic LMAm (which determines photosynthetic capacity and dark respiration) and structural LMAs (which determines leaf toughness and potential leaf lifespan) - using leaf trait data from tropical forest sites in Panama and a global leaf-trait database.

- Decomposing LMA into LMAm and LMAs provides improved predictions of leaf trait variation (photosynthesis, respiration, and lifespan).
We show that strong area-dependence of metabolic traits across species can result from multiple of factors, including high LMAs variance and/or a slow increase in photosynthetic capacity with increasing LMAm.
In contrast, strong mass-dependence of metabolic traits within species across light levels results from LMAm increasing from sun to shade.
LMAm and LMAs were nearly independent of each other in both global and Panama datasets.

- Our results suggest that leaf functional variation is multi-dimensional and that biogeochemical models should treat metabolic and structural leaf components separately.

# Introduction

Leaf functional traits play an important role in ecological and physiological tradeoffs [@Wright2004a; @Reich2014; @Onoda2017] and in carbon and nutrient cycling [e.g., @Huntingford2017; @Tcherkez2017].
Thus, understanding the causes and consequences of leaf trait variation is an important goal in ecology, plant physiology, and biogeochemistry [@Bonan2002; @Poorter2009].
Different leaf assemblages exhibit markedly different patterns of trait variation.
For example, across global species, whole-leaf values of traits related to photosynthesis and metabolism (e.g., the photosynthetic capacity, respiration rate, or nutrient content of entire leaves) tend to increase more strongly with leaf area than leaf mass (@Osnas2013),
whereas the opposite pattern (strong mass-dependence of these same traits) is observed within species across canopy light gradients [@Osnas2018; @Osada2001].
Functional groups (e.g., deciduous vs. evergreen angiosperms) also differ from each other in terms of how photosynthetic and metabolic traits variation depend on leaf mass and area [@Osnas2018].
These divergent patterns suggest the presence of multiple drivers of trait variation.

Strong interspecific correlations among leaf mass per area (LMA), leaf lifespan (LL), and mass-normalized leaf traits related to photosynthesis and metabolism (hereafter, 'metabolic traits') have been interpreted as evidence for a single dominant axis of leaf functional variation [@Wright2004a].
However, the interpretation of these strong correlations, which result from mass-normalization of area-dependent traits [@Lloyd2013; @Osnas2013], is controversial.
On the one hand, mass-normalization has been justified based on economic principles alone [@Westoby2013], because leaf mass provides a simple index of investment that underlies the leaf economics spectrum (LES), ranging from cheap, low-LMA leaves with a fast rate of return per-unit investment to expensive, high-LMA leaves with a slow rate of return [@Wright2004a].
On the other hand, because the lifetime return on investment depends on LL [@Westoby2000; @Falster2012], and normalizing traits by their annualized construction costs (or mass/LL) should be used as a normalizer in leaf economics analyses [@Osnas2013].
In global analyses, normalizing leaf traits by mass/LL yields similarly weak correlations as area-normalization, because leaf area is roughly proportional to mass/LL across global species [@Osnas2013].
Area-dependence of metabolic traits and the inconsistent correlation strengths obtained from different normalizers [@Lloyd2013; @Osnas2013] do not invalidate mass-normalization or the LES [@Westoby2013]
These considerations do, however, lead us to question the evidence for a single dominant axis of leaf functional diversity, which has important implications for how trait variation is represented in ecosystem models [@Bonan2002; @Sakschewski2016].

To understand why the same data can be interpreted as either supporting or opposing the existence of a single dominant axis of leaf trait variation, consider the conceptual model proposed by @Osnas2018, in which LMA is comprised of two additive components:
photosynthetic LMA (LMAm) -- the mass per area of chloroplasts and other metabolically active leaf components that contribute directly to photosynthesis and respiration -- and structural LMA (LMAs) -- the mass per area of structural leaf components that contribute to toughness and durability [@Kitajima2012; @Kitajima2016; @Onoda2017].
Suppose these two LMA components are independent axes of functional variation, so that LMAm and LMAs are uncorrelated across species ([@fig-Hplt]a).
These two independent axes can be translated into either a two-dimensional trait space (if metabolic traits are area-normalized; [@fig-Hplt]b) or a one-dimensional trait space (if metabolic traits are mass-normalized; [@fig-Hplt]c).
While both [@fig-Hplt]b and [@fig-Hplt]c are both ‘correct’ representations of the same data, they lead to different perceptions about the dimensionality of leaf functional variation.
If LMAm, LMAs, and/or other axes are largely independent and have distinct functional consequences, then it would not be possible to accurately represent functional variation with a single axis.

The hypothetical example in [@fig-Hplt] shows how mass-normalization can, in principle, make a two-dimensional trait space appear one-dimensional, but the dimensionality of functional variation in real leaf assemblages remains an open question.
One way to better understand the dimensionality of leaf trait variation is to compare models with different numbers of dimensions, and to ask if models with multiple dimensions provide improved statistical fits and conceptual insights compared to a single axis.
For example, the two-dimensional ‘LMAm + LMAs’ model proposed by @Osnas2018 could be compared to a one-dimensional LMA model in terms of their capacities to explain variation in other traits.
However, implementing the ‘LMAm + LMAs’ model is challenging, because although certain leaf mass components can be neatly classified as ‘metabolic’ or ‘structural’ [@Poorter2009; @Osnas2018], other leaf mass components cannot.
For example, thick cell walls contribute to structural toughness [@Onoda2015], but at least some cell wall mass is required for the biomechanical support that enables photosynthesis.
Thus, partitioning LMA into different functional components requires novel empirical or modeling approaches.

In this paper, we present a statistical modeling framework to partition LMA into metabolic and structural components: LMAm and LMAs.
We develop and test this two-dimensional model using leaf trait data from two tropical forest sites (sun and shade leaves from wet and dry sites in Panama) and the GLOPNET global leaf traits database [@Wright2004a].
We use the model to address the following questions:
(1) Are measured leaf traits (including photosynthetic capacity, dark respiration rate, LL, and concentrations of nutrients and cellulose) better predicted by a one-dimensional (total LMA) or two-dimensional (LMAm-LMAs) model?
(2) What are the relative contributions of LMAm and LMAs to total LMA variance in different leaf assemblages (Panama sun leaves, Panama shade leaves, and the global flora)?
(3) Do LMAm and LMAs differ between evergreen and deciduous species, and between sun and shade leaves? If so, how? and
(4) How do the answers to the preceding questions inform our understanding of empirical patterns of trait variation (e.g., relationships among measured leaf traits)?

# Material and Methods

## Overview

We considered multiple approaches to modeling datasets that include observations of leaf mass per area (LMA), leaf lifespan (LL), net photosynthetic capacity (*A*), and dark respiration rate (*R*).
These traits comprise four of the six traits in the global leaf economics spectrum (LES) analysis of @Wright2004a.
For simplicity, we did not include the other two LES traits – leaf nitrogen (*N*) and phosphorus (*P*) concentrations – in our modeling framework.
Instead, we reserved observations of leaf *N* and *P* concentrations (and cellulose content, when available) for independent model tests.
We considered a simple one-dimensional model that predicts LL, *A*, and *R* from LMA alone, as well as two-dimensional models that predict LL, *A*, and *R* from two additive LMA components: structural and metabolic and structural leaf mass per area (LMAm and LMAs, respectively).
We formulate the models in terms of area-normalized *A* and *R* (*A*~area~ and *R*~area~, respectively), as these trait forms share the same denominator as LMA and its components.

To implement the two-dimensional models, we developed a statistical modeling framework to partition LMA into additive LMAm and LMAs components.
We fit the models to two datasets: the GLOPNET global leaf traits dataset [@Wright2004a], which primarily represents interspecific variation; and the Panama dataset described by @Osnas2018, which includes traits for both sun and shade leaves at wet and dry tropical forest sites.
Because LMAm and LMAs are modeled (rather than observed), the two-dimensional models require one parameter per analysis unit (a species in GLOPNET or a species $\times$ canopy-position in the Panama dataset) to partition observed LMA into LMAm and LMAs.
Given the large number of parameters in these models, we performed tests with randomized data to evaluate if our two-dimensional models were prone to overfitting, which could lead to spurious conclusions.
These tests suggested that our two-dimensional modeling approach revealed meaningful patterns in the observed trait data.

## Modeling leaf lifespan, photosynthetic capacity, and dark respiration rate

We considered five types of models, ranging from simple models with LMA as the sole predictor, to more complex models in which LMA was partitioned into LMAm and LMAs.
In all models, the unit of analysis is a ‘leaf sample’, defined as a species in the GLOPNET dataset or a species $\times$ canopy position in the Panama dataset (the datasets are described below).

First, we considered a simple set of models with LMA as the sole predictor for *A*~area~, *R*~area~, and LL according to power-law relationships.

Next, we considered models in which LMA is partitioned into additive metabolic and structural components for leaf sample *i*:

$$
\begin{aligned}
  &\mathrm{LMA}_{i} =\mathrm{LMAm}_{i} + \mathrm{LMAs}_{i} \\
  &\mathrm{LMAm}_{i} = f_{i} \mathrm{LMA}_{i} \\
  &\mathrm{LMAs}_{i} = (1 - f_{i})  \mathrm{LMA}_{i}
\end{aligned}
$$ {#eq-LMA}

where the *f~i~* values -- the fractions of LMA comprised of LMAm for each sample *i* --  are estimated as latent variables in our modeling framework (see details below).
We assumed that the observed values of *A*~area~, *R*~area~ and LL were related to the unobserved values LMAm and LMAs according to multivariate power-laws:

\begin{align}
& \mathrm{E}[A_{\mathrm{area} \, i}]
= \alpha_0\mathrm{LMAm}_{i}^{\alpha_m}\mathrm{LMAs}_i^{\alpha_s}  =  \alpha_0 (f_i \mathrm{LMA}_{i})^{\alpha_m} \bigl\{(1-f_i) \mathrm{LMA}_{i}\bigr\}^{\alpha_s} \label{eq:Aarea} \\
& \mathrm{E}[R_{\mathrm{area} \, i}]
= \gamma_0\mathrm{LMAm}_{i}^{\gamma_p} \mathrm{LMAs}_{i}^{\gamma_s}
= \gamma_0 (f_i \mathrm{LMA}_{i})^{\gamma_p} \bigl\{(1-f_i)\mathrm{LMA}_{i}\bigr\}^{\gamma_s} \label{eq:Rarea} \\
& \mathrm{E}[\mathrm{LL}_i] = \beta_0\mathrm{LMAm}_{i}^{\beta_m} \mathrm{LMAs}_{i}^{\beta_s}  = \beta_0 (f_i \mathrm{LMA}_{i})^{\beta_m} \bigl\{(1-f_i) \mathrm{LMA}_{i}\bigr\}^{\beta_s} \label{eq:LL_pot} \tag{4a}  \\
& \mathrm{E[LL}_i] = \beta_0\mathrm{LMAm}_{i}^{\beta_m} \mathrm{LMAs}_{i}^{\beta_s} \mathrm{exp}(\theta \mathrm{Light}_i)  \stepcounter{equation} \label{eq:LL_opt} \tag{4b}
\end{align}

where E[$\cdot$] indicates expected value;
the $\alpha$, $\beta$, and $\gamma$ terms are fitted parameters;
and the logarithms of *A*~area~, LL, and *R*~area~ are assumed to have a multivariate normal distribution (Appendix S1).
In preliminary analyses, we also considered alternative (non-power-law) forms, but none of these performed better than the power-law forms and are not discussed further.

To ensure that the model was identifiable, we imposed two broad assumptions:
(i) *A*~area~ depends more strongly on metabolic leaf mass (LMAm: parameter $\alpha_m$ in \ref{eq:Aarea}) than structural leaf mass (LMAs: parameter $\alpha_s$), and
(ii) LL depends more strongly on LMAs ($\beta_s$ in Eqs.\ref{eq:LL_pot}-\ref{eq:LL_opt}) than LMAm ($\beta_m$).
The first assumption was implemented in different model versions either by setting $\alpha_s$ = 0 or by imposing the constraint $\alpha_m$ > $\alpha_s$.
Similarly, the second assumption was implemented either by setting $\beta_m$ = 0 or by imposing the constraint  $\beta_s$ > $\beta_m$.
The weaker form of these assumptions ($\alpha_m$ > $\alpha_s$ and $\beta_s$ > $\beta_m$) is primarily a labeling constraint and only weakly constrains the possible biological model outcomes by excluding the possibility that a single LMA component could be the primary determinant of both *A*~area~ and LL.
The stronger form of the assumptions ($\alpha_s$ = 0 and $\beta_m$ = 0) leads to a more parsimonious model (fewer parameters).
We considered different combinations of the strong and weak forms of the assumptions for *A*~area~ ($\alpha_m$ and $\alpha_s$) and LL ($\beta_s$ and $\beta_m$) using cross-validation (see details below).
We did not impose any constraints on *R*~area~ ($\gamma_s$ and $\gamma_m$).

We also considered models in which LMAs in Eqs. \ref{eq:LL_pot}-\ref{eq:LL_opt} was replaced by leaf structural density (LMAs/LT, where LT is leaf thickness), which is based on the observation that cellulose density is a good predictor for LL in some species assemblages [@Kitajima2012; @Kitajima2013].
These models with LMAs/LT yielded qualitatively similar results as those based on LMAs, but did not perform as well in the model evaluation (see below).
Therefore, we only present results from the models with LMAs.

Finally, we considered a functional form that accounts for the effects of light availability on LL.
This light-dependent model is motivated by optimal LL theory, which predicts decreasing LL with increasing light availability [@Kikuzawa1991], and also by the often-observed ‘LMA counter-gradient’, whereby LL and LMA positively covary across species but negatively covary within species across light gradients [@Lusk2008; @Russo2016; @Osnas2018].
Mechanistically modeling how light affects LL via leaf carbon balance [@Xu2017] is beyond the scope of our study.
Instead, we introduced light effects in a simple way by modifying Eq. \ref{eq:LL_pot} to Eq. \ref{eq:LL_opt} where the dummy variable Light~*i*~ is set to 1 for sun leaves and 0 for shade leaves, and exp($\theta$) is the sun:shade LL ratio.

## Datasets

We fit the models described above using observations of LMA (g m^-2^), *A*~area~ (mol s^-1^ m^-2^), *R*~area~ (mol s^-1^ m^-2^) and LL (months) from the GLOPNET global leaf traits database [@Wright2004a] and from two tropical forest sites in Panama: Monumental Natural Metropolitano (MNM, “dry site”) and Bosque Protector San Lorenzo (SL, “wet site”).
The GLOPNET data primarily represents interspecific variation (the dataset reports 2,548 species $\times$ site combinations, with 2,021 unique species and only 350 species occurring at more than one site) and only reports data for sun leaves if data for both sun and shade leaves are available [@Wright2004a].
In contrast, the Panama dataset represents both inter- and intraspecific variation, including leaves sampled at two canopy positions (“sun”: full sun at the top of the canopy; and “shade”: well shaded, sampled within 2 m of the forest floor) from trees within reach of  canopy cranes.
The dry MNM site is a semi-deciduous coastal Pacific forest with a 5-month dry season from December-April and 1740 mm of annual rainfall [@Wright2003].
The MNM crane is 40 m tall with a 51 m long boom. The wet SL site is an evergreen Caribbean coastal forest with 3100 mm of annual rainfall [@Wright2003].
The SL crane is 52 m tall with a 54 m long boom.
Additional details of the Panama dataset are described in @Osnas2018.

We restricted our analysis to database records for which all four traits (LMA, *A*~area~, *R*~area~, and LL; each typically averaged over multiple leaves) were available.
Each database record corresponds to an analysis unit (or ‘leaf sample’) as described above; i.e., a species in GLOPNET, or a species $\times$ canopy position in the Panama dataset.
After excluding database records that lacked one or more of the four traits, 198 samples for 198 unique species were available for GLOPNET, and 130 samples for 104 unique species were available for Panama (dry and wet sites combined; 26 species sampled in both sun and shade; no species with all four traits available at both sites).
In addition to LMA, *A*~area~, *R*~area~, and LL, the model based on structural leaf density also requires observations of leaf thickness, which was not available in GLOPNET but was available for 106/130 Panama samples.
Both the GLOPNET and Panama datasets include additional traits that we used to interpret model results, but which were not used to fit models. These traits include nitrogen and phosphorus content per leaf unit area (*N*~area~ and *P*~area~; g m^-2^) and leaf habit(deciduous or evergreen), and cellulose content per unit area (*CL*~area~; g m^-2^) in the Panama dataset.

## Model estimation and evaluation

We modeled *A*~area~ and *R*~area~ using Eqs. \ref{eq:Aarea} and \ref{eq:Rarea}, respectively, for both GLOPNET and Panama.
To model LL for GLOPNET, we used Eq. \ref{eq:LL_pot} (no light effects), because GLOPNET does not report canopy position (and prioritizes data for sun leaves, as described above).
To model LL for Panama, we used Eq. \ref{eq:LL_opt} (light effects model), which was motivated by the negative intraspecific LL-LMA relationship observed in Panama [@Xu2017; @Osnas2018] and elsewhere [@Lusk2008; @Russo2016].

Posterior distributions of all parameters were estimated using the Hamiltonian Monte Carlo algorithm (HMC) implemented in Stan [@Carpenter2017].
We used non-informative or weakly informative prior distributions [@Lemoine2019].
Prior distributions for the latent variables *f~i~* (which are used to partition LMA into LMAm and LMAs according to @eq-LMA were non-informative uniform(0, 1) distributions, (i.e., LMA was partitioned based on patterns in the data).
See Appendix S1 for more detail.
The Stan code use to fit models is available from Github at: [https://github.com/mattocci27/LMAms](https://github.com/mattocci27/LMAms).
Convergence of the posterior distribution was assessed with the Gelman-Rubin statistic with a convergence threshold of 1.05 for all parameters [@Gelman2013].

Because our two-dimensional (LMAm-LMAs) modeling approach includes many parameters (one latent variable *f~i~* to partition LMA into LMAm and LMAs for each leaf sample), we implemented tests with randomized data to assess potential overfitting.
We generated 10 different randomized datasets by randomizing all trait values (LMA, *A*~area~, *R*~area~ and LL) across species.
Thus, the randomized datasets had zero expected covariance among traits.
Models fit to the randomized datasets either did not converge or showed divergent transitions (Appendix S2), which do not provide reliable inferences [@Betancourt2016].
In simple terms, the two-dimensional models failed when fit to randomized data.
In contrast, when fit to the observed (non-randomized) data, the two-dimensional models converged without divergent transitions and out-performed one-dimensional models (total LMA) for both GLOPNET and Panama (see Results).
Thus, the tests with randomized data indicate that our two-dimensional approach is not inherently prone to overfitting or to creating spurious results.
We therefore assume that estimates of LMAm and LMAs obtained from the GLOPNET and Panama datasets reflect meaningful patterns in the observations and allow for a meaningful exploration of our questions.

We used Pareto-smoothed importance sampling leave-one-out cross-validation (PSIS-LOO; @Vehtari2017) to compare the performance of different models.
PSIS-LOO is an accurate and reliable approximation to standard leave-one-out cross-validation (LOO), which is a robust method for comparing models with different numbers of parameters [@Vehtari2017].
LOO requires *n* model fits (for a dataset with *n* observations) and was therefore impractical in our study due to our computationally-demanding modeling approach.
We used PSIS-LOO to calculate the LOO Information Criterion (LOOIC) for each model.
A lower LOOIC indicates a better model in terms of predictive accuracy [@Vehtari2017].


## Understanding relationships between photosynthetic capacity and LMA

We applied our LMAm-LMAs model to simulated data to better understand relationships between photosynthetic capacity (*A*) and LMA.
The causes and interpretation of relationships between *A* and LMA are controversial [@Westoby2013].
Although *A* is often mass-normalized [e.g., @Wright2004a; @Shipley2006; @Blonder2011], it has been argued that *A* should be area-normalized when exploring trait relationships, because photosynthesis is an area-based process [@Lloyd2013].
Consistent with this argument, @Osnas2013 showed that across global species, variation in whole-leaf *A* is strongly dependent on leaf area, but only weakly dependent on leaf mass (after controlling for variation in leaf area).
@Osnas2018 further showed that the relationship between *A* and LMA (i.e., the degree of mass- vs. area-dependence) is sensitive to the amount of LL variation in an assemblage, which we hypothesize depends on the fraction of total LMA variance in the assemblage that is due to LMAs variance.

To better understand the factors affecting *A* vs. LMA relationships, we created simulated datasets in which we varied the following factors: the sensitivity of *A*~area~ to variation in LMAm (parameter $\alpha_m$ in Eq. \ref{eq:Aarea}); the sensitivity of *A*~area~ to LMAs (parameter $\alpha_s$ in Eq. \ref{eq:Aarea}), and the total LMA variance due to variance in LMAs.
For each simulated dataset, we quantified the *A* vs. LMA relationship following @Osnas2018:

$$
A_{\mathrm{area} \, i} = c (LMA_i)^{b}\epsilon_i
$$ {#eq-mass}

where LMA is the sum of LMAm and LMAs (@eq-LMA), *c* is a fitted constant, and *b* is an index of mass-dependence as illustrated by the following cases [@Osnas2013; @Osnas2018]: if *b* = 0, then *A*~area~ is independent of LMA, which implies that whole-leaf A is proportional to leaf area;
conversely, if *b* = 1, then *A*~area~ is proportional to LMA, which implies that whole-leaf A is proportional to leaf mass.
Intermediate cases (0 < *b* < 1), as well as more extreme cases (*b* $\le$ 0 or *b* $\geq$ 1), are also possible, but empirical estimates tend to be between 0 and 1 [@Osnas2018].
Note that applying @eq-mass to either mass- or area-normalized A yields equivalent results [@Osnas2018].

For these simulations, we generated LMAm and LMAs values based on their estimated distributions in our GLOPNET and Panama analyses, while varying the variance of LMAs (so that LMAs contributed different fractions of total LMA variance in different simulated datasets).
Although Panama sun and shade leaves were analyzed together when fitting the LMAm-LMAs models Eqs. \ref{eq:Aarea} and \ref{eq:LL_opt}, they were analyzed separately here due to their different estimated covariance structures (see details below).

For GLOPNET and Panama sun leaves, LMAm and LMAs were generated independently of each other, consistent with the low estimated correlations between LMAm and LMAs in these assemblages (see Results).
Specifically, LMAm values were randomly sampled from a lognormal distribution with the log-scale mean and standard deviation given by our model outputs; and LMAs values were randomly sampled from a lognormal distribution with the log-scale mean given by our model outputs and the log-scale standard deviation varying between log(1.01) and log(10) in different simulated datasets.
In contrast to GLOPNET and Panama sun leaves, our analysis revealed a negative correlation of r = −0.4 between LMAm and LMAs for Panama shade leaves (see Results).
Therefore, we used a multivariate normal distribution with r = −0.4 to generate log(LMAm) and log(LMAs) for the Panama shade simulations.
Otherwise, the simulation protocol was as described above for GLOPNET and Panama sun leaves.

For each simulated set of LMAm and LMAs values, we generated *A*~area~ values using the estimated values of $\alpha_m$ and $\alpha_s$ (Eq. \ref{eq:Aarea}) from the corresponding best-fitting model (GLOPNET or Panama); i.e., the model with the lowest LOOIC.
We generated 1000 simulated datasets for GLOPNET, Panama sun leaves, and Panama shade leaves, with each dataset having a sample size of 100 leaves.
For each simulated dataset, we used @eq-mass to quantify the relationship between *A*~area~ and LMA.

## Variance partitioning

To estimate the contributions of LMAm and LMAs to total LMA variance (where LMA = LMAm + LMAs), we used the following identity:

$$
\mathrm{Var}(Y = X1 + X2) = \mathrm{Cov}(Y, X1+X2) = \mathrm{Cov}(Y,X1) + \mathrm{Cov}(Y,X2)
$$ {#eq-var}

where Var($\cdot$) is variance and Cov($\cdot$) is covariance.
Thus, the fractions of total LMA variance due to variance in LMAm and LMAs are Cov(LMA, LMAm)/Var(LMA) and Cov(LMA, LMAs)/Var(LMA), respectively.
We applied this method to the simulated datasets described above and also to estimates of LMAm and LMAs for GLOPNET and Panama sun and shade leaves.

# Results

**1. Decomposing LMA into metabolic and structural components leads to improved predictions of *A*~area~, *R*~area~ and LL.**
For both the GLOPNET global datasets and the Panama datasets, two-dimensional model (LMAm and LMAs) performed better in cross validation than one-dimensional (LMA) models (Table 1), suggesting that decomposing LMA into metabolic and structural components leads to improved predictions of *A*~area~, *R*~area~ and LL.
Additionally, the effects of LMAm on LL for the GLOPNET and the Panama dataset and the effect of LMAs on the Panama dataset were not selected in the best models (Table 1).

For the GLOPNET global dataset, *A*~area~ had a positive correlation with LMAm, a negative correlation with LMAs, and a non-significant correlation with total LMA ([@fig-GLplt]a-c; Table 2).
*R*~area~ in GLOPNET also had a positive correlation with LMAm, which was stronger than the correlation between *R*~area~ and either LMA or LMAs ([@fig-GLplt]d-f).
Finally, LL in GLOPNET had a strong positive correlation with LMAs, which was stronger than the correlation between LL and LMA ([@fig-GLplt]g-i).

For the Panama dataset, including the effect of light (sun vs. shade) imporved model performance (Table 1).
All Panama results we report are for models including light effects unless stated otherwise.
*A*~area~ and *R*~area~ had stronger and more positive correlations with LMAm than with LMA ([@fig-PAplt]a-f and Table 2).
LL was not significantly correlated with LMA when all leaves were combined, but was strongly correlated with LMA for shade leaves at the dry site ([@fig-PAplt]g).
There was a positive correlation between LL and LMAs ([@fig-PAplt]i), and this relationship was strengthened by accounting the effects of LL (@fig-LLplt).

**2. Nearly all leaf dark respiration is associated with metabolic leaf tissue mass.**
According to our model results, metabolic leaf mass (LMAm) accounted for nearly all leaf dark respiration; i.e., estimated dark respiration rate per-unit structural mass ($\gamma_s$) was close to zero in analyses of both GLOPNET and Panama data (Table 1).
Thus, although building costs are likely similar for different leaf chemical components and tissues [@Williams1989; @Villar2001], our results suggest that leaf mass associated with photosynthetic function accounts for nearly all leaf maintenance respiration.

**3. Evergreen leaves have greater LMAs than deciduous leaves, and sun leaves have both greater LMAm and LMAs than shade leaves.**
Evergreen vs. decidous explains the large variance
(`r  get_vpart_para(here("data/vpart_intra.csv"), "LMAs", "Eve/Dec", "gl_de", digits = 1, nsmall = 1)`%)
in LMAs but not LMAm in GLOPNET ([@fig-vpart]).
Thus, the higher total LMA in evergreen leaves in GLOPNET ([@fig-boxplt_de]) was primarily due to differences in LMAs.
In the Panama dataset, light (sun vs. shade) explains the majority of variance in LMAm
(`r  get_vpart_para(here("data/vpart_intra.csv"), "LMAm", "Sun/Shade", "pa_leaf", digits = 1, nsmall = 1)`%)
but not LMAs
(`r  get_vpart_para(here("data/vpart_intra.csv"), "LMAs", "Sun/Shade", "pa_leaf", digits = 1, nsmall = 1)`%)
, and site (wet/dry) explains
`r  get_vpart_para(here("data/vpart_intra.csv"), "LMAs", "Wet/Dry", "pa_leaf", digits = 1, nsmall = 1)`%
of variance in LMAs ([@fig-vpart]).
Sun leaves have both greater LMAm and LMAs than shade leaves, but the differences in LMAm were more apparent ([@fig-boxplt_pa]).
Thus, variation in LMAm plays a more important role within species than among species.
Evergreen leaves have higher fraction of LMAs compared to deciduous leaves both in the Panama and the GLOPNET dataset (Fig. S\ref{fig-box_frac}).

**4. LMA variance components determine the area- vs. mass-dependence of leaf photosynthetic capacity (*A*).**
Analysis of simulated data showed that the percent of LMA variation due to LMAm vs. LMAs variation determined the area- vs. mass-dependence of *A* (i.e., the degree to which whole-leaf *A* depends on leaf area or mass; @Osnas2013; @Osnas2018).
Generally, mass-dependence of *A* decreases with LMAs variance ([@fig-massplt]).
Other factors, including the senstivity of *A* to variation in LMAm and LMAs ($\alpha_m$ and $\alpha_s$) and covariance between LMAm and LMAs, also affect the relationship between mass-dependence of *A* and LMAs variance ([@fig-massplt], Fig. S\ref{fig-mass_prop_sim}-\ref{fig-mass_prop_comp}).
Although relative variance of LMAs was intermediate in GLOPNET (`r var_val$GL`%), interspecific variation in *A* among leaves in GLOPNET tended to be area-dependent because of their lower $\alpha_m$ and negative $\alpha_s$.
Interspecific variation in *A* in sun and shade leaves in the Panama data was area-dependent because of their large variance in LMAs ([@fig-massplt]).

**5. Nitrogen and phosphorus per-unit leaf area are strongly correlated with LMAm, and cellulose per-unit leaf area is strongly correlated with LMAs.**
In the GLOPNET dataset, *N*~area~ and *P*~area~ had strong positive correlations with LMAm, but only weak correlations with LMAs (Fig. S\ref{fig-gl_point_np2}).
Similarly, in the Panama dataset, *N*~area~ and *P*~area~ had strong positive correlations with LMAm, but were not correlated with LMAs ([@fig-PA-NPC]).
In contrast, cellulose per-unit leaf area (*CL*~area~), which was available for the Panama dataset but not for GLOPNET, had a strong positive correlation with LMAs, and a weak positive correlation with LMAm ([@fig-PA-NPC]).
*CL*~area~ was more strongly correlated with LMA than with LMAm or LMAs, but sun and shade leaves aligned along a common *CL*~area~-LMAs relationship, as opposed to being offset for LMA and LMAm ([@fig-PA-NPC]g-i).

# Discussion

Our analyses demonstrate that decomposing LMA variation into separate metabolic and structural components (LMAm and LMAs, respectively) leads to improved predictions of photosynthetic capacity (*A*), dark respiration rate (*R*), and leaf lifespan (LL), as well as clear relationships with traits used for independent model evaluation (nitrogen, phosphorus, and cellulose concentrations).
The model with LMAm and LMAs showed better predictive accuracy than model with LMA alone for both intraspecific variation in relation to light and interspecific variation.
Below, we elaborate on the insights gained from our analysis and the implications of our results for the representation of leaf functional diversity in global ecosystem models.

Decomposing LMA into LMAm and LMAs provides insights into why interspecific variation in leaf traits related to photosynthesis and metabolism are primarily area-dependent (i.e., primarily independent of LMA when expressed per-unit area rather than mass-dependent [@Osnas2013; @Osnas2018]).
We expected that leaf metabolic traits tend to be area-depend when variance in LMAm is larger than variance in LMAs.
However, our results suggest that the scaling slopes ($\alpha_m$ and $\alpha_s$) between LMAm, LMAs and *A*~area~ also affect trait area proportionality.
Leaf metabolic traits tend to be area-dependent when (i) variance in LMAm is small or similar to variance in LMAs, (ii) the effects of LMAm and LMAs on *A*~area~ ($\alpha_m$ and $\alpha_s$) are small, (iii) or combinations of those (Fig. [@fig-massplt]).
Consistent with this explanation, the assemblage we examined where LMAs counted for the highest fraction of total LMA variation (`r var_val$shade`% for Panama shade leaves) is also the assembled with the highest degree of trait area-dependence [@Osnas2018].
Our results also suggest that high degree of trait area-dependence in GLOPNET is due to the negative effect of LMAs on *A*~area~.

The scaling slope $\alpha_m$ less than 1 (Table 2) suggests a diminishing return on addition LMAm (i.e., doubling LMAm does not simply result in doubling *A*).

1. High LMAm might cause lower mesophyll conductance
1. High LMAm (thick leaves) might cause chloroplast self-shading within a leaf
1. LMAm = (purely) metabolic active leaf mass + something we couldn't estimate in our analysis. All the leaf mass needs to be allocated to LMAm or LMAs in our model. However, variation in non-structural carbohydrates in leaves, which is not estimated in our model, is relatively large [@Martinez-Vilalta2016]. Simply partitioning leaf mass into two functions might underestimate how much *A* depending on the metabolic leaf mass .

Decomposing LMA also provides insights as to why intraspecific patterns of trait variation differ from those observed across species.
In contrast to interspecific LMA variation, our analysis suggests that LMAm contributes half or more of the intraspecific increase in LMA from shade to sun (Figs. [@fig-massplt] and SX).
The increase in LMAm from shade to sun – which likely reflects an increase in the size and number of palisade mesophyll cells with increasing light availability [@Onoda2008; @Terashima2011] – is also associated with an increase in LMAs from shade to sun (Fig. [@fig-boxplt_pa]).
This positive covariance between LMAm and LMAs within species means that per-area values of LMAm-proportional traits (e.g., *A*~area~) have a strong, positive relationship with total LMA, which implies trait mass-proportionality [@Osnas2018].

The improved predictions and understanding provided by decomposing LMA into photosynthetic and structural components challenge the view that leaf functional diversity can be accurately represented by a single leaf economics spectrum (LES) axis [@Wright2004a].
@Lloyd2013 argued that the apparent dominance of a single LES axis is an artifact of expressing area-dependent leaf traits on a per-mass basis, and @Osnas2013 demonstrated that across the global flora, traits related to photosynthesis and metabolism are indeed area-dependent.
Intraspecific patterns in trait variation, which contrast with interspecific patterns, pose additional challenges for a one-dimensional view of leaf functional diversity. Our analysis shows that considering two primary axes of leaf trait variation (photosynthesis and structure) provides improved quantitative predictions and insights compared to LES of a single dominant axis.
Our results point to a simple two-dimensional framework for representing leaf functional diversity in global ecosystem models: a LMAm axis that determines *A*, accounts for nearly all *R* (see rp and rs estimates in Table S1) and a LMAs axis that determines potential LL through its effects on leaf toughness [@Kleyer2012].
In the datasets we analyzed (the global flora and tropical trees in Panama), these two axes are only weakly correlated with each other (Fig. S\ref{fig-LMAm_LMAs}), which suggests that trait-based approaches to global ecosystem modeling [@Scheiter2013; @Wullschleger2014] could consider these as independent axes.
For example, to simulate individual competing trees which from diverse community of growth strategies [@Sakschewski2015; @Sakschewski2016], it would be possible to draw LMAm and LMAs values from distributions with realistic ranges instead of a single LMA distribution.

Our statistical decomposition of LMA into metabolic and structural components provides important insights, but additional insights and accuracy could be gained by a more mechanistic modeling approach.
For example, @John2017 decomposed interspecific LMA variation into anatomical components such as the size, number of layers, and mass density of cells in different leaf tissues.
If such detailed information became available for a large number of leaves, representing both intra- and interspecific variation, it should be possible to quantify how these anatomical traits scale up to leaf-level *A* , *R*, and LL.
A simpler alternative would be to modify our model to account for variation in cell wall thickness (TCW): for a given cell size, increasing TCW would lead to an increase in lamina density, cellulose per volume, toughness, and LL [@Kitajima2010; @Kitajima2012; @Kitajima2016], and a decrease in mesophyll conductance and *A* [@Evans2009; @Terashima2011; @Onoda2017].
Indeed, our model might capture those patterns. The negative effects of LMAs on *A*~area~ in GLOPNET data indicate that high amount of photosynthetic proteins is not always translated into high photosynthetic capacity.
High LMAs that may be associated with a particular anatomical components might lead to lower mesophyll conductance.


# Conclusions

It is widely recognized that LMA variation is associated with multiple tissues and functions, including metabolically active mesophyll that largely determines photosynthetic capacity, as well as structural and chemical components that contribute primarily to leaf toughness and defense [@Roderick1999; @Shipley2006; @Lusk2010].
It should not be surprising then, that partitioning LMA into metabolic and structural components yields enhanced predictions and improved understanding of patterns of leaf trait variation both within and among species.
Yet for over a decade, the vast literature on leaf traits has been strongly influenced by the view that leaf trait variation can usefully be represented by a single dominant axis of LES.
Our results provide quantitative evidence that this one-dimensional view of leaf trait variation is insufficient, and our model provides a biological explanation for previous statistical analyses that have demonstrated area-dependence of leaf traits across species [@Osnas2013], while also explaining mass-dependence within species. Our results suggest that small scaling slope between metabolic leaf mass and photosynthetic rate, and similar variance in LMAm and in LMAs leads to area-proportionality of interspecific leaf traits.
Thus, strong interspecific correlations between LMA and mass-normalized photosynthetic capacity (and related traits, such as respiration rate, and nitrogen and phosphorus concentrations) are likely driven by mass-normalization itself, rather than any functional dependence of these photosynthetic and metabolic traits on LMA [@Lloyd2013; @Osnas2013].
In contrast, intraspecific variation in LMA is driven by coordinated changes in structural and metabolic mass components, which explains why mass-normalized photosynthetic and metabolic traits vary little from sun to shade within species [@Aranda2004; @Poorter2006b; @Niinemets2015].


# Acknowledgments

We thank Jonathan Dushoff for statistical advice and Martijn Slot for helpful comments that improved the paper.
Mirna Samaniego and Milton Garica provided indispensable assistance in data collection.
We thank the Smithsonian Tropical Research Institute (STRI), the Tropical Canopy Biology Program at STRI and the Andrew W. Mellon Foundation for supporting this work. MK was supported by a Postdoctoral Fellowship for Research Abroad from the Japan Society for the Promotion of Science and the CAS President’s International Fellowship for Young Staff.

# Author contributions

MK, JWL and JLDO conceived of the study; KK, SJW and SAVB contributed data; MK devised the analytical approach and performed analyses; MK and JWL wrote the first draft of the manuscript, and all authors contributed to revisions.

# References

<div id="refs"></div>

\newpage

# Tables

Table 1. Summary of the model comparison.
Lower approximate leave-one-out cross-validation information criterion (LOOIC) indicates better predictive accuracy.
The models here are the five moodel forms representing LL:
(1) LMA,
(2) LMAm and LMAs,
(3) LMA and light,
(4) LMAm, LMAs and light, and
(5) LMAm, LMAs/LT and light, with four types of parameter restrictions:
(a) $\alpha_s = 0$  and $\beta_m = 0$,
(b) $\beta_m = 0$,
(c) $\alpha_s = 0$, and
(d) $\alpha_m > \alpha_s$  and $\beta_s > \beta_m$.

```{r lootab, eval=TRUE, echo=FALSE}
loo_tbl <- read_csv(here("data/model_selection.csv"))  |>
  # arrange(Data) |>
  dplyr::select(Model, Constraints, N, LOOIC)

if ("html" == knitr::opts_knit$get("rmarkdown.pandoc.to")) {
  knitr::kable(loo_tbl, format = "html", escape = FALSE) |>
    kable_classic(full_width = FALSE) |>
    pack_rows("GLOPNET", 1, 5) |>
    pack_rows("Panama", 6, 23)
} else if ("docx" == knitr::opts_knit$get("rmarkdown.pandoc.to")) {
  tmp <- c("**GLOPNET**", rep("", 4), "**Panama**", rep("", 17))
  tmp2 <- cbind(tmp, loo_tbl)
  colnames(tmp2)[1] <- ""
  tmp2 |>
    kable(format = "pipe", escape = FALSE)
} else if ("latex" == knitr::opts_knit$get("rmarkdown.pandoc.to")) {
  loo_tbl |>
    kable(format = "latex", escape = TRUE, booktabs = TRUE) |>
    kable_classic(full_width = FALSE) |>
    pack_rows("GLOPNET", 1, 5) |>
    pack_rows("Panama", 6, 23)
}
```

\newpage

Table 2.
Posterior means [95% credible interval] of parameters for the best models. Model 1b (LMAm and LMAs) and Model 4a (LMAm, LMAs and light) are shown for GLOPNET and Panama data, respectively.
Bold values are significantly different from zero based on the 95% credible interval.

```{r paratab, echo=FALSE, eval=TRUE}
GLPAtab <- read_csv(here("data/para_tbl.csv"))

if ("html" == knitr::opts_knit$get("rmarkdown.pandoc.to")) {
  knitr::kable(GLPAtab, format = "html", escape = FALSE) |>
    kable_classic(full_width = FALSE)
} else if ("docx" == knitr::opts_knit$get("rmarkdown.pandoc.to")) {
  GLPAtab |>
    kable(format = "pipe", escape = FALSE) |>
    kable_classic(full_width = FALSE)
} else if ("latex" == knitr::opts_knit$get("rmarkdown.pandoc.to")) {
  GLPAtab |>
    kable(format = "pipe", escape = FALSE) |>
    kable_classic(full_width = FALSE)
}
```

\newpage

![
Example of a two-dimensional functional space that can result in either a two- or one-dimensional trait space, depending on how metabolic traits are normalized.
(A) Hypothetical independent variation in two leaf mass per area (LMA) components: metabolic leaf mass per area (LMAm, which largely determines per-area values of photosynthesis, respiration, and nutrient concentrations) and structural leaf mass per area (LMAs, which determines leaf toughness but has little effect on metabolic traits).
(B) Two-dimensional relationship between photosynthetic capacity (*A*) per-unit leaf area and total LMA (equal to LMAm + LMAs).
(C) One-dimensional relationship between *A* per-unit leaf mass and total LMA.
LMAm and LMAs are simulated from a hypothetical scenario of lognormal distributions with medians of 80 and standard deviations of exp(0.8) and exp(0.7), and zero covariance.
Variation in *A* values are derived from our analysis of the GLOPNET database. $A_{\mathrm{area} \, i}=1.77 \times \mathrm{LMAm}_i^{0.28}\mathrm{LMAs}^{-0.13}\epsilon_i$, where $\epsilon_i$ is log-normally distributed with log-mean with 0, log-scale parameter with 0.31 (see Methods and Results).
](`r here::here("figs/hypo.png")`){#fig-Hplt}


![
Observed and estimated leaf-trait relationships in the global GLOPNET dataset.
Estimates are from the best GLOPNET model (Table 1).
Leaf life span (LL), net photosynthetic rate per unit leaf area (*A*~area~), and dark respiration rate per unit leaf area (*R*~area~) are plotted against observed LMA, posterior means of LMAm and LMAs.
Pearson correlation coefficients (*r*) for LMA (left column) and posterior means of Pearson correlation coefficients ($\bar{r}$) or partial correlation coefficients ($\bar{\rho}$) for LMAm (middle column) and LMAs (right column) are shown.
Note that LL was modeled by LMAs alone due to improved model performance with this parameter constraint (Table 1).
Therefore, *r* is reported for LL (single predictor variable), whereas ($\bar{\rho}$)  is reported for *A*~area~ and *R*~area~ (multiple predictor variables).
](`r here::here("figs/gl_point.png")`){#fig-GLplt}

\newpage

![
Observed and estimated leaf-trait relationships in the Panama dataset.
Estimates are from the best Panama model (Table 1), which included effects of light on LL.
Details as for [@fig-GLplt].
Results for other LL models are summarized in Table SX.
Note that *A*~area~ and LL were modeled by LMAp and LMAs alone, respectively, due to improved model performance with these parameter constraints (Table 1).
The results shown here include all leaves in the Panama dataset.
The observed separation of LL between sun and shade leaves is accounted for in the model predictions [@fig-LLplt].
](`r here::here("figs/pa_point.png")`){#fig-PAplt}

\newpage

![
Relationship between leaf lifespan (LL) and LMAs in the Panama dataset, afeter accounting for the effects of light (suv vs. shade leaves; see Eq. \ref{eq:LL_opt}).
The dashed line indicates the 1:1 relationship expected for residuals on the log-scale.
The posterior means of the partial correlation coefficient ($\bar{\rho}$) is shown.
The results shown here include all leaves in the Panama dataset.
](`r here::here("figs/pa_point_par_ll.png")`){#fig-LLplt}

\newpage

![
Variance partitioning on LMA components between and within leaf habits (evergreen vs. deciuous) for the GLOPNET dataset and the Panama dataset, and between and within sites (wet vs. dry) and light (sun vs. shade) for the Panama dataset.
To isolate the effects of intraspecific variation, the Panama results shown here only include species for which both sun and shade leaves were available.
](`r here::here("figs/vpart_intra.png")`){#fig-vpart}

\newpage

![
Boxplots comparing leaf mass per area (LMA) and posterior medians of photosynthetic and structural LMA components (LMAm and LMAs, respectively) across deciduous (D) and evergreen (E) leaves in the GLOPNET dataset (a) and in the Panama dataset (b).
The center line in each box indicates the median, upper and lower box edges indicate the interquartile range, whiskers show 1.5 times the interquartile range, and points are outliers.
Groups sharing the same letters are not significantly different (P > 0.05; t-tests).
The Panama results only include species for which both sun and shade leaves were available.
Qualitatively similar results were obtained when all Panama species were included (Fig. S\ref{fig-box_inter}).
Note that the vertical axis is on a log~10~ scale.
](`r here::here("figs/box_de.png")`){#fig-boxplt_de}

\newpage

![
Boxplots comparing leaf mass per area (LMA) and posterior medians of metabolic and structural LMA components (LMAm and LMAs, respectively) across sites (wet and dry) and canopy strata (sun and shade) in the Panama dataset.
These results only include species for which both sun and shade leaves were available.
Qualitatively similar results were obtained when all Panama species were included (Fig. S\ref{fig-box_inter}).
Details as for [@fig-boxplt_de].
](`r here::here("figs/box_pa.png")`){#fig-boxplt_pa}

\newpage

![
Relationships between mass dependency (*b* in @eq-var) and LMAs variance (relative to total LMA variance) for the global GLPNET datase,t un leaves in Panama, and shade leaves in Panama.
Solid lines indicate simulated means and shaded regions indicate 95% CIs.
Each point indicates the estimated values from the empirical data and represents interspecific variation (e.g., across species within a canopy position in the Panama dataset).
The y-axis values indicates if photosynthetic capacity (*A*) is is primarily mass-dependent (*b* > 0.5) or primarily area-dependent (0.5 > *b* > 0), with *b* near 0 indicating  purely area-dependendence [@Osnas2018].
If *b* > 1, then *A* increases exponentially with LMA, which is not consistent with observed relationships [@Osnas2018].
](`r here::here("figs/mass_prop_mv.png")`){#fig-massplt}

\newpage

![
Measured traits in the Panama dataset related to photosynthesis and metabolism (nitrogen and phosphorus per-unit leaf area; *N*~area~ and *P*~area~) are better correlated with estimates (posterior medians) of the metabolic LMA component (LMAm) than the strctural component (LMAs), whereas the opposite patten occurs for a measured structural trait (cellulose per-unit leaf area; CL~area~).
*N*~area~, *P*~area~, and CL~area~ data were not used to fit the models, and are presented here as independent support for the model results.
Pearson correlation coefficients (*r*) for LMA (left column) and partial correlation coefficients ($\rho$) of LMAm (middle column) and LMAs (right column) are shown.
Analogous results were obtained for *N*~area~ and *P*~area~ for GLOPNET (Fig. S\ref{fig-glnp}).
The results shown here include all leaves for which *N*~area~, *P*~area~  and *CL*~area~ are available.
](`r here::here("figs/pa_point_npc.png")`){#fig-PA-NPC}
