---
output:
  bookdown::pdf_book:
    fig_caption: true
    keep_tex: yes
    toc: no
    number_sections: no
    includes:
        in_header: format.sty
    pandoc_args:
      - "--filter"
      - "pandoc-crossref"
  bookdown::word_document2:
    fig_width: 6
    fig_caption: no
    reference_docx: rmd_style.docx
    pandoc_args:
        - "--filter"
        - "pandoc-crossref"
        - "--mathml"
fontsize: 12pt
geometry: margin=1in
link-citations: yes
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/ecology.csl
bibliography: [MyLibrary.bib]
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r}
library(methods)
library(knitr)
library(kableExtra)
library(png)
library(grid)
library(pander)
library(kfigr)
source("rmd_func.r")
```

**Running title**: Metabolic and structural leaf mass

**Decomposing leaf mass into metabolic and structural components explains divergent patterns of trait variation within and among plant species**

Masatoshi Katabuchi^1,2,6^, Kaoru Kitajima^1,3,4^, S. Joseph Wright^4^, Sunshine A. Van Bael^4,5^, Jeanne L. D. Osnas^1^ and Jeremy W. Lichstein^1^

^1^ Department of Biology, University of Florida, Gainesville, FL 32611, USA

^2^ Xishuangbanna Tropical Botanical Garden, Chinese Academy of Sciences, Menglun, Yunnan 666303 China

^3^ Graduate School of Agriculture, Kyoto University, Kitashirakawa Oiwake-Cho, Kyoto 606-8502 Japan

^4^ Smithsonian Tropical Research Institute, 9100 Panama City Pl., Washington, DC 20521

^5^ Department of Ecology and Evolutionary Biology, Tulane University, New Orleans, LA 70118 USA

^6^**Corresponding Author**: E-mail: mattocci27@gmail.com

`r page_break("Abstract")`

- Across the global flora, photosynthetic and metabolic rates depend more strongly on leaf area than leaf mass.
In contrast, intraspecific variation in these rates is strongly mass-dependent.
These contrasting patterns suggest that the causes of variation in leaf mass per area (LMA) may be fundamentally different within vs. among species.

- We used statistical methods to decompose LMA into two conceptual components – ‘metabolic LMAm (which determines photosynthetic capacity and metabolic rates, and also affects optimal leaf lifespan) and ‘structural’ LMAs (which determines leaf toughness and potential leaf lifespan) using leaf trait data from tropical forest sites in Panama and a global leaf-trait database.

- Statistically decomposing LMA into LMAm and LMAs provides improved predictions of trait variation (photosynthesis, respiration, and lifespan) across the global flora, and within and among tropical plant species in Panama.
Our analysis shows that small scaling slope between metabolic leaf mass and photosynthetic rate, and similar variance in LMAm and in LMAs leads to area-proportionality of interspecific leaf traits.
In contrast, intraspecific LMA variation is due to changes in LMAm, which explains why photosynthetic and metabolic traits are mass-dependent within species

- Our results suggest that leaf trait variation is multi-dimensional and is not well-represented by the one-dimensional leaf economics spectrum.

# Introduction

Leaf functional traits play an important role in ecological and physiological tradeoffs [@Wright2004a; @Reich2014; @Onoda2017] and in carbon and nutrient cycling (references).
Thus, understanding the causes and consequences of leaf trait variation is an important goal in ecology, plant physiology, and biogeochemistry (references).
Different leaf assemblages exhibit markedly different patterns of trait variation.
For example, across global species, whole-leaf values of traits related to photosynthesis and metabolism (e.g., the photosynthetic capacity, respiration rate, or nutrient content of an entire leaf) tend to increase with leaf area, but not leaf mass per area (LMA; @Osnas2013). In contrast, within species, these same whole-leaf trait values tend to increase with LMA from shade to full sunlight [@Osnas2013; @Osada2001].
Functional groups (e.g., deciduous vs. evergreen angiosperms) also differ from each other in terms of how interspecific trait variation depends on leaf mass and area [@Osnas2018].
These divergent patterns suggest the presence of multiple drivers of trait variation.

Strong interspecific correlations among LMA, leaf lifespan (LL), and mass-normalized leaf traits related to photosynthesis and metabolism have been interpreted as evidence for a single dominant axis of leaf functional variation [@Wright2004a].
However, several studies suggest that these strong correlations result from mass-normalization of area-dependent traits [@Lloyd2013; @Osnas2013].
Thus, although mass-normalization and the LES can be justified from economic principles [@Westoby2013], the evidence for a single dominant axis is controversial.

To understand why the same data can be interpreted as either supporting or opposing the existence of a single dominant axis of leaf trait variation, consider the conceptual model proposed by @Osnas2018, in which LMA is comprised of two additive components: structural leaf mass per area (LMAs) and photosynthetic (hereafter ‘metabolic’) leaf mass per area (LMAm).
Suppose that these two LMA components are independent axes of functional variation (Fig. 1a).
These two independent axes can be translated into either a two-dimensional trait space (if metabolic traits are area-normalized; Fig. 1b) or a one-dimensional trait space (if metabolic traits are mass-normalized; Fig. 1c).
Essentially, mass-normalization can collapse one of the functional dimensions, because leaves that differ substantially in their metabolic investment per-unit area can have similar values of mass-normalized metabolic traits (Fig. 1).

The hypothetical example in Fig. 1 shows how mass-normalization can, in principle, make a two-dimensional trait space appear one-dimensional, but this example does not resolve the dimensionality of functional variation in real leaf assemblages.
One way to better understand the dimensionality of leaf trait variation is to compare models with different numbers of dimensions, and to ask if models with multiple dimensions provide improved statistical fits and conceptual insights compared to a single axis.
For example, the two-dimensional ‘LMAm + LMAs’ model proposed by @Osnas2018 could be compared to a one-dimensional LMA model in terms of their capacities to explain variation in other traits.
Doing so would require implementing the ‘LMAm + LMAs’ model in a quantitative framework.
Implementing this model is challenging, because although certain leaf mass components can be neatly classified as ‘metabolic’ or ‘structural’ [@Poorter2009; @Osnas2018], other leaf mass components cannot.
For example, thick cell walls contribute to structural toughness (references), but at least some cell wall mass is required for the biomechanical support that enables photosynthesis (references).
One way forward is to use statistical methods to partition LMA into different components, rather than relying on direct measurements of morphological components (e.g., @John2017).

In this paper, we present a statistical modeling framework to partition LMA into metabolic and structural components: LMAm and LMAs.
We develop and test the model using leaf trait data from two tropical forest sites (sun and shade leaves from wet and dry sites in Panama) and the GLOPNET global leaf traits database [@Wright2004a].
We use the model to address the following questions: (1) How does variation in LMA components related with mass and area proportionality of leaf traits? (2) Do LMAm and LMAs differ between evergreen and deciduous species, and between sun and shade leaves? and (3) How are measurable leaf photosynthetic and structural traits (e.g., concentrations of nitrogen and cellulose) related to LMAm and LMAs?

# Material and Methods

## Model overview

We developed a statistical modeling framework to partition LMA into additive LMAs and LMAm components (see below and Supplement S3).
For sample i (e.g., sun or shade leaves of a given species; see Datasets below), we partition LMA~*i*~ into LMAm~*i*~ = *f~i~* × LMA~*i*~ and LMAs~*i*~ = (1 – *f~i~*) × LMA~*i*~ by estimating a latent variable *f~i~*.
The latent variables *f~i~* are not directly observed, but they can be constrained by available data using Bayesian methods [@Bishop2006; @Gelman2006].
For example, we would expect high values of *f~i~* (high LMAm~*i*~) for leaves with high *A*~area~, and low values of *f~i*~ (high LMAs~*i*~) for leaves with high LL.
Given the large number of free parameters (i.e., one latent variable per leaf sample), it is possible for the model to over-fit the data, which could lead to spurious inferences.
Therefore, we performed tests with randomized data (see below) and with simulated data to evaluate model performance under a range of conditions (Supplement S4). Tests with simulated data suggest that our modeling approach is robust and not prone to producing artefactual results.

## Modeling leaf lifespan, photosynthetic capacity, and dark respiration rate

We considered five sets of models, ranging from simple models with LMA as the sole predictor, to more complex models in which LMA was partitioned into metabolic and structural components (LMAm and LMAs).
The models are summarized in Table 1, and are explained in detail below.

First, we considered a simple set of models with LMA as the sole predictor for *A*~area~, *R*~area~, and LL.
In this model, the observed LMA for leaf sample i (where ‘sample’ is a species in the GLOPNET dataset, or a species $\times$ canopy position combination in the Panama dataset) are assumed to have scaling relationships with *A*~area~, *R*~area~ and LL.

Next, we considered models in which LMA is partitioned into additive metabolic and structural functions; i.e., we assume that the sum of metabolic leaf mass per area (LMAm) and structural leaf mass per area (LMAs) is equal to total observed LMA for leaf sample *i* (where a ‘sample’ is a species in the GLOPNET dataset, or a species × canopy position combination in the Panama dataset):

\begin{align}
  &\mathrm{LMA}_{i} =\mathrm{LMAm}_{i} + \mathrm{LMAs}_{i} (\#eq:LMA)\\
  &\mathrm{LMAm}_{i} = f_{i} \mathrm{LMA}_{i} (\#eq:LMAm)\\
  &\mathrm{LMAs}_{i} = (1 - f_{i})  \mathrm{LMA}_{i}(\#eq:LMAs)
\end{align}

where, *f~i~* is the fraction of LMA~*i*~ that is comprised of LMAm~*i*~.
The *f~i~* terms are not directly observed but can be estimated as latent variables in a Bayesian modeling framework (see details below).
In our model, net photosynthetic capacity, LL and total leaf dark respiration are determined by both metabolic and structural tissues:

In our model, net photosynthetic capacity (*A*~area~) is determined by LMAm, LL is determined by LMAs, and total leaf dark respiration is determined by both metabolic and structural tissues:

\begin{align}
& \mathrm{E}[A_{\mathrm{area} \, i}]
= \alpha_0\mathrm{LMAm}_{i}^{\alpha_m}\mathrm{LMAs}_i^{\alpha_s}  =  \alpha_0 (f_i \mathrm{LMA}_{i})^{\alpha_m} \bigl\{(1-f_i) \mathrm{LMA}_{i}\bigr\}^{\alpha_s} (\#eq:E-A) \\
& \mathrm{E}[\mathrm{LL}_i] = \beta_0\mathrm{LMAm}_{i}^{\beta_m} \mathrm{LMAs}_{i}^{\beta_s}   = \beta_0 (f_i \mathrm{LMA}_{i})^{\beta_m} \bigl\{(1-f_i) \mathrm{LMA}_{i}\bigr\}^{\beta_s} (\#eq:E-LL) \\
& \mathrm{E}[R_{\mathrm{area} \, i}]
= \gamma_0\mathrm{LMAm}_{i}^{\gamma_m}ı \mathrm{LMAs}_{i}^{\gamma_s}
= \gamma_0 (f_i \mathrm{LMA}_{i})^{\gamma_m}  \bigl\{(1-f_i)\mathrm{LMA}_{i}\bigr\}^{\gamma_s} (\#eq:E-R)
\end{align}

where, *A*~area~ ~*i*~, *R*~area~ ~*i*~, and LL~*i*~, are, respectively, the net photosynthetic rate (*A*~max~) per unit area, dark respiration rate (*R*~dark~) per unit area, and leaf life span of leaf *i*; $\alpha_0$, $\beta_0$, and $\gamma_0$ are fitted constants; $\alpha_m$, $\beta_m$, and $\gamma_m$ are fitted exponents that quantify the sensitivities of *A*~area~, LL and *R*~area~, respectively, to variation in LMAm; $\alpha_s$, $\beta_s$, and $\gamma_s$ are fitted exponents that quantify the sensitivities of *A*~area~, LL and *R*~area~, respectively, to variation in LMAs.
The logarithms of *A*~area~, LL, and *R*~area~ are assumed to have a multivariate normal distribution (MVN) (see supplement).

We also considered models in which LMAs in Eq. (5) was replaced by leaf structural density, LMAs/LT, where LT is leaf thickness.
Leaf density has also been shown in some cases to be a good predictor for LL [@Kitajima2012; @Kitajima2013].
In our analysis, models with leaf density yielded qualitatively similar results as those based on LMAs, but did not perform as well in cross-validation (see below).
Therefore, we only present results of the models with LMAs.

Finally, we considered models that explicitly account for the effects of light availability on LL, consistent with optimall LL theory [@Kikuzawa1991; Kikuzawa2004].
Eqs. 4-6 implicitly account for light availability, because LMA tends to increase from shade to sun within species [@Kenzo2006; @Onoda2008; @Terashima2011] that can be allocated to metabolic and/or structural functions.
However, LMA cannot account for both inter- and intraspecific variation in LL, because LL increases as LMA increases across species but decreases as LMA increases within species from shade to sun [@Lusk2008; @Russo2016] different light conditions.
Accounting for the details of leaf carbon balance – e.g., diurnal and seasonal cycles [@Xu2017] – would complicate our modeling framework, but we can account for light effects in a simple way by modifying Eq. 5 as follows:

\begin{align}
\mathrm{E[LL}_i] = \beta_0\mathrm{LMAm}_{i}^{\beta_m} \mathrm{LMAs}_{i}^{\beta_s} exp(\theta Light_i) (\#eq:E-LL2)
\end{align}

where $\theta$ is the effect of light for sun leaves and the dummy variable $Light_i$ is set to 1 for sun leaves and 0 for shade leaves.

## Datasets

To fit the models described above, we used LMA (g m^-2^), net photosynthetic capacity per unit leaf area (*A*~area~; mol s^-1^ m^-2^), dark respiration rate per unit leaf area (*R*~area~; mol s^-1^ m^-2^) and LL (months) from the GLOPNET global leaf traits database [@Wright2004a] and from two tropical forest sites in Panama: Monumental Natural Metropolitano (MNM, “dry site”) and Bosque Protector San Lorenzo (SL, “wet site”)[@Osnas2018].
The GLOPNET data primarily represent interspecific variation, whereas the Panama data represent both inter- and intraspecific variation, including leaves sampled at two canopy positions (“sun”: full sun at the top of the canopy; and “shade”: well shaded, sampled within 2 m of the forest floor) from trees within reach of  canopy cranes.
The dry MNM site is a semi-deciduous coastal Pacific forest with a 5-month dry season from December-April and 1740 mm of annual rainfall [@Wright2003].
The MNM crane is 40 m tall with a 51 m long boom. The wet SL site is an evergreen Caribbean coastal forest with 3100 mm of annual rainfall [@Wright2003]. The SL crane is 52 m tall with a 54 m long boom.

After deleting leaf samples (i.e., database records, which typically average over multiple individual leaves) that lacked one of the four traits (LMA, *A*~area~, *R*~area~, or LL), 198 samples for 198 unique species were available for GLOPNET, and 130 samples for 104 unique species were available for Panama (dry and wet sites combined; 26 species sampled in both sun and shade; no species with all four traits available at both sites). Both datasets include additional traits that we used to interpret model results, but which were not used to fit models. These traits include nitrogen and phosphorus content per leaf unit area (*N*~area~ and *P*~area~; g m^-2^) in both datasets, leaf habit in GLOPNET (deciduous or evergreen), and cellulose content per unit area (*CL*~area~; g m^-2^) in Panama.

## Model estimation

We modeled *A*~area~ and *R*~area~ using Eqs. (4) and (6), respectively, for both GLOPNET and Panam.
To model LL for GLOPNET, we used Eq. (5) (no light effects) for GL, because GLOPNET includes no information on light availability.
To model LL for Panama, we used Eq. (7) (light effects model), which was motivated by the negative intraspecific LL-LMA relationship observed in Panama [@Xu2017; @Osnas2018] and elsewhere [@Lusk2008; @Russo2016] and elsewhere.

Posterior distributions of all parameters were estimated using the Hamiltonian Monte Carlo algorithm (HMC) implemented in Stan [@Carpenter2017].
Prior distributions for the latent variables *f*~i~ (which are used to partition LMA into LMAm and LMAs according to Eqs. 1-3) were beta(0.5, 0.5), which are non-informative; i.e., LMA was partitioned based on patterns in the data.
See supplement X for more detail.
The Stan code use to fit models are also available from Github at: [https://github.com/mattocci27/XXX](https://github.com/mattocci27/LMAmLMAs).
Some parameters are exchangeable in our model.
For example, $(f_i, \alpha_m, \alpha_s)$ = (0.2, 0.5, 1.0) and (0.8, 1.0, 0.5) yield the identical likelihood, which results in bimodal distributions of the posterior distributions of parameters (See X).
To avoid this non-identifiability, we explore two methods based on our assumption that *A*~area~ depends more strongly on metabolic leaf mass (LMAm; parameter $\alpha_m$ in Eq. 4) than structural leaf mass (LMAs; parameter $\alpha_s$ in Eq. 4).
In the first method, we switched the $\alpha_m$ and $\alpha_s$ labels wihtin a chain when the posterior means for $\alpha_m$ was smaller than that for $\alpha_s$ after the posterior distributions converged.
In the second method, we constrained α_m > α_s.
Both methods yielded qualitatively similar results; we report results from the second method.
Convergence of the posterior distribution was assessed with the Gelman-Rubin statistic with a convergence threshold of 1.1 for all diagnostics [@Gelman2013].

Because our statistical approach includes one latent variable *f~i~* to partition LMA into LMAm and LMAs for each leaf sample, one might expect a good match between predictions and observations simply due to the large number of free parameters, whether or not the model captured important biological mechanisms.
To evaluate model performance while controlling for the number of free parameters, we compared the fit of our model to observed and randomized datasets.
Specifically, we generated randomized datasets by randomizing all the trait values (LMA, *A*~area~, *R*~area~ and LL) across species.
Thus, the randomized datasets had zero epected covariance among traits.
Model results obtained from the randomized datasets did not produce any significant patterns (Supplement X).
Basically, the posterior distributions of $f_i$ were changed little from the prior (i.e., all the average $f_i$ tend to be 0.5 for all the samples).
Thus, we assume that estimates of LMAm and LMAs reflect meaningful patterns in the data.

## Model selection

We applied 10-fold cross- validation to determine the best model.
Cross-validation can tell us wheter the performance on the train data is not a good indicators of predictive performance of unseen data due to the problem of over-fitting [@Bishop2006].
Because our model does not estimate the prior distributions of the latent variable *f~i~* in the train datasets, it requires additional calculation to estimate the latent variable *f~i~* in the test datasets.
Thus, less computationally intensive methods such as widely applicable information criterion (WAIC; @Watanabe2010), and Pareto-smoothed importance sampling leave-one-out cross-validation (PSIS-LOO; @Vehtari2014) are not appropriate.
In order to estimate the latent variable *f~i~*, we assigned the prior distributions to the latent variable *f~i~* in the test datasets based on the mean and variance of *f~i~* for each functional group in the train datasets.
Then, we calculated the expected out-of-sample log point-wise predictive density (ELPD) for each fold.
We select the best model based on the ELPD.

## Partitioning LMA variance into metabolic and structural components

We used the following identity to estimate the relative contributions of LMAm and LMAs to LMA variance, where again LMA = LMAm + LMAs:

\begin{align}
\mathrm{Var}(Y = X1 + X2) = \mathrm{Cov}(Y, X1+X2) = \mathrm{Cov}(Y,X1) + \mathrm{Cov}(Y,X2) (\#eq:var)
\end{align}

Thus, the fractions of total LMA variance due to variance in LMAm and LMAs were determined by the covariances Cov(LMA, LMAm) and Cov(LMA, LMAs), respectively, taken as proportions of the total variance Var(LMA).

## Understanding relationships between photosynthetic capacity and LMA

We applied our LMAm+LMAs model (Eq. 4) to simulated data to better understand relationships between photosynthetic capacity (*A*~max~) and LMA.
The causes and interpretation of *A*~max~ vs. LMA relationships are controversial [@Westoby2013].
Although *A*~max~ is often mass-normalized (e.g., @Wright2005; @Shipley2006; @Blonder2011), @Lloyd2013 argued that photosynthesis is an area-based process, and therefore *A*~max~ should be area-normalized rather than mass-normalized when exploring trait relationships.
Consistent with this argument, @Osnas2013 showed that across global species, *A*~max~ variation is strongly dependent on leaf area, but only weakly dependent on leaf mass (after controlling for interspecific variation in leaf area).

To better understand the factors affecting *A*~max~ vs. LMA relationships, we created simulated datasets in which we varied the following factors: the sensitivity of *A*~area~ to variation in LMAm (parameter $\alpha_m$ in Eq. 4); the sensitivity of *A*~area~ to LMAs (parameter $\alpha_s$ in Eq. 4), and the degree to which LMAm and LMAs varied among leaves.
For each simulated dataset, we quantified the *A*~max~ vs. LMA relationship following @Osnas2018:

\begin{align}
A_{\mathrm{area} \, i} = a (LMA_i)^{b}\epsilon_i (\#eq:mass)
\end{align}

where LMA is the sum of LMAm and LMAs (Eq. 1), *a* is a fitted constant, and *b* is an index of mass-dependence as illustrated by the following cases [@Osnas2013; @Osnas2018]: if *b* = 0, then *A*~area~ is independent of LMA, which implies that whole-leaf *A*~max~ is proportional to leaf area; conversely, if *b* = 1, then *A*~area~ is proportional to LMA, which implies that the whole-leaf *A*~max~ is proportional to leaf mass.
Intermediate cases (0 < b < 1), as well as more extreme cases (b < 0 or b > 1), are also possible.
Note that although Eq. (9) uses area-normalized *A*~max~, equivalent results can be obtained from mass-normalized *A*~max~ [@Osnas2018].

We created simulated datasets based on a range of assumptions for the LMA variance components: (i) LMAm variance << LMAs variance, (ii) LMAm variance = LMAs variance, (iii) LMAm variance >> LMAs variance, (iv) our estimates (posterior means) of LMAm and LMAs variances from GLOPNET, (v) our estimates of LMAm and LMAs variances for Panama sun leaves, and (vi) our estimates of LMAm and LMAs variances for Panama shade leaves.
For each of the above cases, we considered a range of values for  and  (which determine the sensitivity of Aarea to LMAm and LMAs in Eq.
4), as well as the empirical estimates of these values based on Glopnet and Panama datasets.
Each simulation had a sample size of 100 leaves, where the values of LMAm and LMAs for each leaf was a random draw from a lognormal distribution with median and variance as follows: for simulations (i)-(iii),  median log(40) and  variance  log(1) or log(4) (for low and high variance, respectively); and for simulations (iv)-(vi), medians and variances were the posterior means from our Glopnet and Panama analyses.

# Results

**1. Nearly all leaf dark respiration is associated with metabolic leaf tissue mass.**
Metabolic leaf mass (LMAm) accounted for nearly all leaf dark respiration; i.e., estimated dark respiration rate per-unit structural mass ($\gamma_s$) was close to zero in analyses of both GLOPNET and Panama data (Table 1)
. Thus, although building costs are likely similar for different leaf chemical components and tissues [@Williams1989; @Villar2001], our results suggest that leaf mass associated with photosynthetic function accounts for nearly all leaf maintenance respiration.

**2. Decomposing LMA into metabolic and structural components leads to improved predictions of *A*~area~, *R*~area~ and LL.**
For both the GLOPNET global datasets and the Panama datasets, model including LMAm and LMAs showed better predictions than model including LMA based on cross-validation (Table 1), suggesting that decompose LMA into metabolic and structural components leads improved predictions of *A*~area~, *R*~area~ and LL.

For the GLOPNET global dataset, *A*~area~ had a strong positive correlation with LMAm, a weak negative correlation with LMAs, and a non-significant correlation with total LMA (Figs. 1a-c; Table 2).
*R*~area~ in GLOPNET also had a strong positive correlation with LMAm, which was stronger than the correlation between *R*~area~ and either LMA or LMAs (Figs. 1d-f).
Finally, LL in GLOPNET had a strong positive correlation with LMAs, which was stronger than the correlation between LL and either LMA or LMAm (Figs. 1g-i).

For the Panama dataset, we evaluated multiple models due to the availability of both sun and shade leaves.
The model including the light effects fit the data better than the model without the light effects according to cross-validation (Table 1).
All Panama results we report are for the model including the light effects unless stated otherwise.
*A*~area~ and *R*~area~ had stronger and more positive correlations with LMAm than with LMA or LMAs (Figs. 2a-f and Table S1).
LL was not significantly correlated with LMA when all leaves were combined, but was strongly correlated with LMA for shade leaves at the dry site (Fig. 2g).
There was an apparent negative correlation between LMAm and LL (Fig. 2h) but the effect of LMAm on LL was not significant (Table 2), implying that a weak covariance between LMAm and LMAs produced this negative correlation.
There was a positive correlation between LL and LMAs (Fig. 2i and Table 2).
Because the LL vs. LMAs relationship ignores important factors (i.e., effects of light on realized photosynthetic rates), the explained variance of LL from the model was higher (r^2^ = 0.51) than expected from the correlation between LMAs and LL.

**3. Interspecific mass proportionality of *A*~max~ depended on both scaling slope and variance in LMA components.**
Percent of interspecific variation in LMA explained by structural components were 45%, 90% and 66% for GLPNET, shade leaves in Panama and sun leaves in Panama, respectively.

Large variance in LMAm and LMAs among leaves leads to mass and area proportionality of *A*~max~, respectively (Fig. \@ref(fig:massplt)a). Additionally, increasing the effects of LMAm on *A*~area~ ($\alpha_m$) leads to mass proportionality and increasing the effects of LMAs on *A*~area~ ($\alpha_s$) leads to area proportionality (Fig. \@ref(fig:massplt)a).

Interspecific variation in LMA among leaves in GLOPNET tended to be area proportional because of their negative $\alpha_s$ and small $\alpha_m$ values (Fig. \@ref(fig:massplt)b)).
Interspecific variation in LMA in sun leaves in the Panama data was area depend because of their large variance in LMAs (Fig. \@ref(fig:massplt)ab).
Sun leaves in the Panama data showed weak mass proportionality (\@ref(fig:massplt)b) instead of their relatively large variance in LMAs because of relatively large $\alpha_m$.

**4. Evergreen leaves have both greater LMAm and LMAs than deciduous leaves, and sun leaves have both greater LMAm and LMAs than shade leaves.**
In the GLOPNET dataset, altough evergreen leaves had higher both LMAm and LMAs than decidous leaves, differences in LMAs between the two groups were more aparent (Fig. 5).
Thus, the higher total LMA in evergreen leaves in GLOPNET was primarily due to differences in LMAs (which comprised a greater fraction of LMA in evergreen than in deciduous leaves; Fig. S1a).
Similar results were obtained for the Panama dataset (evergreen leaves have higher LMAs but similar LMAm compared to deciduous leaves; Fig. S1b).
In the Panama dataset, sun leaves have both greater LMAm and LMAs than shade leaves, but the differcnes in LMAm were larger (Fig. 5).
Thus, in contrast to interspecific variation in LMA (which is driven by variation in both LMAm and LMAs; Fig. 4), intraspecific variation in LMA reflects changes in LMAm.

**5. Nitrogen and phosphorus per-unit leaf area are strongly correlated with LMAm, and cellulose per-unit leaf area is strongly correlated with LMAs.**
In the GLOPNET dataset, *N*~area~ and *P*~area~ had strong positive correlations with LMAm, but only weak correlations with LMAs (Fig. S3).
Similarly, in the Panama dataset, *N*~area~ and *P*~area~ had strong positive correlations with LMAm, but were not correlated with LMAs (Fig. 6).
In contrast, cellulose per-unit leaf area (*CL*~area~), which was available for the Panama dataset but not for GLOPNET, had a strong positive correlation with LMAs, and a weak positive correlation with LMAm (Fig. 6).
*CL*~area~ was more strongly correlated with LMA than with LMAm or LMAs, but sun and shade leaves aligned along a common *CL*~area~-LMAs relationship, as opposed to being offset for LMA and LMAm (Figs. 6g-i).

# Discussion

Our analyses demonstrate that decomposing LMA variation into separate metabolic and structural components (LMAm and LMAs, respectively) leads to improved predictions of photosynthetic capacity (*A*~max~), dark respiration rate (*R*~dark~), and leaf lifespan (LL), as well as clear relationships with traits used for independent model evaluation (nitrogen, phosphorus, and cellulose concentrations). The model with LMAm and LMAs showed better predictive accuracy than model with LMA alone for both intraspecific variation in relation to light and interspecific variation. Below, we elaborate on the insights gained from our analysis and the implications of our results for the representation of leaf functional diversity in global ecosystem models.

Decomposing LMA into LMAm and LMAs provides insights into why interspecific variation in leaf traits related to photosynthesis and metabolism are primarily area-dependent (i.e., primarily independent of LMA when expressed per-unit area rather than mass-dependent [@Osnas2013; @Osnas2018]). We expected that leaf metabolic traits tend to be area-depend when variance in LMAm is larger than variance in LMAs. However, our results suggest that the scaling slopes ($\alpha_m$ and $\alpha_s$) between LMAm, LMAs and *A*~area~ also affect trait area proportionality. Leaf metabolic traits tend to be area-dependent when (i) LMAm is small or similar to variance in LMAs, (ii) the effects of LMAm and LMAs on *A*~area~ are small or combinations of those (Fig. \@ref(fig:massplt)a). Consistent with this explanation, the assemblage we examined where LMAs counted for the highest fraction of total LMA variation (90% for Panama shade leaves) is also the assembled with the highest degree of trait area-dependence [@Osnas2018]. Our results also suggest that high degree of trait area-dependence in GLOPNET is due to the negative effect of LMAs on *A*~area~.

Decomposing LMA also provides insights as to why intraspecific patterns of trait variation differ from those observed across species. In contrast to interspecific LMA variation, our analysis suggests that LMAp contributes half or more of the intraspecific increase in LMA from shade to sun (Figs. 5 and SX). The increase in LMAm from shade to sun – which likely reflects an increase in the size and number of palisade mesophyll cells with increasing light availability [@Onoda2008; @Terashima2011] – is also associated with an increase in LMAs from shade to sun (Fig. 5). This positive covariance between LMAm and LMAs within species means that per-area values of LMAm-proportional traits (e.g., *A*~area~) have a strong, positive relationship with total LMA, which implies trait mass-proportionality (Supplement S1-2).

The improved predictions and understanding provided by decomposing LMA into photosynthetic and structural components challenge the view that leaf functional diversity can be accurately represented by a single leaf economics spectrum (LES) axis [@Wright2004a]. @Lloyd2013 argued that the apparent dominance of a single LES axis is an artifact of expressing area-dependent leaf traits on a per-mass basis, and @Osnas2013 demonstrated that across the global flora, traits related to photosynthesis and metabolism are indeed area-dependent. Intraspecific patterns in trait variation, which contrast with interspecific patterns, pose additional challenges for a one-dimensional view of leaf functional diversity. Our analysis shows that considering two primary axes of leaf trait variation (photosynthesis and structure) provides improved quantitative predictions and insights compared to LES of a single dominant axis. Our results point to a simple two-dimensional framework for representing leaf functional diversity in global ecosystem models: a LMAm axis that determines *A*~max~, accounts for nearly all *R*~dark~ (see rp and rs estimates in Table S1) and a LMAs axis that determines potential LL through its effects on leaf toughness [@Kleyer2012]. In the datasets we analyzed (the global flora and tropical trees in Panama), these two axes are only weakly correlated with each other (Fig. S5), which suggests that trait-based approaches to global ecosystem modeling [@Scheiter2013; @Wullschleger2014] could consider these as independent axes. For example, to simulate individual competing trees which from diverse community of growth strategies [@Sakschewski2015; @Sakschewski2016], it would be possible to draw LMAm and LMAs values from distributions with realistic ranges instead of a single LMA distribution.

Our statistical decomposition of LMA into metabolic and structural components provides important insights, but additional insights and accuracy could be gained by a more mechanistic modeling approach. For example, @John2017 decomposed interspecific LMA variation into anatomical components such as the size, number of layers, and mass density of cells in different leaf tissues. If such detailed information became available for a large number of leaves, representing both intra- and interspecific variation, it should be possible to quantify how these anatomical traits scale up to leaf-level *A*~max~ , *R*~dark~, and LL. A simpler alternative would be to modify our model to account for variation in cell wall thickness (TCW): for a given cell size, increasing TCW would lead to an increase in lamina density, cellulose per volume, toughness, and LL [@Kitajima2010; @Kitajima2012; @Kitajima2016], and a decrease in mesophyll conductance and *A*~max~ [@Evans2009; @Terashima2011; @Onoda2017]. Indeed, our model might capture those patterns. The negative effects of LMAs on *A*~area~ in GLOPNET data indicate that high amount of photosynthetic proteins is not always translated into high photosynthetic capacity. High LMAs that may be associated with a particular anatomical components might lead to lower mesophyll conductance.


# Conclusions

It is widely recognized that LMA variation is associated with multiple tissues and functions, including metabolically active mesophyll that largely determines photosynthetic capacity, as well as structural and chemical components that contribute primarily to leaf toughness and defense [@Roderick1999; @Shipley2006; @Lusk2010]. It should not be surprising then, that partitioning LMA into metabolic and structural components yields enhanced predictions and improved understanding of patterns of leaf trait variation both within and among species. Yet for over a decade, the vast literature on leaf traits has been strongly influenced by the view that leaf trait variation can usefully be represented by a single dominant axis of LES. Our results provide quantitative evidence that this one-dimensional view of leaf trait variation is insufficient, and our model provides a biological explanation for previous statistical analyses that have demonstrated area-dependence of leaf traits across species [@Osnas2013], while also explaining mass-dependence within species. Our results suggest that small scaling slope between metabolic leaf mass and photosynthetic rate, and similar variance in LMAm and in LMAs leads to area-proportionality of interspecific leaf traits. Thus, strong interspecific correlations between LMA and mass-normalized photosynthetic capacity (and related traits, such as respiration rate, and nitrogen and phosphorus concentrations) are likely driven by mass-normalization itself, rather than any functional dependence of these photosynthetic and metabolic traits on LMA [@Lloyd2013; @Osnas2013]. In contrast, intraspecific variation in LMA is driven by coordinated changes in structural and metabolic mass components, which explains why mass-normalized photosynthetic and metabolic traits vary little from sun to shade within species [@Aranda2004; @Poorter2006b; @Niinemets2015].


# Acknowledgments

We thank Jonathan Dushoff for statistical advice and Martijn Slot for helpful comments that improved the paper. Mirna Samaniego and Milton Garica provided indispensable assistance in data collection. We thank the Smithsonian Tropical Research Institute (STRI), the Tropical Canopy Biology Program at STRI and the Andrew W. Mellon Foundation for supporting this work. MK was supported by a Postdoctoral Fellowship for Research Abroad from the Japan Society for the Promotion of Science.

# Author contributions

MK, JWL and JLDO conceived of the study; KK, SJW and SAVB contributed data; MK devised the analytical approach and performed analyses; MK and JWL wrote the first draft of the manuscript, and all authors contributed to revisions.

# References

<div id="refs"></div>

`r page_break("Tables")`


```{r modeltab, anchor = "table", tab.cap = "Five model forms compared to predict *A*~area~, LL and *R*~area~."}

tab1 <- data_frame(
  Model = c("LMA",
            "",
            "",
            "LMA + light",
            "",
            "",
            "LMAm + LMAs",
            "",
            "",
            "LMAm + LMAs + light",
            "",
            "",
            "LMAm + LMA/LT + light",
            "",
            ""
            ),
  `Expected trait values`
    = c("$\\mathrm{E}[A_{\\mathrm{area} \\, i}] = \\alpha_0\\mathrm{LMA}_{i}^{\\alpha}$",
    "$\\mathrm{E}[\\mathrm{LL}_i] = \\beta_0\\mathrm{LMA}_{i}^{\\beta}$",
    "$\\mathrm{E}[R_{\\mathrm{area} \\, i}] = \\gamma_0\\mathrm{LMA}_{i}^{\\gamma}$",

    "$\\mathrm{E}[A_{\\mathrm{area} \\, i}] = \\alpha_0\\mathrm{LMA}_{i}^{\\alpha}$",
    "$\\mathrm{E}[\\mathrm{LL}_i] = \\beta_0\\mathrm{LMA}_{i}^{\\beta}exp(\\theta Light_i)$",
    "$\\mathrm{E}[R_{\\mathrm{area} \\, i}] = \\gamma_0\\mathrm{LMA}_{i}^{\\gamma}$",

    "$\\mathrm{E}[A_{\\mathrm{area} \\, i}] = \\alpha_0\\mathrm{LMAm}_{i}^{\\alpha_m} \\mathrm{LMAs}_{i}^{\\alpha_s}$",
    "$\\mathrm{E}[\\mathrm{LL}_i] = \\beta_0\\mathrm{LMAm}_{i}^{\\beta_m} \\mathrm{LMAs}_{i}^{\\beta_s}$",
    "$\\mathrm{E}[R_{\\mathrm{area} \\, i}] = \\gamma_0\\mathrm{LMAm}_{i}^{\\gamma_m} \\mathrm{LMAs}_{i}^{\\gamma_s}$",

    "$\\mathrm{E}[A_{\\mathrm{area} \\, i}]= \\alpha_0\\mathrm{LMAm}_{i}^{\\alpha_m} \\mathrm{LMAs}_{i}^{\\alpha_s}$",
    "$\\mathrm{E}[\\mathrm{LL}_i] = \\beta_0\\mathrm{LMAm}_{i}^{\\beta_m} \\mathrm{LMAs}_{i}^{\\beta_s}exp(\\theta Light_i)$",
    "$\\mathrm{E}[R_{\\mathrm{area} \\, i}] = \\gamma_0\\mathrm{LMAm}_{i}^{\\gamma_m} \\mathrm{LMAs}_{i}^{\\gamma_s}$",

    "$\\mathrm{E}[A_{\\mathrm{area} \\, i}]= \\alpha_0\\mathrm{LMAm}_{i}^{\\alpha_m} (\\mathrm{LMAs}_{i}/\\mathrm{LT}_{i})^{\\alpha_s}$",
    "$\\mathrm{E}[\\mathrm{LL}_i] = \\beta_0\\mathrm{LMAm}_{i}^{\\beta_m} (\\mathrm{LMAs}_{i}/\\mathrm{LT}_{i})^{\\beta_s}exp(\\theta Light_i)$",
    "$\\mathrm{E}[R_{\\mathrm{area} \\, i}] = \\gamma_0\\mathrm{LMAm}_{i}^{\\gamma_m} \\mathrm{LMAs}_{i}^{\\gamma_s}$"
  ))

if ("docx" == knitr::opts_knit$get("rmarkdown.pandoc.to")) {
  knitr::kable(tab1, format = "markdown") #%>%
} else if ("html" == knitr::opts_knit$get("rmarkdown.pandoc.to")) {
  knitr::kable(tab1, format = "html", escape = FALSE)
} else {
  knitr::kable(tab1, format = "latex")
}

```



```{r elpdtab, anchor = "table", tab.cap = "Summary of the model comparison based on exact 10-fold cross-validation. Higher expected log point-wise predictive density (ELPD) indicates better predictive accuracy. N; number of samples."}

GLtab <- read_csv("../data/GL_elpd.csv")

GLtab2 <- GLtab %>%
  mutate(LMA = ifelse(LMA == "LMAms",
         "LMAm and LMAs", LMA)) %>%
  arrange(elpd %>% desc) %>%
  mutate(elpd = round(elpd, 0)) %>%
  mutate(elpd = as.character(elpd)) %>%
  mutate(N = as.character(N)) %>%
  bind_rows(data.frame(LMA = "**GLOPNET**", elpd = "", N = "", site = ""),
            .)

PAtab <- read_csv("../data/PA_elpd.csv")

PAtab2 <- PAtab %>%
  mutate(LMA = ifelse(LMA == "LMAms",
         "LMAm + LMAs", LMA)) %>%
  mutate(LMA = ifelse(LMA == "LMAms_L",
         "LMAm + LMAs + light", LMA)) %>%
  mutate(LMA = ifelse(LMA == "LMA_L",
         "LMA + light", LMA)) %>%
  mutate(LMA = ifelse(LMA == "LD_L",
         "LMAm + LMAs/LT + light", LMA)) %>%
  arrange(elpd %>% desc) %>%
  mutate(elpd = round(elpd, 0)) %>%
  mutate(elpd = as.character(elpd)) %>%
  mutate(N = as.character(N)) %>%
  bind_rows(data.frame(LMA = "**Panama**", elpd = "", N = "", site = ""),
            .)

tab3 <- bind_rows(GLtab2, PAtab2) %>%
  rename(Model = LMA) %>%
  rename(ELPD = elpd) %>%
  dplyr::select(-site)

colnames(tab3) <- c("",
                    "ELPD",
                    "N")

if ("docx" == knitr::opts_knit$get("rmarkdown.pandoc.to")) {
  knitr::kable(tab3, format = "markdown") #%>%
} else if ("html" == knitr::opts_knit$get("rmarkdown.pandoc.to")) {
  knitr::kable(tab3, format = "html", escape = FALSE)
} else {
  knitr::kable(tab3, format = "latex")
}

```

`r page_break("")`


```{r paratab, anchor = "table", tab.cap = "Posterior medians [95% credible interval] of parameters for the best models. Bold values are significantly different from zero based on the 95% credible interval."}

s_GL <- read_csv("../data/GLpara.csv")

GLtab <- s_GL %>%
  filter(para == "amas[1]" |
         para == "amas[2]" |
         para == "z[4]" |
         para == "z[5]" |
         para == "z[6]" |
         para == "z[7]") %>%
  mutate(tmp = c(2, 1, 3, 4, 5, 6)) %>%
  arrange(tmp) %>%
  mutate(mid = round(`X50.`, 3)) %>%
  mutate(low = round(`X2.5.`, 3)) %>%
  mutate(up = round(`X97.5.`, 3)) %>%
  mutate(est = paste0(mid, " [", low, ", ", up, "]")) %>%
  mutate(para = c("Effect of LMAm on *A*~area~ ($\\alpha_m$)",
                  "Effect of LMAs on *A*~area~ ($\\alpha_s$)",
                  "Effect of LMAm on LL ($\\beta_m$)",
                  "Effect of LMAs on LL ($\\beta_s$)",
                  "Effect of LMAm on *R*~area~ ($\\gamma_m$)",
                  "Effect of LMAs on *R*~area~ ($\\gamma_s$)"
                  ))  %>%
  dplyr::select(para, GLOPNET = est, sig1 = sig)


s_PA <- read_csv("../data/PApara.csv")

PAtab <- s_PA %>%
  filter(para == "am" |
         para == "as" |
         para == "z[8]" |
         para == "z[4]" |
         para == "z[5]" |
         para == "z[6]" |
         para == "z[7]") %>%
  mutate(tmp = c(3, 5, 4, 6, 7, 1, 2)) %>%
  arrange(tmp) %>%
  mutate(mid = round(`X50.`, 3)) %>%
  mutate(low = round(`X2.5.`, 3)) %>%
  mutate(up = round(`X97.5.`, 3)) %>%
  mutate(est = paste0(mid, " [", low, ", ", up, "]")) %>%
  mutate(para = c("Effect of LMAm on *A*~area~ ($\\alpha_m$)",
                  "Effect of LMAs on *A*~area~ ($\\alpha_s$)",
                  "Effect of LMAm on LL ($\\beta_m$)",
                  "Effect of LMAs on LL ($\\beta_s$)",
                  "Effect of LMAm on *R*~area~ ($\\gamma_m$)",
                  "Effect of LMAs on *R*~area~ ($\\gamma_s$)",
                  "Effect of light on LL ($\\theta$)"
                  ))  %>%
  dplyr::select(para, Panama = est, sig2 = sig)


GLPAtab <- full_join(GLtab, PAtab, by = "para") %>%
  dplyr::rename(Parameters = para) %>%
  mutate(GLOPNET = ifelse(sig1 == "sig",
                        paste0("**", GLOPNET, "**"),
                        GLOPNET)) %>%
  mutate(Panama = ifelse(sig2 == "sig",
                        paste0("**", Panama, "**"),
                        Panama)) %>%
  dplyr::select(Parameters, GLOPNET, Panama)

GLPAtab[is.na(GLPAtab)] <- "-"

if ("docx" == knitr::opts_knit$get("rmarkdown.pandoc.to")) {
  knitr::kable(GLPAtab, format = "markdown") #%>%
} else if ("html" == knitr::opts_knit$get("rmarkdown.pandoc.to")) {
  knitr::kable(GLPAtab, format = "html", escape = FALSE)
} else {
  knitr::kable(GLPAtab, format = "latex")
}

```

`r page_break("Figures")`

```{r GLplt, echo = F, dpi = 300, fig.width = 6, fig.height = 6, fig.cap = "Observed and estimated leaf-trait relationships in the global GLOPNET dataset. Leaf life span (LL), net photosynthetic rate per unit leaf area (*A*~area~), and dark respiration rate per unit leaf area (*R*~area~) are plotted against observed LMA and estimates (posterior means) of photosynthetic and structural LMA components (LMAm and LMAs, respectively). Pearson correlation coefficients are for LMA (left column) and posterior distributions of LMAm (middle column) and LMAs (right column)."}

img <- readPNG("../figs/GL_2.png", native = TRUE, info = TRUE)
grid.raster(img, interpolate=F)

```

`r page_break()`

```{r PAplt, echo = F, dpi = 300, fig.width = 6, fig.height = 6, fig.cap = "Observed and estimated leaf-trait relationships in the Panama dataset. Estimates are from the Optimal LL Model (Eq. 8). Details as in Fig. 1. Results for other LL models are summarized in Table SX."}

img <- readPNG("../figs/PA_2.png", native = TRUE, info = TRUE)
grid.raster(img, interpolate=F)
```

`r page_break()`

```{r LLplt, echo = F, dpi = 300, fig.width = 6, fig.height = 6, fig.cap = "Observed vs. predicted leaf lifespan (LL) in the Optimal LL Model (Eq. 8). Predicted LL values are posterior medians. The dashed line indicates the 1:1 relationship. The *r*^2^ value is the posterior median of Bayesian *r*^2^ (Gelman et al. 2017)."}

img <- readPNG("../figs/LL_plot.png", native = TRUE, info = TRUE)
grid.raster(img, interpolate=F)

```

`r page_break()`

```{r massplt, echo = F, dpi = 300, fig.width = 6, fig.height = 3, fig.cap = "The relationships among mass proportionality (*b~i~* in Eq. X), variance in LMAm and LMAs, and scaling slopes between LMAm and *A*~area~ ($\\alpha_m$) and LMAs and *A*~area~ ($\\alpha_s$). Blue colors indicate the photosynthetic rate is primarily mass proportional (*b~i~* is large) and red colors indicate the photosynthetic rate is primarily area proportional (*b~i~* is close to 0).**(a)** Simulated mass proportionality gradients for different degrees of LMA components and scaling slopes. **(b)** Simulated mass proportionality gradients based on the estimated LMA components and scaling slopes for the global GLOPNET dataset, sun leaves in Panama, and shade leaves in Panama."}

img <- readPNG("../figs/mass_prop4.png", native = TRUE, info = TRUE)
grid.raster(img, interpolate=F)

```

`r page_break()`

```{r boxplt, echo = F, dpi = 300, fig.width = 6, fig.height = 6, fig.cap = "Boxplots comparing leaf mass per area (LMA) and posterior means of photosynthetic and structural LMA components (LMAm and LMAs, respectively) across deciduous (D) and evergreen (E) leaves in the GLOPNET dataset (top), and across sites (wet and dry) and canopy strata (sun and shade) in the Panama dataset (bottom). The center line in each box indicates the median, upper and lower box edges indicate the interquartile range, whiskers show 1.5 times the interquartile range, and points are outliers. Groups sharing the same letters are not significantly different (P > 0.05; t-tests). To isolate the effects of intraspecific variation (i.e., plastic responses to light), the Panama results shown here only include species for which both sun and shade leaves were available. Qualitatively similar results were obtained when all Panama species were included (Fig. S2). Note that the vertical axis is on a log~10~ scale."}

img <- readPNG("../figs/box_main.png", native = TRUE, info = TRUE)
grid.raster(img, interpolate=F)
```

`r page_break()`

```{r PA-NPC, echo = F, dpi = 300, fig.width = 6, fig.height = 6, fig.cap = "Measured traits related to photosynthesis and metabolism traits (nitrogen and phosphorus per-unit leaf area; *N*~area~ and *P*~area~) are strongly correlated with estimates (posterior means) of the photosynthetic LMA component (LMAm), and a measured structural trait (cellulose per-unit leaf area; *CL*~area~) is strongly correlated with estimates of the structural LMA component (LMAs) for the Panama dataset. Note that sun and shade leaves align along a single relationship for *CL*~area~ vs. LMAs, but not for *CL*~area~ vs. LMA or LMAm. *N*~area~, *P*~area~, and *CL*~area~ data were not used to fit the models, and are presented here as independent support for the model results. Analogous results were obtained for *N*~area~ and *P*~area~ vs. LMAm for GLOPNET (Fig. S3). Others details as in Fig. GL. Results for other LL models are reported in Table SX."}

img <- readPNG("../figs/PA_NPC.png", native = TRUE, info = TRUE)
grid.raster(img, interpolate=F)
```
