---
output:
  bookdown::pdf_document2:
    fig_width: 6
    fig_caption: yes
    number_sections: no
    keep_tex: yes
    includes:
        in_header: format.sty
    pandoc_args:
        - "--filter"
        - "pandoc-crossref"
 # bookdown::html_document2:
 #   fig_width: 6
 #   fig_caption: yes
 #   number_sections: no
 #   includes:
 #       in_header: format.sty
 #   pandoc_args:
 #       - "--filter"
 #       - "pandoc-crossref"
 #       - "--mathml"
figPrefix: Fig.
figureTitle: '**Figure**'
tblPrefix: Table
tableTitle: 'Table'
---

# hoge

tlmgr search --file pcrr8t.tfm --global

- hoge
- hoge


```{r modeltab, eval=FALSE}
source("rmd_func.r")
library(tidyverse)
library(kableExtra)

tab1 <- read_csv("tab1.csv")
tab1 <- tab1 |>
  dplyr::select(`Expected trait values`)

#colnames(tab1)[1] <- ""

if ("docx" == knitr::opts_knit$get("rmarkdown.pandoc.to")) {
  knitr::kable(tab1, format = "markdown") #%>%
} else {
  kbl(tab1, escape = FALSE, booktabs = TRUE, longtable = FALSE) |>
    kable_paper() |>
    pack_rows("Model 1: LMA", 1, 3) |>
    pack_rows("Model 2: LMA + light", 4, 6) |>
    pack_rows("Model 3: LMAp + LMAs", 7, 9) |>
    pack_rows("Model 4: LMAp + LMAs + light", 10, 12) |>
    pack_rows("Model 5: LMAp + LMAs/LT + light", 13, 15)
}

```

\newpage

```{r, eval=FALSE}
kbl(mtcars[1:10, 1:6], caption = "Group Rows", booktabs = T) %>% kable_styling() %>%
  pack_rows("Group 1", 4, 7) %>%
  pack_rows("Group 2", 8, 10)
```

\newpage

| Model | Name                   | Expected trait values                                                                                                             |
|-------|------------------------|-----------------------------------------------------------------------------------------------------------------------------------|
| 1     | LMA                    | $\mathrm{E}[A_{\mathrm{area} \, i}] = \alpha_0\mathrm{LMA}_{i}^{\alpha}$                                                          |
|       |                        | $\mathrm{E}[\mathrm{LL}_i] = \beta_0\mathrm{LMA}_{i}^{\beta}$                                                                     |
|       |                        | $\mathrm{E}[R_{\mathrm{area} \, i}] = \gamma_0\mathrm{LMA}_{i}^{\gamma}$                                                          |
| 2     | LMA + light            | $\mathrm{E}[A_{\mathrm{area} \, i}] = \alpha_0\mathrm{LMA}_{i}^{\alpha}$                                                          |
|       |                        | $\mathrm{E}[\mathrm{LL}_i] = \beta_0\mathrm{LMA}_{i}^{\beta}exp(\theta Light_i)$                                                  |
|       |                        | $\mathrm{E}[R_{\mathrm{area} \, i}] = \gamma_0\mathrm{LMA}_{i}^{\gamma}$"                                                         |
| 3     | LMAp + LMAs            | $\mathrm{E}[A_{\mathrm{area} \, i}] = \alpha_0\mathrm{LMAp}_{i}^{\alpha_p} \mathrm{LMAs}_{i}^{\alpha_s}$                          |
|       |                        | $\mathrm{E}[\mathrm{LL}_i] = \beta_0\mathrm{LMAp}_{i}^{\beta_p} \mathrm{LMAs}_{i}^{\beta_s}$                                      |
|       |                        | $\mathrm{E}[R_{\mathrm{area} \, i}] = \gamma_0\mathrm{LMAp}_{i}^{\gamma_p} \mathrm{LMAs}_{i}^{\gamma_s}$                          |
| 4     | LMAp + LMAs + light    | $\mathrm{E}[A_{\mathrm{area} \, i}]= \alpha_0\mathrm{LMAp}_{i}^{\alpha_p} \mathrm{LMAs}_{i}^{\alpha_s}$                           |
|       |                        | $\mathrm{E}[R_{\mathrm{area} \, i}] = \gamma_0\mathrm{LMAp}_{i}^{\gamma_p} \mathrm{LMAs}_{i}^{\gamma_s}$                          |
|       |                        | $\mathrm{E}[\mathrm{LL}_i] = \beta_0\mathrm{LMAp}_{i}^{\beta_p} \mathrm{LMAs}_{i}^{\beta_s}exp(\theta Light_i)$                   |
| 5     | LMAp + LMAs/LT + light | $\mathrm{E}[A_{\mathrm{area} \, i}]= \alpha_0\mathrm{LMAp}_{i}^{\alpha_p} (\mathrm{LMAs}_{i}/\mathrm{LT}_{i})^{\alpha_s}$         |
|       |                        | $\mathrm{E}[\mathrm{LL}_i] = \beta_0\mathrm{LMAp}_{i}^{\beta_p} (\mathrm{LMAs}_{i}/\mathrm{LT}_{i})^{\beta_s}exp(\theta Light_i)$ |
|       |                        | $\mathrm{E}[R_{\mathrm{area} \, i}] = \gamma_0\mathrm{LMAp}_{i}^{\gamma_p} \mathrm{LMAs}_{i}^{\gamma_s}$                          |
